---
title: "``boxTidwell`` Box-Tidwell Transformations"
author: "Kevin O Brien"
date: "21 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```


####Description

Computes the Box-Tidwell power transformations of the predictors in a linear model.

####Usage
<pre><code>

boxTidwell(y, ...)
## S3 method for class 'formula'
boxTidwell(formula, other.x=NULL, data=NULL, subset,
na.action=getOption("na.action"), verbose=FALSE, tol=0.001,
max.iter=25, ...)
## Default S3 method:
boxTidwell(y, x1, x2=NULL, max.iter=25, tol=0.001,
verbose=FALSE, ...)
## S3 method for class 'boxTidwell'
print(x, digits=getOption("digits") - 2, ...)

</code></pre>

####Arguments

formula two-sided formula, the right-hand-side of which gives the predictors to be transformed.
other.x one-sided formula giving the predictors that are not candidates for transformation, including (e.g.) factors.

* ``data  `` : an optional data frame containing the variables in the model. By default the variables are taken from the environment from which boxTidwell is called.
* ``subset `` :  an optional vector specifying a subset of observations to be used.
* na.action a function that indicates what should happen when the data contain NAs. The default is set by the na.action setting of options.
* verbose if TRUE a record of iterations is printed; default is FALSE.
* tol if the maximum relative change in coefficients is less than tol then convergence is declared.
* `` max.iter ``:   maximum number of iterations.
* y response variable.
* x1 matrix of predictors to transform.
* x2 matrix of predictors that are not candidates for transformation.
* ... not for the user.
* x boxTidwell object.
* digits number of digits for rounding.


####Examples

```{r}

boxTidwell(prestige ~ income + education, ~ type + poly(women, 2), data=Prestige)
```
####Details

The maximum-likelihood estimates of the transformation parameters are computed by Box and Tidwell's
(1962) method, which is usually more efficient than using a general nonlinear least-squares
routine for this problem. Score tests for the transformations are also reported.
Value
an object of class boxTidwell, which is normally just printed.
Author(s)
John Fox <jfox@mcmaster.ca>

####References

Box, G. E. P. and Tidwell, P.W. (1962) Transformation of the independent variables. Technometrics
4, 531-550.
Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.
Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.

