---
title: "``brief`` Print Abbreviated Ouput"
author: "Kevin O Brien"
date: "22 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```

####Description

Print data objects and statistical model summaries in abbreviated form.

####Usage
<pre><code>

brief(object, ...)
## S3 method for class 'data.frame'
brief(object, rows = if (nr <= 10) c(nr, 0) else c(3, 2),
cols, head=FALSE, tail=FALSE, elided = TRUE,
classes = inherits(object, "data.frame"), ...)
## S3 method for class 'matrix'
brief(object, rows = if (nr <= 10) c(nr, 0) else c(3, 2), ...)
## S3 method for class 'numeric'
brief(object, rows = c(2, 1), elided = TRUE, ...)
## S3 method for class 'integer'
brief(object, rows = c(2, 1), elided = TRUE, ...)
## S3 method for class 'character'
brief(object, rows = c(2, 1), elided = TRUE, ...)
## S3 method for class 'factor'
brief(object, rows=c(2, 1), elided=TRUE, ...)
## S3 method for class 'list'
brief(object, rows = c(2, 1), elided = TRUE, ...)
## S3 method for class 'function'
brief(object, rows = c(5, 3), elided = TRUE, ...)
## S3 method for class 'lm'
brief(object, terms = ~ .,
intercept=missing(terms), pvalues=FALSE,
digits=3, horizontal=TRUE, vcov.=vcov, ...)
## S3 method for class 'glm'
brief(object, terms = ~ .,
intercept=missing(terms), pvalues=FALSE,
digits=3, horizontal=TRUE, vcov., dispersion, exponentiate, ...)
## S3 method for class 'multinom'
brief(object, terms = ~ .,
intercept=missing(terms), pvalues=FALSE,
digits=3, horizontal=TRUE, exponentiate=TRUE, ...)
26 brief
## S3 method for class 'polr'
brief(object, terms = ~ .,
intercept, pvalues=FALSE,
digits=3, horizontal=TRUE, exponentiate=TRUE, ...)
## Default S3 method:
brief(object, terms = ~ .,
intercept=missing(terms), pvalues=FALSE,
digits=3, horizontal=TRUE, ...)

</code></pre>

####Arguments

* `` object ``:    a data or model object to abbreviate.
rows for a matrix or data frame, a 2-element integer vector with the number of rows to print at the beginning and end of the display; for a vector or factor, the number
of lines of output to show at the beginning and end; for a list, the number of elements to show at the beginning and end; for a function, the number of lines to show at the beginning and end.
cols for a matrix or data frame, a 2-element integer vector with the number of columns to print at the beginning (i.e., left) and end (right) of the display.
head, tail alternatives to the rows argument; if TRUE, print the first or last 6 rows; can also be the number of the first or last few rows to print; only one of heads and tails should be specified; ignored if FALSE (the default).
* elided controls whether to report the number of elided elements, rows, or columns;
default is TRUE.
classes show the class of each column of a data frame at the top of the column; the classes are shown in single-character abbreviated form-e.g., [f] for a factor,
[i] for an integer variable, [n] for a numeric variable, [c] for a character variable.
* ``terms`` a one-sided formula giving the terms to summarize; the default is ~ .-i.e., to
summarize all terms in the model.
intercept whether or not to include the intercept; the default is TRUE unless the terms argument is given, in which case the default is FALSE; ignored for polr models.
pvalues include the p-value for each coefficient in the table; default is FALSE.
exponentiate for a "glm" or "glmerMod" model using the log or logit link, or a "polr"
or "multinom" model, show exponentiated coefficient estimates and confidence
bounds.
* ``digits`` significant digits for printing.
* ``horizontal`` if TRUE (the default), orient the summary produced by brief horizontally, which typically saves space.
dispersion see summary.glm
* ``vcov.`` either a matrix giving the estimated covariance matrix of the estimates, or a function that when called with object as an argument returns an estimated covariance matrix of the estimates. The default of vcov. = vcov uses the usual estimated covariance matrix. Other choices include the functions documented
at hccm. For the glm method, if the vcov. or dispersion argument is specified, then Wald-based confidence limits are computed; otherwise the reported confidence limits are computed by profiling the likelihood. NOTE: The dispersion and vcov. arguments may not both be specified.
* ``...`` arguments to pass down.

#### Value
Invisibly returns object for a data object, or summary for a model object.
Note
The method brief.matrix calls brief.data.frame.
#### Author(s)
John Fox <jfox@mcmaster.ca>

####References

Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.
See Also
S

####Examples

```{r}

brief(rnorm(100))
brief(Duncan)
brief(OBrienKaiser, elided=TRUE)
brief(matrix(1:500, 10, 50))
brief(lm)
mod.prestige <- lm(prestige ~ education + income + type, Prestige)
brief(mod.prestige, pvalues=TRUE)
brief(mod.prestige, ~ type)
mod.mroz <- glm(lfp ~ ., data=Mroz, family=binomial)
brief(mod.mroz)
```
