---
title: "``compareCoefs``: Print estimated coefficients and their standard errors in a table for
several regression models."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```




####Description

This function extracts estimates of regression parameters and their standard errors from one or more
models and prints them in a table.

####Usage
<pre><code>

compareCoefs(..., se = TRUE, zvals = FALSE, pvals = FALSE, vcov.,
print = TRUE, digits = 3)

</code></pre>

####Arguments

... One or more regression-model objects. These may be of class lm, glm, nlm,
or any other regression method for which the functions coef and vcov return
appropriate values, or if the object inherits from the mer class created by the lme4 package or lme in the nlme package.
* se If TRUE, the default, show standard errors as well as estimates.
* zvals If TRUE (the default is FALSE), print Wald statistics, the ratio of each coefficient to its standard error.
* pvals If codeTRUE (the default is FALSE), print two-sided p-values from the standard normal distribution corresponding to the Wald statistics.
* `` vcov.  ``:   an optional argument, specifying a function to be applied to all of the models,
returning a coefficient covariance matrix for each, or a list with one element for each model, with each element either containing a function to be applied to the corresponding model or a coefficient covariance matrix for that model. If omitted, vcov is applied to each model.
* print If TRUE, the defualt, the results are printed in a nice format using printCoefmat.
If FALSE, the results are returned as a matrix
* digits Passed to the printCoefmat function for printing the result.
Value
This function is mainly used for its side-effect of printing the result. It also invisibly returns a matrix
of estimates, standard errors, Wald statistics, and p-values.

####Examples

```{r}

mod1 <- lm(prestige ~ income + education, data=Duncan)
mod2 <- update(mod1, subset=-c(6,16))
mod3 <- update(mod1, . ~ . + type)
mod4 <- update(mod1, . ~ . + I(income + education)) # aliased coef.
compareCoefs(mod1)
compareCoefs(mod1, mod2, mod4)
compareCoefs(mod1, mod2, mod3, zvals=TRUE, pvals=TRUE)
compareCoefs(mod1, mod2, se=FALSE)
compareCoefs(mod1, mod1, vcov.=list(vcov, hccm))
```
#### Author(s)
John Fox <jfox@mcmaster.ca>

####References

Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.

