---
title: "`` hccm``:  Heteroscedasticity-Corrected Covariance Matrices"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```

####Description

Calculates heteroscedasticity-corrected covariance matrices linear models fit by least squares or
weighted least squares. These are also called “White-corrected” or “White-Huber” covariance matrices.
hccm 55

####Usage
<pre><code>

hccm(model, ...)
## S3 method for class 'lm'
hccm(model, type=c("hc3", "hc0", "hc1", "hc2", "hc4"),
singular.ok=TRUE, ...)
## Default S3 method:
hccm(model, ...)

</code></pre>

####Arguments

model a unweighted or weighted linear model, produced by lm.
type one of "hc0", "hc1", "hc2", "hc3", or "hc4"; the first of these gives the classic
White correction. The "hc1", "hc2", and "hc3" corrections are described in
Long and Ervin (2000); "hc4" is described in Cribari-Neto (2004).
singular.ok if FALSE (the default is TRUE), a model with aliased coefficients produces an
error; otherwise, the aliased coefficients are ignored in the coefficient covariance
matrix that’s returned.
... arguments to pass to hccm.lm.




####Examples

```{r}

options(digits=4)
mod<-lm(interlocks~assets+nation, data=Ornstein)
vcov(mod)
```

```{r}
hccm(mod)

```
####Details

The original White-corrected coefficient covariance matrix ("hc0") for an unweighted model is
V (b) = (X0X)??1X0diag(e2i
)X(X0X)??1
where e2i
are the squared residuals, and X is the model matrix. The other methods represent adjustments
to this formula. If there are weights, these are incorporated in the corrected covariance
matrix.
The function hccm.default simply catches non-lm objects.
Value
The heteroscedasticity-corrected covariance matrix for the model.
#### Author(s)
John Fox <jfox@mcmaster.ca>


####References

* Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.
Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.
* Cribari-Neto, F. (2004) Asymptotic inference under heteroskedasticity of unknown form. Computational
Statistics and Data Analysis 45, 215–233.
* Long, J. S. and Ervin, L. H. (2000) Using heteroscedasity consistent standard errors in the linear
regression model. The American Statistician 54, 217–224.

* White, H. (1980) A heteroskedastic consistent covariance matrix estimator and a direct test of heteroskedasticity.
Econometrica 48, 817–838.