---
title: "``invTranPlot``:  invTranPlot Choose a Predictor Transformation Visually or Numerically"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```
``invTranPlot``:   Choose a Predictor Transformation Visually or Numerically

####Description

invTranPlot draws a two-dimensional scatterplot of Y versus X, along with the OLS fit. 
nvTranEstimate finds the nonlinear least squares estimate and its standard error.

####Usage
<pre><code>

invTranPlot(x, ...)
## S3 method for class 'formula'
invTranPlot(x, data, subset, na.action, id=FALSE, ...)

## Default S3 method:
invTranPlot(x, y, lambda=c(-1, 0, 1), robust=FALSE,
  lty.lines=rep(c("solid", "dashed", "dotdash", "longdash", "twodash"),
  length=1 + length(lambda)), lwd.lines=2,
  col=carPalette()[1], col.lines=carPalette(),
  xlab=deparse(substitute(x)), ylab=deparse(substitute(y)),
  family="bcPower", optimal=TRUE, key="auto", id=FALSE,
  grid=TRUE, ...)
  
invTranEstimate(x, y, family="bcPower", confidence=0.95, robust=FALSE)

</code></pre>

####Arguments

* x The predictor variable, or a formula with a single response and a single predictor
* y The response variable
* data An optional data frame to get the data for the formula
* subset Optional, as in lm, select a subset of the cases
* na.action Optional, as in lm, the action for missing data
* lambda The powers used in the plot. The optimal power than minimizes the residual sum of squares is always added unless optimal is FALSE.
* robust If TRUE, then the estimated transformation is computed using Huber M-estimation with the MAD used to estimate scale and k=1.345. The default is FALSE.
* family The transformation family to use, "bcPower", "yjPower", or a user-defined family.
* confidence returns a profile likelihood confidence interval for the optimal transformation with this confidence level. If FALSE, or if robust=TRUE, no interval is returned.
* optimal Include the optimal value of lambda?
* lty.lines line types corresponding to the powers
* lwd.lines the width of the plotted lines, defaults to 2 times the standard col color(s) of the points in the plot. If you wish to distinguish points according to the levels of a factor, we recommend using symbols, specified with the pch argument, rather than colors.
* col.lines color of the fitted lines corresponding to the powers. The default is to use the colors returned by carPalette
key The default is "auto", in which case a legend is added to the plot, either above the top marign or in the bottom right or top right corner. Set to NULL to suppress the legend.
xlab Label for the horizontal axis.
ylab Label for the vertical axis.
id controls point identification; if FALSE (the default), no points are identified; can be a list of named arguments to the showLabels function; TRUE is equivalent to list(method=list(method="x", n=2, cex=1, col=carPalette()[1], location="lr"),
which identifies the 2 points with the most extreme horizontal values-i.e., the response variable in the model.
... Additional arguments passed to the plot method, such as pch.
* grid If TRUE, the default, a light-gray background grid is put on the graph


####Examples

```{r}
with(UN, invTranPlot(ppgdp, infantMortality))
```

```{r}
with(UN, invTranEstimate(ppgdp, infantMortality))
```

#### Value
invTranPlot plots a graph and returns a data frame with _ in the first column, and the residual sum
of squares from the regression for that _ in the second column.
invTranEstimate returns a list with elements lambda for the estimate, se for its standard error,
and RSS, the minimum value of the residual sum of squares.

#### Author(s)
Sanford Weisberg, <sandy@umn.edu>

#### References

Fox, J. and Weisberg, S. (2011) An R Companion to Applied Regression, Second Edition, Sage.
Prendergast, L. A., & Sheather, S. J. (2013) On sensitivity of inverse response plot estimation and
the benefits of a robust estimation approach. Scandinavian Journal of Statistics, 40(2), 219-237.
Weisberg, S. (2014) Applied Linear Regression, Fourth Edition, Wiley, Chapter 7.

#### See Also
inverseResponsePlot,optimize