---
title: "``leveneTest`` Levene's Test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```
####Description

Computes Levene's test for homogeneity of variance across groups.

####Usage
<pre><code>

leveneTest(y, ...)
## S3 method for class 'formula'
leveneTest(y, data, ...)
## S3 method for class 'lm'
leveneTest(y, ...)
## Default S3 method:
leveneTest(y, group, center=median, ...)

</code></pre>

####Arguments

* y response variable for the default method, or a lm or formula object. If y is a linear-model object or a formula, the variables on the right-hand-side of the model must all be factors and must be completely crossed.
* group factor defining groups.
* center The name of a function to compute the center of each group; mean gives the original Levene's test; the default, median, provides a more robust test.
* data a data frame for evaluating the formula.
* ... arguments to be passed down, e.g., data for the formula and lm methods; can also be used to pass arguments to the function given by center (e.g., center=mean and trim=0.1 specify the 10% trimmed mean).

####Examples

```{r}

with(Moore, leveneTest(conformity, fcategory))
with(Moore, leveneTest(conformity, interaction(fcategory, partner.status)))
leveneTest(conformity ~ fcategory*partner.status, data=Moore)
leveneTest(lm(conformity ~ fcategory*partner.status, data=Moore))
leveneTest(conformity ~ fcategory*partner.status, data=Moore, center=mean)
leveneTest(conformity ~ fcategory*partner.status, data=Moore, center=mean, trim=0.1)
```

#### Value
returns an object meant to be printed showing the results of the test.

#### Note
adapted from a response posted by Brian Ripley to the r-help email list.

#### Author(s)
John Fox <jfox@mcmaster.ca>; original generic version contributed by Derek Ogle

####References

Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.
Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.
