---
title: "``logit``:  Logit Transformation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```


####Description

Compute the logit transformation of proportions or percentages.

####Usage

<pre><code>

logit(p, percents=range.p[2] > 1, adjust)

</code></pre>

####Arguments

* ``p``: numeric vector or array of proportions or percentages.
* ``percents``: TRUE for percentages.
* ``adjust``: adjustment factor to avoid proportions of 0 or 1; defaults to 0 if there are no such
proportions in the data, and to 0.025 if there are.


#### Examples

```{r}
options(digits=4)
logit(.1*0:10)
```

```{r}
logit(.1*0:10, adjust=0)
```

####Details

Computes the logit transformation logit = log[p=(1 ?? p)] for the proportion p.
If p = 0 or 1, then the logit is undefined. logit can remap the proportions to the interval (adjust, 1 - adjust) prior to the transformation. If it adjusts the data automatically, logit
will print a warning message.

#### Value

a numeric vector or array of the same shape and size as p.

#### Author(s)

John Fox <jfox@mcmaster.ca>

####References

Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.
See Also
probabilityAxis


