---
title: "``qqPlot``:  Quantile-Comparison Plot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```



####Description

Plots empirical quantiles of a variable, or of studentized residuals from a linear model, against
theoretical quantiles of a comparison distribution. Includes options not avaiable in the qqnorm
function.

####Usage
<pre><code>

qqPlot(x, ...)
qqp(...)
## Default S3 method:
qqPlot(x, distribution="norm", groups, layout,
ylim=range(x, na.rm=TRUE), ylab=deparse(substitute(x)),
xlab=paste(distribution, "quantiles"), glab=deparse(substitute(groups)),
main=NULL, las=par("las"),
envelope=.95, col=carPalette()[1], col.lines=carPalette()[2],
lwd=2, pch=1, cex=par("cex"), line=c("quartiles", "robust", "none"), id=TRUE, grid=TRUE, ...)

## S3 method for class 'formula'
qqPlot(formula, data, subset, id=TRUE, ylab, glab, ...)

## S3 method for class 'lm'
qqPlot(x, xlab=paste(distribution, "Quantiles"),
ylab=paste("Studentized Residuals(",
deparse(substitute(x)), ")", sep=""),
main=NULL, distribution=c("t", "norm"),
line=c("robust", "quartiles", "none"), las=par("las"),
simulate=TRUE, envelope=.95, reps=100,
col=carPalette()[1], col.lines=carPalette()[2], lwd=2, pch=1, cex=par("cex"),
id=TRUE, grid=TRUE, ...)

</code></pre>

####Arguments

x vector of numeric values or lm object.
distribution root name of comparison distribution – e.g., "norm" for the normal distribution;
t for the t-distribution.
groups an optional factor; if specified, a QQ plot will be drawn for x within each level
of groups.
layout a 2-vector with the number of rows and columns for plotting by groups – for
example c(1, 3) for 1 row and 3 columns; if omitted, the number of rows
and columns will be selected automatically; the specified number of rows and
columns must be sufficient to accomodate the number of groups; ignored if there
is no grouping factor.
formula one-sided formula specifying a single variable to be plotted or a two-sided formula
of the form variable ~ factor, where a QQ plot will be drawn for
variable within each level of factor.
data optional data frame within which to evaluage the formula.
subset optional subset expression to select cases to plot.
ylim limits for vertical axis; defaults to the range of x. If plotting by groups, a common
y-axis is used for all groups.
* ylab label for vertical (empirical quantiles) axis.
* xlab label for horizontal (comparison quantiles) axis.
glab label for the grouping variable.
main label for plot.
envelope confidence level for point-wise confidence envelope, or FALSE for no envelope.
las if 0, ticks labels are drawn parallel to the axis; set to 1 for horizontal labels (see
par).
* `` col  ``:  color for points; the default is the first entry in the current car palette (see
carPalette and par).
98 qqPlot
col.lines color for lines; the default is the second entry in the current car palette.
pch plotting character for points; default is 1 (a circle, see par).
cex factor for expanding the size of plotted symbols; the default is 1.
id controls point identification; if FALSE, no points are identified; can be a list of
named arguments to the showLabels function; TRUE is equivalent to list(method="y", n=2, cex=1, col=which identifies the 2 points with the 2 points with the most extreme verical values
— studentized residuals for the "lm" method. Points labels are by default
taken from the names of the variable being plotted is any, else case indices are
used. Unlike most graphical functions in car, the default is id=TRUE to include
point identification.
lwd line width; default is 2 (see par).
line "quartiles" to pass a line through the quartile-pairs, or "robust" for a robustregression
line; the latter uses the rlm function in the MASS package. Specifying
line = "none" suppresses the line.
simulate if TRUE calculate confidence envelope by parametric bootstrap; for lm object
only. The method is due to Atkinson (1985).
reps integer; number of bootstrap replications for confidence envelope.
... arguments such as df to be passed to the appropriate quantile function.
grid If TRUE, the default, a light-gray background grid is put on the graph

####Examples

```{r}

x<-rchisq(100, df=2)
qqPlot(x)
qqPlot(x, dist="chisq", df=2)
qqPlot(~ income, data=Prestige, subset = type == "prof")
qqPlot(income ~ type, data=Prestige, layout=c(1, 3))
qqPlot(lm(prestige ~ income + education + type, data=Duncan),
envelope=0.99)
```

####Details

Draws theoretical quantile-comparison plots for variables and for studentized residuals from a linear
model. A comparison line is drawn on the plot either through the quartiles of the two distributions,
or by robust regression.
Any distribution for which quantile and density functions exist in R (with prefixes q and d, respectively)
may be used. When plotting a vector, the confidence envelope is based on the SEs of the
order statistics of an independent random sample from the comparison distribution (see Fox, 2008).
Studentized residuals from linear models are plotted against the appropriate t-distribution with a
point-wise confidence envelope computed by default by a parametric bootstrap, as described by
Atkinson (1985). The function qqp is an abbreviation for qqPlot.
Value
These functions return the labels of identified points, unless a grouping factor is employed, in which
case NULL is returned invisibly.
#### Author(s)
John Fox <jfox@mcmaster.ca>

####References

* Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.
Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.
Atkinson, A. C. (1985) Plots, Transformations, and Regression. Oxford.
recode 99
See Also
qqplot, qqnorm, qqline, showLabels

