---
title: "``regLine`` Plot Regression Line"
author: "Kevin O Brien"
date: "21 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```


####Description

Plots a regression line on a scatterplot; the line is plotted between the minimum and maximum
x-values.

####Usage
<pre><code>

regLine(mod, col=carPalette()[2], lwd=2, lty=1,...)

</code></pre>

####Arguments

mod a model, such as produced by lm, that responds to the coef function by returning
a 2-element vector, whose elements are interpreted respectively as the intercept
and slope of a regresison line.
* `` col  ``:  color for points and lines; the default is the second entry in the current car
palette (see carPalette and par).
lwd line width; default is 2 (see par).
lty line type; default is 1, a solid line (see par).
... optional arguments to be passed to the lines plotting function.



####Examples

```{r}

plot(repwt ~ weight, pch=c(1,2)[sex], data=Davis)
regLine(lm(repwt~weight, subset=sex=="M", data=Davis))
regLine(lm(repwt~weight, subset=sex=="F", data=Davis), lty=2)
```
####Details

In contrast to abline, this function plots only over the range of the observed x-values. The x-values
are extracted from mod as the second column of the model matrix.
Value
NULL. This function is used for its side effect: adding a line to the plot.
Author(s)
John Fox <jfox@mcmaster.ca>
See Also
abline, lines