---
title: "``ScatterplotSmoothers ``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```



### Smoothers to Draw Lines on Scatterplots

####Description

These smoothers are used to draw nonparametric-regression lines on scatterplots produced by the
scatterplot, scatterplotMatrix and several other car functions. The functions are not meant
to be called directly by the user, although the user can supply options via the smoother.args
argument, the contents of which vary by the smoother (see 
details below). The gamLine smoother
uses the ``gam()`` function in the mgcv package, the loessLine smoother uses the ``loess()`` function in the
stats package, and the quantregLine smoother uses the ``rqss()`` function in the quantreg package.


####Usage
<pre><code>

gamLine(x, y, col=carPalette()[1], log.x=FALSE, log.y=FALSE, var=FALSE, spread=var,
smoother.args=NULL, draw=TRUE, offset=0)
loessLine(x, y, col=carPalette()[1], log.x=FALSE, log.y=FALSE, var=FALSE, spread=var,
smoother.args=NULL, draw=TRUE, offset=0)
quantregLine(x, y, col=carPalette()[1], log.x=FALSE, log.y=FALSE, var=FALSE, spread=var,
smoother.args=NULL, draw=TRUE, offset=0)

</code></pre>

####Arguments

x horizontal coordinates of points.
y vertical coordinates of points.
col line color.
log.x should be set to TRUE (default is FALSE) if the horizontal axis is logged.
log.y should be set to TRUE (default is FALSE) if the vertical axis is logged.
spread, var the default is to plot only an estimated mean or median function. If either of these arguments is TRUE, then a measure of variability is also plotted.
smoother.args additional options accepted by the smoother, in the form of a list of named values (see details  below).
draw if TRUE, the default, draw the smoother on the currently active graph. If FALSE,
return a list with coordinates x and y for the points that make up the smooth and if requested x.pos, y.pos, x.neg, y.neg for the spread smooths.
offset For use when spread=TRUE, the vertical axis is sqrt(offset^2 + variance smooth).

####Examples

```{r}

scatterplot(prestige ~ income, data=Prestige)
scatterplot(prestige ~ income, data=Prestige, smooth=list(smoother=gamLine))
scatterplot(prestige ~ income, data=Prestige,
smooth=list(smoother=quantregLine))
scatterplot(prestige ~ income | type, data=Prestige)
scatterplot(prestige ~ income | type, data=Prestige,
smooth=list(smoother=gamLine))
scatterplot(prestige ~ income | type, data=Prestige,
smooth=list(smoother=quantregLine))
scatterplot(prestige ~ income | type, data=Prestige, smooth=FALSE)
scatterplot(prestige ~ income | type, data=Prestige, spread=TRUE)
scatterplot(prestige ~ income | type, data=Prestige,
smooth=list(smoother=gamLine), spread=TRUE)
scatterplot(prestige ~ income | type, data=Prestige,
smooth=list(smoother=quantregLine), spread=TRUE)
scatterplot(weight ~ repwt | sex, spread=TRUE, data=Davis,
smooth=list(smoother=loessLine))
scatterplot(weight ~ repwt | sex, spread=TRUE, data=Davis,
smooth=list(smoother=gamLine)) # messes up
scatterplot(weight ~ repwt | sex, spread=TRUE, data=Davis,
smooth=list(smoother=quantregLine)) # robust


set.seed(12345)
w <- 1 + rpois(100, 5)
x <- rnorm(100)

p <- 1/(1 + exp(-(x + 0.5*x^2)))
y <- rbinom(100, w, p)
scatterplot(y/w ~ x, smooth=list(smoother=gamLine, family="binomial",
weights=w))
scatterplot(y/w ~ x, smooth=list(smoother=gamLine, family=binomial,
link="probit", weights=w))
scatterplot(y/w ~ x, smooth=list(smoother=loessLine), reg=FALSE)
y <- rbinom(100, 1, p)
scatterplot(y ~ x, smooth=list(smoother=gamLine, family=binomial))
```
####Details

The function loessLine is a re-implementation of the loess smoother that was used in car prior
to September 2012. The main enhancement is the ability to set more arguments through the
smoother.args argument.
The function gamLine is more general than the loess fitting because it allows fitting a generalized
additive model using splines. You can specify an error distribution and link function.
The function quantregLine fits an additive model using splines with estimation based on L1 regression
for the median and quantile regression if you ask for the spread. It is likely to be more
robust than the other smoothers.
The argument smoother.args is a list of named elements used to pass additional arguments to the
smoother. As of November, 2016, the smoother is evaluated at an equally spaced grid of 50 points
in the range of the horizontal variable. With any of the smoothers you can change to say 100 evaluation
points by using the argument smoother.args=list(evaluation=100). As of version 3.0-1,
the arguments in smoother.args col.var, lty.var and lwd.var are equivalent to col.spread,
lty.spread and lwd.spread, respectively.
For loessLine the default value is smoother.args=list(lty.smooth=1, lwd.smooth=2, lty.spread=4, lwd.spread=2, The arguments lty.smooth and lwd.smooth are the type and width respectively of the mean or median
smooth, lty.spread and lwd.spred are the type and color of the spread smooths if requested.
ScatterplotSmoothers 123
The arguments span, degree and family are passed to the loess function, iterations=4 robustness
iterations.
For gamLine the default is smoother.args=list(lty.smooth=1, lwd.smooth=2, lty.spread=4, lwd.spread=2,k=-1, The first four arguments are as for loessLine. The next two arguments are passed to the gam function
to control the smoothing: k=-1 allows gam to choose the number of splines in the basis function;
bs="tp" provides the type of spline basis to be used with "tp" for the default thin-plate splines.
The last three arguments allow providing a family, link and weights as in generalized linear models.
See examples below. The spread argument is ignored unless family="gaussian" and link=NULL.
For quantregLine the default is smoother.args=list(lty.smooth=1, lwd.smooth=2, lty.spread=4, lwd.spread=2,The first four arguments are as for loessLine. The last argument is passed to the qss function in
quantreg. It is a smoothing parameter, here a robust estimate of the scale of the horizontal axis
variable. This is an arbitrary choice, and may not work well in all circumstances.
#### Author(s)
John Fox <jfox@mcmaster.ca> and Sanford Weisberg <sandy@umn.edu>.
See Also
scatterplot, scatterplotMatrix, gam, loess, and rqss.

