---
title: "``vif``: Variance Inflation Factors"
author: "Kevin O Brien"
date: "22 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```

####Description

Calculates variance-inflation and generalized variance-inflation factors for linear, generalized linear, and other models.

####Usage
<pre><code>

vif(mod, ...)
## Default S3 method:
vif(mod, ...)
## S3 method for class 'merMod'
vif(mod, ...)

</code></pre>

####Arguments
* ``mod``: for the default method, an object that responds to coef, vcov, and model.matrix, such as an lm or glm object.
* ``...``: not used.


####Examples

```{r}

vif(lm(prestige ~ income + education, data=Duncan))
vif(lm(prestige ~ income + education + type, data=Duncan))
```

####Details

If all terms in an unweighted linear model have 1 df, then the usual variance-inflation factors are
calculated.
If any terms in an unweighted linear model have more than 1 df, then generalized variance-inflation
factors (Fox and Monette, 1992) are calculated. These are interpretable as the inflation in size of
the confidence ellipse or ellipsoid for the coefficients of the term in comparison with what would
be obtained for orthogonal data.
The generalized vifs are invariant with respect to the coding of the terms in the model (as long as
the subspace of the columns of the model matrix pertaining to each term is invariant). To adjust for
the dimension of the confidence ellipsoid, the function also prints GV IF1=(2_df) where df is the
degrees of freedom associated with the term.
Through a further generalization, the implementation here is applicable as well to other sorts of
models, in particular weighted linear models, generalized linear models, and mixed-effects models.
Value
A vector of vifs, or a matrix containing one row for each term in the model, and columns for the
GVIF, df, and GV IF1=(2_df).
#### Author(s)
Henric Nilsson and John Fox <jfox@mcmaster.ca>

####References

* Fox, J. and Monette, G. (1992) Generalized collinearity diagnostics. JASA, 87, 178-183.
* Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.
Fox, J. and Weisberg, S. (2018) An R Companion to Applied Regression, Third Edition, Sage.
140 wcrossprod

