---
title: "``wcrossprod`` Weighted Matrix Crossproduct"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```

####Description

Given matrices x and y as arguments and an optional matrix or vector of weights, w, return a
weighted matrix cross-product, t(x) w y. If no weights are supplied, or the weights are constant,
the function uses crossprod for speed.

####Usage
<pre><code>

wcrossprod(x, y, w)

</code></pre>

####Arguments

x,y x, y numeric matrices; missing(y) is taken to be the same matrix as x. Vectors
are promoted to single-column or single-row matrices, depending on the context.
w A numeric vector or matrix of weights, conformable with x and y.


####Examples

```{r}

set.seed(12345)
n <- 24
drop <- 4
sex <- sample(c("M", "F"), n, replace=TRUE)
x1 <- 1:n
x2 <- sample(1:n)
extra <- c( rep(0, n - drop), floor(15 + 10 * rnorm(drop)) )
y1 <- x1 + 3*x2 + 6*(sex=="M") + floor(10 * rnorm(n)) + extra
y2 <- x1 - 2*x2 - 8*(sex=="M") + floor(10 * rnorm(n)) + extra
# assign non-zero weights to 'dropped' obs

wt <- c(rep(1, n-drop), rep(.2,drop))
X <- cbind(x1, x2)
Y <- cbind(y1, y2)
wcrossprod(X)
wcrossprod(X, w=wt)
wcrossprod(X, Y)
wcrossprod(X, Y, w=wt)
wcrossprod(x1, y1)
wcrossprod(x1, y1, w=wt)
```

#### Value
A numeric matrix, with appropriate dimnames taken from x and y.

#### Author(s)
Michael Friendly, John Fox <jfox@mcmaster.ca>

#### See Also
crossprod
