

\section*{Dummy variables}

*  When an explanatory variable is categorical we can use \textbf{dummy variables} to contrast
the different categories. 
*  For each variable we choose a \textbf{baseline category} and then
contrast all remaining categories with the base line.
*  If an explanatory variable
has $k$ categories, we need $k-1$ dummy variables to investigate all the differences in
the categories with respect to the dependent variable.


For example suppose the explanatory variable was \textbf{\textit{housing}} coded like this:

	* [1:] Owner occupier
	* [2:] renting from a private landlord
	* [3:] renting from the local authority


\noindent We would therefore need to choose a baseline category and create two dummy
variables. For example if we chose owner occupier as the baseline category we
would code the dummy variables (\texttt{House1} and \texttt{House2}) like this

\begin{center}
\begin{tabular}{|l|c|c|} \hline 
	Tenure: &House1 &House2\\ \hline \hline
	Owner occupier &0& 0\\ \hline
	Rented private &1 &0\\ \hline
	Rented local authority &0 &1\\ \hline
\end{tabular}

-----------------------------------------------------------------------------------

\section*{One-Hot Encoding}
One hot encoding is a process by which categorical variables are converted into a form that could be provided to ML algorithms to do a better job in prediction.
\begin{verbatim}
╔════════════╦═════════════════╦════════╗ 
║ CompanyName Categoricalvalue ║ Price  ║
╠════════════╬═════════════════╣════════║ 
║ VW         ╬      1          ║ 20000  ║
║ Acura      ╬      2          ║ 10011  ║
║ Honda      ╬      3          ║ 50000  ║
║ Honda      ╬      3          ║ 10000  ║
╚════════════╩═════════════════╩════════╝
\end{verbatim}

One hot encoding is a process by which categorical variables are converted into a form that could be provided to ML algorithms to do a better job in prediction.

Example: Suppose you have ‘flower’ feature which can take values ‘daffodil’, ‘lily’, and ‘rose’. One hot encoding converts ‘flower’ feature to three features, ‘is\_daffodil’, ‘is\_lily’, and ‘is\_rose’ which all are binary.
