---
title: "Untitled"
author: "Kevin O'Brien"
date: "1 November 2023"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy  = FALSE
                     ,cache = FALSE
                     ,message = FALSE
                     ,warning = FALSE
                     ,fig.height =  8
                     ,fig.width  = 11)

#library(conflicted)
#library(tidyverse)
library(scales)
library(ggplot2)
library(survival)
library(survminer)
#library(muhaz)
library(broom)
```

# Estimations of the Survival Functions

## The Kaplan-Meier Estimator

Kaplan-Meier is the standard method for estimating the survival function of a
given dataset. Formally, it is defined as follows

$$
\hat{S}(t) = \prod_{t_i \leq t} (1 - \hat{q}_i) = \prod_{t_i \leq t} \left(1 - \frac{d_i}{n_i}\right)
$$

where $n_i$ is the number of subjects at risk at time $t$, and $d_i$ is the
number of individuals who fail at that time.

### Using Kaplan-Meier

In R, we construct KM estimators using the `survfit()` function.

Before we move on to our datasets, we start with a small set of data.

```{r km_estimator_intro, echo=TRUE}
tt   <- c(7, 6, 6, 5, 2, 4)
cens <- c(0, 1, 0, 0, 1, 1)
grp  <- c(0, 0, 0, 1, 1, 1)

Surv(tt, cens)

sample_tbl <- tibble(tt = tt, cens = cens, grp = grp)

example_km <- survfit(Surv(tt, cens) ~ 1, data = sample_tbl, conf.type = 'log-log')

plot(example_km)
```

Basic plotting routines are worth trying, but the `survminer` package has
specialised plots that use `ggplot2` to create them.

```{r km_estimator_survminer, echo=TRUE}
ggsurvplot(example_km)
```

Printing out the 'fitted' object gives us some basic statistics:

```{r km_estimator_print, echo=TRUE}
example_km %>% print()
```

We get more details from the `summary()` function:

```{r km_estimator_summary, echo=TRUE}
example_km %>% summary()
```

#### Exercises

  1. Construct the KM estimator for the telco churn data
  1. What is the median survival time for this data?
  1. What is the mean survival time?
  1. Repeat the above for the other two datasets.