---
title: Friedman Test
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Friedman Rank Sum Test

Performs a Friedman rank sum test with unreplicated blocked data.

<tt>friedman.test</tt> can be used for analyzing unreplicated complete block designs (i.e., there is exactly one observation in **y** for each combination of levels of groups and blocks) where the normality assumption may be violated.

The null hypothesis is that apart from an effect of **blocks**, the location parameter of **y** is the same in each of the groups.

### warpbreaks

* The number of Breaks in Yarn during Weaving
* This data set gives the number of warp breaks per loom, where a loom corresponds to a fixed length of yarn.

A data frame with 54 observations on 3 variables.

*	***breaks*** (numeric) The number of breaks
*	***wool***	(factor) The type of wool (A or B)
* ***tension***	(factor) The level of tension (L, M, H)

There are measurements on 9 looms for each of the six types of warp (AL, AM, AH, BL, BM, BH).

```{r}
head(warpbreaks)

```

```{r fig.width=7, fig.height=6}
wb <- aggregate(warpbreaks$breaks,
                by = list(w = warpbreaks$wool,
                          t = warpbreaks$tension),
                FUN = mean)
wb
```

### Formulations

```{r}
friedman.test(wb$x, wb$w, wb$t)
```

### Alternative Formulations

```{r}
friedman.test(x ~ w | t, data = wb)
```
