---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Wilcoxon Mann-Whitney Test
The Wilcoxon Mann-Whitney Test is one of the most powerful of the nonparametric tests for comparing two populations. It is used to test the null hypothesis that two populations have identical distribution functions against the alternative hypothesis that the two distribution functions differ only with respect to ***location*** (i.e. median), if at all.

The Wilcoxon Mann-Whitney test does not require the assumption that the differences between the two samples are normally distributed.
In many applications, the Wilcoxon Mann-Whitney Test is used in place of the two sample t-test when the normality assumption is questionable.
This test can also be applied when the observations in a sample of data are ranks, that is, ordinal data rather than direct measurements.

#### Remark 
- Non-parametric procedures often give warning errors when tied values occur. In this module you may disregard these.


#### {Example}
In the data frame column \textit{\textbf{mpg}} of the \textit{\textbf{mtcars}} data set , there are gas mileage data of various 1974 U.S. automobiles.

<code>
> mtcars$mpg 
 [1] 21.0 21.0 22.8 21.4 18.7 ...
<\code> 

Meanwhile, another data column in mtcars, named am, indicates the transmission type of the automobile model (0 = automatic, 1 = manual). In other words, it is the differentiating factor of the transmission type.
<code>
> mtcars$am 
 [1] 1 1 1 0 0 0 0 0 ...
<\code> 
In particular, the gas mileage data for manual and automatic transmissions are independent.

#### Exercise
Without assuming the data to have normal distribution, decide at .05 significance level if the gas mileage data of manual and automatic transmissions in mtcars have identical data distribution.

The null hypothesis is that the gas mileage data of manual and automatic transmissions are identical populations. 
To test the hypothesis, we apply the ``wilcox.test()`` function to compare the independent samples. 

```{r}
wilcox.test(mpg ~ am, data=mtcars) 
```

<pre><code>
        Wilcoxon rank sum test with continuity correction 
 
data:  mpg by am 
W = 42, p-value = 0.001871 
alternative hypothesis: true location shift is not equal to 0 
 
..............
<\code></pre>

As the p-value turns out to be 0.001817, and is less than the .05 significance level, we reject the null hypothesis.
At 0.05 significance level, we conclude that the gas mileage data of manual and automatic transmissions in mtcar are nonidentical populations.

#### Boxplot to accompany Wilcoxon Test Analysis 
<pre>
<code>
boxplot(mpg~am,data=mtcars,horizontal=TRUE,
    col=c("lightblue","lightpink"))
title("Miles Per Gallon by Transmission Type")
<\code>
</pre>
