---
title: "CEL Data Set Visualizations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

cel <- drop_na(read_csv(url("https://www.dropbox.com/s/4ebgnkdhhxo5rac/cel_volden_wiseman%20_coursera.csv?raw=1")))

```

### Plot 1

The purpose of this visualization is assess the change in
the ideology of the members of Congress (as measured by the DW-Nominate
score) over a time period of the last 10 Congresses included in the dataset. A series of boxplots is created for both parties. The $i$-th box will show the
distribution of the `dwnom1` variable corresponding to the $i$-th Congress.

* Plot Type: Boxplots factted by Political party



```{r exercise1,echo=FALSE}


  
# Get only the data corresponding to the last 10 Congresses.
# Select only the `congress`, `dwnom1` and `dem` columns.
prep_plot_1 <- cel %>%
    filter(congress > 105) %>%
    select(congress, dwnom1, dem) %>%
  rename("Party" =dem)
# Recode the `congress` variable:
prep_plot_1<- prep_plot_1 %>% mutate( congress = recode(
    congress,
    `106` = "106th",    `107` = "107th",
    `108` = "108th",    `109` = "109th",
    `110` = "110th",    `111` = "111th",
    `112` = "112th",    `113` = "113th",    
    `114` = "114th",    `115` = "115th"
), Party = recode(
     Party,
    `0` = "Republican",    `1` = "Democrat"
))

# Plot the figure:

ggplot(prep_plot_1, aes(x = congress, y = dwnom1, fill = Party)) +
    geom_boxplot() +
    labs(
        x = "Congress",y = "Ideology (DW-Nominate Score)",
        title = "Change in Ideology of Members of Congress",
        subtitle = "Variations over last 10 congresses by party",
        caption = "Source: CEL Data Set"
    ) +
  geom_hline(yintercept=0, linetype="dashed", color = "black") + 
    facet_wrap(.~Party) +
    theme_bw(base_size = 16) +
    theme(
        plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 14, face = "bold"),
        plot.caption = element_text(color = "gray", face = "italic"),
        legend.direction = "horizontal",
        legend.position = "top",
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.text.x = element_text(angle = 90)
    ) +
    scale_fill_manual(
        values = c("Democrat" = "#0000FF", "Republican" = "#FF0000")
    )
```
