---
title: "Diamonds Data set"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Draft

### ggplot2 visualizations with the Diamonds Data set

```{r warning=FALSE,echo=FALSE,message=FALSE}
library(dplyr)
library(magrittr)
library(ggplot2)

data(diamonds)

```


### Make a Subset of the Data

```{r}
diamonds2 <- diamonds %>% filter(color %in% c("D","E","F"))

diamonds.report <- diamonds2 %>% 
  group_by(cut,color)  %>% 
  summarize(mean.depth = mean(depth)) 
```

```{r}
diamonds.report 
```

```{r}
p <- ggplot(data=diamonds.report, 
       aes(x = cut,
           y = mean.depth, 
           fill = color)) 
```

```{r}
p+ geom_bar(stat="identity",position = "dodge") + 
         scale_fill_brewer(palette = "Greens" , direction = -1) +
         ggtitle("Diamonds") + 
         ylab("Mean Depth") + 
         theme_bw() +
         theme(axis.title.x = element_text(color="black", size=14, face="bold"),
               axis.title.y = element_text(color="black", size=14, face="bold"))
```

```{r}              
 p+ geom_bar(stat="identity",position = "stack") + 
         scale_fill_brewer(palette = "Greens" , direction = -1) +
         ggtitle("Diamonds") + 
         ylab("Mean Depth") + 
         theme_bw() +
         theme(axis.title.x = element_text(color="black", size=14, face="bold"),
               axis.title.y = element_text(color="black", size=14, face="bold"))
```

The Diamonds Data Set
=============================

First we will install ggplot2 and then have a quick look at the in-built data set; ***diamonds***.
```{r} 
# install.packages("ggplot2")
library(ggplot2)
data(diamonds)
head(diamonds)
``` 
Let's construct a scatter-plot of two of the numeric variables.
```{r} 
qplot(diamonds$cut, diamonds$carat)
``` 


library(dplyr)
library(magrittr)
library(ggplot2)

data(diamonds)


diamonds2 <- diamonds %>% filter(color %in% c("D","E","F"))

diamonds.report <- diamonds2 %>% group_by(cut,color)  %>% summarize(mean.depth = mean(depth)) 

p <- ggplot(data=diamonds.report, 
       aes(x = cut,
           y = mean.depth, 
           fill = color)) 

p+ geom_bar(stat="identity",position = "dodge") + 
         scale_fill_brewer(palette = "Greens" , direction = -1) +
         ggtitle("Volume by Species and BAU") + 
         ylab("Total volume (1000s) ") + 
         theme_bw() +
         theme(axis.title.x = element_text(color="black", size=14, face="bold"),
               axis.title.y = element_text(color="black", size=14, face="bold"))
              
 p+ geom_bar(stat="identity",position = "stack") + 
         scale_fill_brewer(palette = "Greens" , direction = -1) +
         ggtitle("Volume by Species and BAU") + 
         ylab("Total volume (1000s) ") + 
         theme_bw() +
         theme(axis.title.x = element_text(color="black", size=14, face="bold"),
               axis.title.y = element_text(color="black", size=14, face="bold"))

This code does the same thing, with slightly different syntax. Importantly, a dataframe must be specified.
```{r}
qplot(carat, price, data = diamonds)
``` 
We can use some of the categorical variables to gain more insight into the data. We are going to use some different numeric variables,
```{r}
qplot(carat, price, data = diamonds, colour=clarity)
``` 

```{r}
qplot(carat, price, data = diamonds, geom=c("point", "smooth"), method=lm)
 
``` 
```{r}
 
qplot(diamonds$cut, diamonds$carat)
``` 
```{r}
qplot(carat, price, data = diamonds)
``` 
```{r}
qplot(carat, price, data = diamonds, colour=clarity)
qplot(carat, price, data = diamonds, geom=c("point", "smooth"), method=lm)
 
 
 ``` 
```{r}
qplot(carat, data = diamonds¨  geom="histogram")
``` 
```{r} 
qplot(carat, data = diamonds¨  geom="histogram", binwidth = 1)
``` 
```{r}
qplot(carat, data = diamonds¨  geom="histogram", binwidth = 0.1)
``` 
```{r} 
qplot(carat, data = diamonds¨  geom="histogram", binwidth = 0.01)
 
 

 
``` 
#### Using `ggplot()`
```{r} 
d <- ggplot(diamonds, aes(x=carat, y=price))
d + geom_point()
``` 
```{r} 
d + geom_point(aes(colour = carat))
``` 
```{r} 
d + geom_point(aes(colour = carat)) + scale_colour_brewer()
``` 
```{r} 
 
ggplot(diamonds) + geom_histogram(aes(x=price)) 
 ``` 
```{r}
 
p <- ggplot(diamonds, aes(x=price))
``` 
```{r}
p + geom_histogram()
p + stat_bin(geom="area")
``` 
```{r}
p + stat_bin(geom="point")
``` 
```{r}
p + stat_bin(geom="line")
 
``` 
```{r}
p + geom_histogram(aes(fill = clarity))
``` 
```{r}
p + geom_histogram(aes(y = ..density..))
``` 
```{r}
 
 
Setting vs mapping
 
 
p <- ggplot(diamonds, aes(x=carat,y=price))
 
# What will this do?
``` 
```{r}
p + geom_point(aes(colour = "green"))
p + geom_point(colour = "green")
p + geom_point(colour = colour)


p <- ggplot(diamonds, aes(carat, price, fill = )) +
  stat_density2d(aes(fill = ..density..), contour = F, geom = "tile") + 
  scale_x_log2()+
  scale_y_log10()

p + scale_fill_gradient(high = "black", low = "white")
