---
title: "Smoothed Regression Fits with ggplot2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

### How to smooth R data in ggplot2

The ggplot2 package also makes it very easy to create regression lines through your data. You use the <tt>stat_smooth()</tt> function to create this type of line.

The interesting thing about <tt>stat_smooth()</tt> is that it makes use of local regression by default. R has several functions that can do this, but ggplot2 uses the loess() function for local regression. 

This means that if you want to create a linear regression model you have to tell <tt>stat_smooth()</tt> to use a different smoother function. You do this with the method argument.

### Unemployment Dataset 

To illustrate the use of a smoother, start by creating a scatterplot of unemployment in the longley dataset:

```{r} 
ggplot(longley, aes(x=Year, y=Employed)) + geom_point()
```
Next, add a smoother. This is as simple as adding stat_smooth() to your line of code.

```{r}
ggplot(longley, aes(x=Year, y=Employed)) +
 geom_point() + stat_smooth()
```

Finally, tell stat_smooth to use a linear regression model. You do this by adding the argument method="lm".

```{r}
ggplot(longley, aes(x=Year, y=Employed)) +
  geom_point() + stat_smooth(method="lm")
```
