---
title: "trelliscopejs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(plotly)
library(trelliscopejs)
```

### {trelliscopejs}

The trelliscopejs R package is an interface for creating Trelliscope displays with the goal of providing methods that are convenient to use in traditional visual / analytical workflows. 

For example, the package provides methods that allow you to easily generate Trelliscope displays from ggplot2 or “tidy” workflows.


### mpg data
This data set contains a subset of the fuel economy data.It contains only models which had a new release every year between 1999 and 2008 .
```{r}
head(mpg)
```


```{r}
ggplot2:: ggplot(data = mpg)+aes(x=displ,y= hwy) +geom_point()
```

### Sample Plot

This is an interactive plot that does not run in an R Markdown environment.

```{r eval=FALSE}
ggplot(mpg,aes(displ,hwy))+
  geom_point(size=4)+
  geom_smooth(se=FALSE,span=1)+
  facet_trelliscope(~manufacturer,ncol=4,nrow=3)
```
