---
title: "Using the`lme` function - Orthodont"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Fitting LME Models

The fitted model is an object of class `lme` representing the linear mixed-effects model fit. Generic functions such as `print`, `plot` and `summary` have methods to show the results of the fit. 

### Orthodont

* The Orthodont data frame has 108 rows and 4 columns of the change in an orthdontic measurement over time for several young subjects.

* Investigators at the University of North Carolina Dental School followed the growth of 27 children (16 males, 11 females) from age 8 until age 14. 

* Every two years they measured the distance between the pituitary and the pterygomaxillary fissure, two points that are easily identified on x-ray exposures of the side of the head.

### Set Up

* Load the {nlme} R package and the Orthodont data set.

* Inspect the data set by looking at the last 6 cases.

```{r}
library(nlme)
data(Orthodont)
tail(Orthodont)
```

### Baseline LM Model

```{r}
lm1 <- lm(distance ~ age, data = Orthodont)
summary(lm1)
```


```{r}
AIC(lm1)
```

### Simple LME Model

```{r}
fm1 <- lme(distance ~ age, data = Orthodont,method="ML") # random is ~ age
summary(fm1)
```

```{r}
AIC(fm1)
```

### More Complex LME Model

```{r}
fm2 <- lme(distance ~ age + Sex, data = Orthodont, random = ~ 1)
summary(fm2)
```



```{r fig.width=11, fig.height=8}
par(mfrow=c(2,2))
plot(fm2,pch=18)
par(mfrow=c(1,1))
```
