---
title: "Machines Data -  {nlme} Example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nlme)
library(janitor)
library(tidyverse)
```

### Computation on a mixed effects model

Pinheiro-Bates describes an experiment whereby the productivity of six
randomly chosen workers are assessed three times on each of three
machines, yielding the 54 observations tabulated below.

```{r}
library(nlme)
data(Machines)
head(Machines)
```
```{r}
Machines %>% group_by(Worker,Machine) %>% tally()
```

(Overall mean score = $59.65$, mean on machine A = $52.35$ , mean
on machine B = $60.32$, mean on machine C = $66.27$)

```{r}
Machines %>% group_by(Machine) %>% summarize( meanscore= mean(score))
```

### Model

* The `worker' factor is modelled with random effects($u_{i}$),
whereas the `machine' factor is modelled with fixed effects
($\beta_{j}$). 
* Due to the repeated nature of the data, interaction
effects between these factors are assumed to be extant, and shall
be examined accordingly. 

* The interaction effect in this case
($\tau_{ij}$) describes whether the effect of changing from one
machine to another is different for each worker. 
* The productivity
score $y_{ijk}$ is the $k$th observation taken on worker $i$ on
machine $j$, and is formulated
 as follows;

$$\begin{equation}
y_{ijk} = \beta_{j} + u_{i} + \tau_{ij} + \epsilon_{ijk}
\end{equation}$$
$$\begin{equation*}
u_{i} \sim N(0, \sigma^{2}_{u}), \epsilon_{ijk} \sim N(0,
\sigma^{2}), \tau_{i} \sim N(0, \sigma^{2}_{\tau})
\end{equation*}$$

### {nlme} R Package
The {nlme} package is incorporated into the R programming to
perform linear mixed model calculations. For the `Machines' data,
Pinheiro-Bates use the following code, with the hierarchical structure
specified in the last argument.

```{r}
lme(score~Machine, data=Machines, random=~1|Worker/Machine)
```


The output of the R computation is given below.

### Summary 
The crucial pieces of information given in the programme output
are the estimates of the intercepts for each of the three
machines. 

Machine A, which is treated as a control case, is
estimated to have an intercept of 52.35. 

The intercept estimates
for machines B and C are found to be $60.32$ and $66.27$ (by
adding the values 7.96 and 13.91 to 52.35 respectively).

Estimatefor the variance components are also given; 
* $\sigma^{2}_{u} = (4.78)^{2}$ 
* $\sigma^{2}_{\tau} = (3.73)^{2}$
* $\sigma^{2}_{\epsilon} = (0.96)^{2}$
