
###  RImplementation (part 1)
We have 3 doors to choose from, so we will dene a numeric sequence between 1 and 3. The
command <tt>sample(,n)<tt>takes a sample of size n from a specified set of values. Here we just
want to select one door to be our ***correct door and another to be ***selected door (i.e. the door that the contestant selects).

These events are independent. For the sake of simplicity, we will perform the selection for both doors separately, but this can be implemented in one command.

```{r}
Doors = 1:3 
Correct = sample(Doors,1)
Choice = sample(Doors,1)
````



```{r}
> Doors
[1] 1 2 3
> Correct
[1] 1
> Choice
[1] 3
````

------------------------------------------

###  R Implementation (part 2)
In this part, we will use the set theory commands <tt>union() and <tt>setdiff(). Another related command is <tt>intersect(), but we will not be using it here.

These commands are very simple, and can be best demonstrated with a simple example.

```{r}
X <- 1:6
Y <- 5:9
union(X,Y)
intersect(X,Y)
setdiff(X,Y)
setdiff(Y,X)
````

The output from these commands looks like this:
```{r}
> X
[1] 1 2 3 4 5 6
> Y
[1] 5 6 7 8 9
>
> union(X,Y)
[1] 1 2 3 4 5 6 7 8 9
> intersect(X,Y)
[1] 5 6
> setdiff(X,Y)
[1] 1 2 3 4
> setdiff(Y,X)
[1] 7 8 9
````
Going back to the Monty Hall problem. The Set Theory commands aver very useful in that they can avoid a lot of programming control statements to account for various scenarios. A wrong door must be selected to be opened. The door selected by the contestant can't be
chosen. First let us select the doors that must stay closed, then find the doors that the contestant can choose from.

```{r}
StayClosed = union(Correct, Choice)
CanOpen = setdiff(Doors, StayClosed)
````

-----------------------------------------------------

### Using the <tt>sample command


The Rcommand <tt>sample() poses an interesting problem. Lets look at the help file to see what it is.
Create a single element vector (let's call it \textbf{Vec, with that single element being 4. 

Now sample a single value from that data set. You may expect to get 4 each time, as it is the only element in the set but Rdoesnt work that way in this instance.

```{r}
Vec=c(4)
sample(Vec,1)
````

Instead Rinterprets the argument 4 as the upper bound of a sequence of four integers (i.e. the sample space is 1 to 4)
```{r}
> sample(Vec,1)
[1] 3
> sample(Vec,1)
[1] 4
> sample(Vec,1)
[1] 1
````

Problems will arise when the number of elements in the \textbf{CanOpen set of numbers is one.
When there are more than one, things will work fine.
-----------------------------------------------------------------------------------
To overcome thisproblem, we need some control statements.

*  If there are two elements in the data set, we just select one of them at randomm using the <tt>sample() command.
* If there is only one element in the data set we avoid using the <tt>sample() command, and instead use the assignment operator.

We wish to select a door for opening, to reveal the goat.

```{r}
if(length(CanOpen)>1)
{
Open = sample(CanOpen,1)
else {Open=CanOpen
Open # This door reveals a goat
````

-----------------------------------------------------------------------------------
### Stick or Switch?
There are two unopened doors. One is the correct door, but which one? Should the contestant \textbf{stick or \textbf{switch ?


```{r}
NotOpen = setdiff(Doors,Open)
Stick = Choice
# Stick with the original choice
Switch = setdiff(NotOpen,Choice)
# Switch to the other door
````


Was sticking the right decision? Or was switching?


```{r}
# Was sticking the right decision? Or was switching?
# The following logicial statements will tell us.
Stick==Correct
Switch==Correct
````


An answer of 1 means that decision would have been correct. Equivalently, a response of 0 indicates a bad decision.
-----------------------------------------------------------------


###  Writing a Function
We are going to create a function <tt>MHfunc() to help us simulate a solution for the Monty Hall Problem. The function is constructed using R code we have seen already. The output of the
function is returned as a data frame, with 7 columns:


* Correct : The number of the correct door.
* Choice : The door that the contestant chose originally, and the door selected if the contestant decided to "stick".
* Open : The door Monty Hall opens to reveal a goat
* StickDoor : The door that the contestant chose originally, and the door selected if the contestant decided to "stick".
* SwitchDoor : The door selected if the contestant chose to "switch".
* Stick: Was Sticking a good decision (1 for yes, 0 for no)
* Switch : Was Switching a good decision (1 for yes, 0 for no)

The output looks like this:
```{r}
> MHfunc()
   Correct     Choice Open(goat)  StickDoor SwitchDoor      Stick     Switch 
         2          2          1          2          3          1          0 
> MHfunc()
   Correct     Choice Open(goat)  StickDoor SwitchDoor      Stick     Switch 
         1          2          3          2          1          0          1 
> MHfunc()
   Correct     Choice Open(goat)  StickDoor SwitchDoor      Stick     Switch 
         1          2          3          2          1          0          1
````



```{r}
MHfunc <- function(numdoors=3){
Doors = 1:numdoors
Correct = sample(Doors,1)
Choice = sample(Doors,1)
StayClosed = union(Correct, Choice)
CanOpen = setdiff(Doors, StayClosed)
#Open the Door with a Goat
if(length(CanOpen)>1)
{
Open = sample(CanOpen,1) #Problem here
else {Open=CanOpen
NotOpen = setdiff(Doors,Open)
#Stick or Switch
Stick = Choice
Switch = setdiff(NotOpen,Choice)
Stick.Decision= as.numeric(Stick==Correct)
Switch.Decision = as.numeric(Switch==Correct)
#Preparing the Output
MHoutput=c(Correct,Choice,Open,Stick,Switch,
 Stick.Decision,Switch.Decision)
names(MHoutput)= c("Correct","Choice","Open(goat)","StickDoor",
 "SwitchDoor","Stick","Switch")
return(MHoutput)

````


-------------------------------------------------------------

###  A Simulation Study of the Monty Hall Problem
Let us perform a simulation study of the Monty Hall problem. We are putting the code we have
written already, placed in a for loop, although we alter the ending to make for more readable
output.

```{r}
#Initialise Counters
StickCount = 0   
SwitchCount = 0
#How many Trials?
M=1000
# a For loop
for(i in 1:M)
	{
	output=MHfunc()
	StickCount = StickCount + output[6]
	SwitchCount = SwitchCount + output[7]
	
StickCount
SwitchCount
```


The output of this program would look like this. The proportion is approximately 2 to 1 in
favour of the switching option.
```{r}
> StickCount
[1] 323
> SwitchCount
[1] 677
````

