---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---



Outliers
===========================================

* *"Outliers are sample values that cause surprise in relation to the majority of the sample"* 

* (W.N. Venables and B.D. Ripley. 2002. Modern applied statistics with S. New York: Springer, p.119).


* Crucially, surprise is in the mind of the beholder and is dependent on some explicit model of the data. 

* Importantly, Normality is only an assumption:There may be another model under which the outlier is not surprising at all, say if the data really are lognormal or 
gamma rather than normal. 


---


```{r}

library(faraway)
data("cheddar")

```

---

Outliers
===========================================



* Data points that diverge in a big way from the overall pattern are referred to as ``outliers".

* In the case of Simple Linear Regression, there are four ways that a data point might be considered an outlier.


  *  It could have an extreme X value compared to other data points.
  *  It could have an extreme Y value compared to other data points.
  *  It could have extreme X and Y values.
  *  It might be distant from the rest of the data, even without extreme X or Y values.

---

Outliers
===========================================



*  After a regression line has been computed for a group of data, a point which lies far from the line 
(and thus has a large residual value) is known as an outlier. 
* Such points may represent erroneous data, or may indicate a poorly fitting regression line. 

*  If a point lies far from the other data in the horizontal direction, it is known as an ***influential observation***. 
* The reason for this distinction is that these points have may have a significant impact on the slope of the regression line.

---

Outliers
===========================================




#### ``outlierTest()``
* Suppose we have some fitted models and we would like to see if there are any outliers. 

* For this purpose, we can use ``outlierTest()`` from {car} R package. 

---

Outliers
===========================================





```{r}
library(car)
outlierTest(Fit_1)   
```

```{r}

outlierTest(Fit_2)   
```

---

Outliers
===========================================



```{r}

outlierTest(Fit_3)   
```

```{r}

outlierTest(Fit_4)   
```