---
title: Diagnostics Plots For Regression Analysis 
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



### Diagnostics Plots for Linear Models

For the purposes of demonstration, we will create a conventional linear model using the <tt>lm()</tt> function and the ***mtcars*** dataset.

```{r}
lmodel <-  lm(mpg ~ cyl + wt,data=mtcars)
summary(lmodel)

# What are the regression coefficients?
coef(lmodel)

# What are the p-values for the regression coefficients?
summary(lmodel)$coefficients



# What are the 95% confidence intervals for this coefficients?
confint(lmodel)

# What is the AIC value for this fitted model?
AIC(lmodel)


```

### Test Residuals for Normality

- compute the residuals of the fitted model using the `resid()` command.
- Test the hypothesis of normality using `shapiro.test()`.
```{r}
Residuals = resid(lmodel)

shapiro.test(Residuals)
```
The p-value is 0.06341. We fail to reject the null hypothesis that the residuals are normally distributed.

#### Use `mfrow` argument to create grid of plots

* `mfrow` is used to create a grid of plots. Simply specify the number of rows and columns of the grid. 

* On each of the plots, which points are identified as being influential?

* Which points, if any, are influential according to all four plots.

```{r fig.width=18, fig.height=20}
par(mfrow=c(2,2))
plot(lmodel,pch=18,col="blue",cex=2)
#Put it back to Normal Mode
par(mfrow=c(1,1))
```

#### Diagnostic Plots for Linear Models with R
Plot Diagnostics for an <tt>{lm}</tt> Object

There are actually six diagnostic plots created using the plot command. The six plots, which are selectable by <tt>which</tt>, are as follows:

1. a plot of residuals against fitted values, 
2. a Scale-Location plot of **{sqrt(| residuals |}** against fitted values, 
3. a Normal Q-Q plot, 
4. a plot of Cook's distances versus row labels, 
5. a plot of residuals against leverages, 
6. a plot of Cook's distances against leverage/(1-leverage).


Plot 4 and 6 are less commonly used however. By default, the first three and 5 are provided.

```{r fig.width=18, fig.height=20}
par(mfrow=c(3,2))
plot(lmodel,pch=18,which=1:6,col="blue",cex=2)
#Put it back to Normal Mode
par(mfrow=c(1,1))
```