---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Outliers and Influential Observations

\begin{quote}
"Outliers are sample values that cause surprise in relation to the majority of the sample" (W.N. Venables and B.D. Ripley. 2002. Modern applied statistics with S. New York: Springer, p.119).
\end{quote}

Crucially, surprise is in the mind of the beholder and is dependent on some explicit model of the data. 

Importantly, Normality is only an assumption:There may be another model under which the outlier is not surprising at all, say if the data really are lognormal or 
gamma rather than normal. 

### Outliers

Data points that diverge in a big way from the overall pattern are referred to as ``outliers".\\ 
In the case of Simple Linear Regression, there are four ways that a data point might be considered an outlier.



* It could have an extreme X value compared to other data points.
* It could have an extreme Y value compared to other data points.
* It could have extreme X and Y values.
* It might be distant from the rest of the data, even without extreme X or Y values.


* After a regression line has been computed for a group of data, a point which lies far from the line 
(and thus has a large residual value) is known as an outlier. 
Such points may represent erroneous data, or may indicate a poorly fitting regression line. 

* If a point lies far from the other data in the horizontal direction, it is known as an \textit{\textbf{influential observation}}. 
The reason for this distinction is that these points have may have a significant impact on the slope of the regression line.

-------------------------------
### Identifying Outliers

Possible outliers can be identified by looking at the diagnostic plot of the residuals

Another approach is to use the <tt>outlierTest()</tt> function of the ***car*** package:

### outlierTest()
Suppose we have a two fitted models and we would like to see if there are any outliers. 

For this purpose, we can use outlierTest() from \texttt{library(car)} in R. 


```{r}
#If package not installed, uncomment next line.
#install.packages("car")

library(car)
outlierTest(Fit1)  
```


```{r}
outlierTest(Fit2)   
```

The row numbers ( here : 21, 15 and 177) indicate the outlier points in the data.


