---
title: "Weibull distribution"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Standard parameterization

The probability density function of a Weibull random variable is:

\[{\displaystyle f(x;\lambda ,k)={\begin{cases}{\frac {k}{\lambda }}\left({\frac {x}{\lambda }}\right)^{k-1}e^{-(x/\lambda )^{k}}&x\geq 0,\\0&x<0,\end{cases}}}\]


where $k > 0$ is the shape parameter and $\lambda > 0$ is the scale parameter of the distribution. Its complementary cumulative distribution function is a stretched exponential function. 


The Weibull distribution is related to a number of other probability distributions; in particular, it interpolates between the exponential distribution ($k = 1$) and the Rayleigh distribution ($k = 2$) and \[{\displaystyle \lambda ={\sqrt {2}}\sigma }\].

#### Density Plot of Random Covariates

 * Shape Parameter set to 2,3,5 and 10.

```{r rweibuill}
Weibull = data.frame(
  k2 <- rweibull(1000,shape=2),
  k3 <- rweibull(1000,shape=3),
  k5 <- rweibull(1000,shape=5),
  k10 <- rweibull(1000,shape=10)
)
  
```

```{r echo=FALSE}
library(ggplot2)
ggplot() + geom_density( data = Weibull, aes(x=k2),col="black") + 
  geom_density( data = Weibull, aes(x=k3),col="red") +
  geom_density( data = Weibull, aes(x=k5),col="green") + 
  geom_density( data = Weibull, aes(x=k10),col="blue") + 
  geom_hline(yintercept=0) + 
  geom_vline(xintercept=0) + 
  xlab("Weibull Distribution") + 
  theme_bw()
  
```