---
title: "OSUN R Users Community"
subtitle: "Kevin O'Brien"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

### EasyStats

* Performance R package

```{r message =FALSE, warning=FALSE}

library(olsrr)

```

---



The olsrr package provides following tools for teaching and learning OLS regression using R:

comprehensive regression output
residual diagnostics
measures of influence
heteroskedasticity tests
collinearity diagnostics
model fit assessment
variable contribution assessment
variable selection procedures

---

### Residual Diagnostics

olsrr offers tools for detecting violation of standard regression assumptions. Here we take a look at residual diagnostics. The standard regression assumptions include the following about residuals/errors:

The error has a normal distribution (normality assumption).
The errors have mean zero.
The errors have same but unknown variance (homoscedasticity assumption).
The error are independent of each other (independent errors assumption).

---

Residual QQ Plot
Graph for detecting violation of normality assumption.


```{r}
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_qq(model)
```

---
Residual Normality Test
Test for detecting violation of normality assumption.

```{r}
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_test_normality(model)
```
---


Correlation between observed residuals and expected residuals under normality.

```{r}
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_test_correlation(model)
```

---

### Residual vs Fitted Values Plot
It is a scatter plot of residuals on the y axis and fitted values on the x axis to detect non-linearity, unequal error variances, and outliers.

---

### Characteristics of a well behaved residual vs fitted plot:

The residuals spread randomly around the 0 line indicating that the relationship is linear.
The residuals form an approximate horizontal band around the 0 line indicating homogeneity of error variance.
No one residual is visibly away from the random pattern of the residuals indicating that there are no outliers.

```{r}
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_fit(model)
```
---

### Residual Histogram
Histogram of residuals for detecting violation of normality assumption.

```{r}
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_hist(model)
```

---
