---
title: "Test for Autocorrelated Errors"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Model Assumptions for Linear Models

Recall the 4 assumptions made about residuals in linear models 

1. The mean value of the residuals is zero,
2. The variance of residuals are constant across the range of measurements,
3. The residuals are normally distributed,
4. Residuals are independent.

We will address the 4th assumption here in this demonstration. 
 
### Non-independence of Errors: Autocorrelation
The Durbin–Watson statistic is a test statistic used to detect the presence of autocorrelation (a relationship between values separated from each other by a given time lag) in the residuals (prediction errors) from a regression analysis. It is named after James Durbin and Geoffrey Watson.

### {car} - The Regression Diagnostics Package

An excellent review of regression diagnostics is provided in John Fox's aptly named ***Overview of Regression Diagnostics***. Dr. Fox's car package provides advanced utilities for regression modeling.

```{r warning=FALSE,message=FALSE}
library(car)
```

#### Linear Model using "mtcars"

```{r}
# Assume that we are fitting a multiple linear regression
# on the MTCARS data
library(car)
fit <- lm(mpg~disp+hp+wt+drat, data=mtcars)
```

### Non-independence of Errors: Autocorrelation
The Durbin–Watson statistic is implemented using the command <tt>durbinWatsonTest()</tt>

```{r}
# Test for Autocorrelated Errors
durbinWatsonTest(fit)
```