---
title: "Measures of Influence"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

## Measures of Influence
The impact of an observation on a regression fitting can be determined by the difference between the estimated regression coefficient of a model with all observations and the estimated coefficient when the particular observation is deleted. 

#### Linear Model using "mtcars"

```{r warning=FALSE,message=FALSE}
# Assume that we are fitting a multiple linear regression
# on the MTCARS data
library(car)
fit <- lm(mpg~disp+hp+wt+drat, data=mtcars)
```

### DFBETA

The DFBETA statistic for measuring the influence of 
the th observation is defined as the one-step approximation to 
the difference in the MLE of the regression parameter vector and 
the MLE of the regression parameter vector without the $i$-th 
observation. This one-step approximation assumes a Fisher 
scoring step, and is given by

$$ DFBETA_{a} = \hat{\beta} - \hat{\beta}_{(i)} \\
= B(Y-Y_{\bar{i}}) $$


```{r include=FALSE,echo=FALSE}
#  $e_{-Q} = y_{Q} - x_{Q}\hat{\beta}^{-Q}$
#  $PRESS_{(U)} = y_{i} - x\hat{\beta}_{(U)}$

```



#### Computing DFBETAs

```{r}
### First Five Cases
dfbeta(fit) %>% head(5)
```

```{r}
## using {car}
dfbetaPlots(fit, pch=18,col = "red")
```

### DFFITS
DFFITS is a statistical measured designed to a show how influential an observation is in a statistical model. It is closely related to the studentized residual.

$$ DFFITS = {\widehat{y_i} -
\widehat{y_{i(k)}} \over s_{(k)} \sqrt{h_{ii}}}$$


#### Computing DFFITS

```{r}
dffits(fit)
```


### PRESS
The prediction residual sum of squares (PRESS) is an value associated with this calculation. When fitting linear models, PRESS can be used as a criterion for model selection, with smaller values indicating better model fits.
$$ PRESS = \sum(y-y^{(k)})^2 $$

#### Computing PRESS

For linear models, <tt>rstandard(, type = "predictive")</tt> provides leave-one-out cross validation residuals, and the “PRESS” statistic (PREdictive Sum of Squares, the same as the CV score) of model "model "fit" is

```{r}
   PRESS <- sum(rstandard(fit, type="pred")^2)
```