

Diagnostic Plots for Linear Models with R
===================================================

### Plot Diagnostics for an `lm()` Object

There are six plots (selectable by `which=`) are currently available: 

*  a plot of residuals against fitted values, 
*  a Normal Q-Q plot, 
*  a Scale-Location plot of *sqrt( $\|$ residuals $\|$ })* against fitted values, 

*  a plot of Cook's distances versus row labels, 
*  a plot of residuals against leverages, 
*  a plot of Cook's distances against *leverage/(1-leverage)*.

* By default, the first three and 5 are provided, if you just type something like `plot(fit)`.


Diagnostic Plots
======================================================
class:smaller-95

### Diagnostic Plot 1

* The first one displays the residuals vs. the fitted values we use this to evlauate the mean, variance and correlation of residuals.

* If our assumptions of constant variance and uncorrelated residuals are violated we **may** be able to correct this with a variance-stabilizing transformation.

* see ``ncevTest()``


Diagnostic Plots
======================================================
class:smaller-95

### Diagnostic Plot 1


  plot(Fit_4,
     which=1,
     pch=16,lwd=1.2)
   

Just increment the "``which=``" argument with any integer between 1 and 6


Diagnostic Plots
======================================================
class:smaller-95

Diagnostic Plot 1

```{r out.width = "80%",echo=FALSE }
plot(Fit_4,which=1,
     pch=16,lwd=1.2)
```




Diagnostic Plots
======================================================

#### Diagnostic Plot 2

* The second plot helps us check the normality of the residuals. If the
residuals are indeed normal, they should fall along the dashed line.
* Remember that the normality assumption for our errors allows us to determine
the standard errors of our coefficients and predictions.

Diagnostic Plot 2
======================================================
class:smaller-95


```{r echo = FALSE,out.width = "80%" }
plot(Fit_4,which=2,
     pch=16,lwd=1.2)

```



 

Diagnostic Plots
======================================================

#### Diagnostic Plot 3

* The ***Scale-Location*** plot, also called 'Spread-Location' (or 'S-L' plot), takes the square root of the absolute residuals in order to diminish skewness (sqrt($|E|$)) is much less skewed than $| E |$ for Gaussian zero-mean E).
 
 
Diagnostic Plots 3
======================================================
class:smaller-95


```{r echo = FALSE,out.width = "80%" }
plot(Fit_4,which=3,
     pch=16,lwd=1.2)

```

Diagnostic Plots
======================================================

#### Diagnostic Plot 4

* This plot details the Cook's Distance for each observation. 
* We will revert to this later.
 
 
Diagnostic Plots 4
======================================================
class:smaller-95


```{r echo = FALSE,out.width = "80%" }
plot(Fit_4,which=4,
     pch=16,lwd=1.2)

```

Extracting Cook's Distance
======================================================

```{r}
cooks.distance(Fit_4) %>% 
  head(5) %>% 
  round(4) %>%
  t()
```

Diagnostic Plots
======================================================

#### Diagnostic Plot 5

* The ***Residual-Leverage*** plot shows contours of equal Cook's distance, for values of `cook.levels` (by default 0.5 and 1) and omits cases with leverage one with a warning. 
* If the leverages are constant the plot uses factor level combinations instead of the leverages for the x-axis. 
* **(The factor levels are ordered by mean fitted value.)**

Diagnostic Plot 5
======================================================
class:smaller-95


```{r echo = FALSE,out.width = "80%" }
plot(Fit_4,which=5,
     pch=16,lwd=1.2)

```


Diagnostic Plots
=====================================================

#### Diagnostic Plots 6

* The final plot will display our residuals vs. their leverage. 
* The dashed red lines are level curves that denote a particular value of Cook's distance.
* We will pay attention to points lying beyond the distance of 1. 
* Notice that when we have data with row labels, the points will be labeled with their names. Otherwise, the row number will be shown.

Diagnostic Plot 6
======================================================
class:smaller-95



```{r echo = FALSE,out.width = "80%" }
plot(Fit_4,which=6,
     pch=16,lwd=1.2)

```

