Linear modelling with R (Cheeses)
======================================================

#### Cheddar Cheese Taste

* As cheese ages, various chemical processes take place that determine the taste of the final product. 
* This dataset contains concentrations of various chemicals in 30 samples of mature cheddar cheese, and a subjective measure of taste for each sample. 
*The variables "Acetic" and "H2S" are the natural logarithm of the concentration of acetic asid and hydrogen sulfide respectively. 
* The variable "Lactic" has not been transformed.

#### Reference: 

* Moore, David S., and George P. McCabe (1989). Introduction to the Practice of Statistics.



Linear modelling with R (Cheeses)
======================================================

* Number of cases: 30

#### Variable Names:

* **Case**: Sample number
* **Taste**: Subjective taste test score, obtained by combining the scores of several tasters
* **Acetic**: Natural log of concentration of acetic acid
* **H2S**: Natural log of concentration of hydrogen sulfide
* **Lactic**: Concentration of lactic acid




Linear modelling with R (Cheeses)
======================================================


```{r}
CheesesData <- read_csv("C:/Users/Kevin/Documents/Ghana_R_Users/Cheeses.csv")

head(CheesesData)
```

Linear modelling with R (Cheeses)
======================================================



```{r}
Fit_1 <- lm(Taste ~ Acetic + Lactic, data = CheesesData)
Fit_2 <- lm(Taste ~ Acetic + H2S, data = CheesesData)
Fit_3 <- lm(Taste ~ H2S + Lactic, data = CheesesData)

Fit_4 <- lm(Taste ~ Acetic + H2S + Lactic, data = CheesesData)
```

Linear modelling with R (Cheeses)
======================================================
class:smaller-95

#### Aikaike Information Criterion

```{r}
AIC(Fit_1)

AIC(Fit_2)

AIC(Fit_3)

AIC(Fit_4)

```



{modelr}
=======================
Compute model quality for a given dataset

Three summaries are immediately interpretible on the scale of the response variable:

* ``rmse()`` is the root-mean-squared-error
* ``mae()`` is the mean absolute error
* ``qae()`` is quantiles of absolute error.

{modelr}
=======================

### Root Mean Square Error

```{r}
library(modelr)
rmse(Fit_4,CheesesData)
```


### mean absolute error

```{r}
mae(Fit_4,CheesesData)
```


{modelr}
=======================

```{r}
qae(Fit_4,CheesesData)
```

{modelr}
=======================
Other summaries

* ``mape()`` mean absolute percentage error.
* ``rsae()`` is the relative sum of absolute errors.
* ``mse()`` is the mean-squared-error.
* ``rsquare()`` is the variance of the predictions divided by the variance of the response.

{modelr}
=======================

```{r}
rsquare(Fit_4,CheesesData)
```