---
title: "Untitled"
output:
  html_document:
    toc: true
    fig_caption: true
---

```{r echo=FALSE, include=FALSE}
require(tidyverse)

```
Truncated regression is used to model dependent variables for which some of the observations are not included in the analysis because of the value of the dependent variable.

```{r message=FALSE,warning=FALSE}
#install.packages("truncreg")
require(foreign)
require(ggplot2)
require(truncreg)
require(boot)

```


### Examples of truncated regression

* ***Example 1.*** A study of students in a special GATE (gifted and talented education) program wishes to model achievement as a function of language skills and the type of program in which the student is currently enrolled. A major concern is that students are required to have a minimum achievement score of 40 to enter the special program. Thus, the sample is truncated at an achievement score of 40.

* ***Example 2.*** A researcher has data for a sample of Americans whose income is above the poverty line. Hence, the lower part of the distribution of income is truncated. If the researcher had a sample of Americans whose income was at or below the poverty line, then the upper part of the income distribution would be truncated. In other words, truncation is a result of sampling only part of the distribution of the outcome variable.
