---
title: "Untitled"
output:
  html_document:
    toc: true
    fig_caption: true
---

```{r echo=FALSE, include=FALSE}
require(tidyverse)

```


require(foreign)
require(ggplot2)
require(GGally)
require(survival)
require(rgl)

####################################################

dat <- read.dta("https://stats.idre.ucla.edu/stat/data/intreg_data.dta")

# summary of the variables
summary(dat)

```{r}
# bivariate plots
ggpairs(dat[, -1], lower = list(combo = "box"), upper = list(combo = "blank"))
```

```{r}
by(dat[, 2:5], dat$type, colMeans, na.rm = TRUE)
```

# setup the survival object with interval censoring
(Y <- with(dat, Surv(lgpa, ugpa, event = rep(3, nrow(dat)), type = "interval")))
##  [1] [2.5, 3.0] [3.4, 3.8] [2.5, 3.0] [0.0, 2.0] [3.0, 3.4] [3.4, 3.8]
##  [7] [3.8, 4.0] [2.0, 2.5] [3.0, 3.4] [3.4, 3.8] [2.0, 2.5] [2.0, 2.5]
## [13] [2.0, 2.5] [2.5, 3.0] [2.5, 3.0] [2.5, 3.0] [3.4, 3.8] [2.5, 3.0]
## [19] [2.0, 2.5] [3.0, 3.4] [3.4, 3.8] [3.8, 4.0] [2.0, 2.5] [3.0, 3.4]
## [25] [3.4, 3.8] [2.0, 2.5] [2.0, 2.5] [2.0, 2.5] [2.5, 3.0] [2.5, 3.0]
m <- survreg(Y ~ write + rating + type, data = dat, dist = "gaussian")

summary(m)


# for the regression surface
f <- function(x, y, type = "vocational") {
    newdat <- data.frame(write = x, rating = y, type = factor(type, levels = levels(dat$type)))
    predict(m, newdata = newdat)
}

```{r}
# Create X, Y, and Z grids
X <- with(dat, seq(from = min(write), to = max(write), length.out = 10))
Y <- with(dat, seq(from = min(rating), to = max(rating), length.out = 10))
Z <- outer(X, Y, f)
```

# Create 3d scatter plot and add the regression surface
open3d(windowRect = c(100, 100, 700, 700))
## [1] 1
with(dat, plot3d(x = write, y = rating, z = ugpa, xlab = "write", ylab = "rating", 
    zlab = "ugpa", xlim = range(write), ylim = range(rating), zlim = range(ugpa)))
par3d(ignoreExtent = TRUE)

# add regression surface for each type of program in a different colour
# with 50 percent transparency (alpha = .5)
surface3d(X, Y, outer(X, Y, f, type = "vocational"), col = "blue", alpha = 0.5)
surface3d(X, Y, outer(X, Y, f, type = "general"), col = "red", alpha = 0.5)
surface3d(X, Y, outer(X, Y, f, type = "academic"), col = "green", alpha = 0.5)

# create an animated movie movie3d(spin3d(axis=c(.5,.5,.5), rpm=5),
# duration=6, dir = 'intreg_fig')
