---
title: "Titanic Survivors"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## What is Multinomial Regression ?

Multinomial Logistic Regression (MLR) is a form of linear regression analysis conducted when the dependent variable is nominal with more than two levels. 
It is used to describe data and to explain the relationship between one dependent nominal variable and one or more continuous-level (interval or ratio scale) independent variables. You can understand nominal variable as, a variable which has no intrinsic ordering.

For example: Types of Forests: ‘Evergreen Forest’, ‘Deciduous Forest’, ‘Rain Forest’. 

As you see, there is no intrinsic order in them, but each forest represent a unique category. 
In other words, multinomial regression is an extension of logistic regression, which analyzes dichotomous (binary) dependents.

```{r}
install.packages(c("nnet","sandwich","msm","reshape2"))

library(foreign)
library(ggplot2)
library(MASS)
library(Hmisc)
library(reshape2)
library(sandwich)
library(msm)
```

Installing packages into '/home/nbcommon/R'
(as 'lib' is unspecified)
Loading required package: lattice
Loading required package: survival
Loading required package: Formula

Attaching package: 'Hmisc'

The following objects are masked from 'package:base':

    format.pval, round.POSIXt, trunc.POSIXt, units

```{r}
library(nnet)
library(ggplot2)
library(reshape2)

ml <- read.csv("https://raw.githubusercontent.com/RWorkshop/workshopdatasets/master/multilog.csv")
with(ml, table(ses, prog))
```
        prog
ses      academic general vocation
  high         42       9        7
  low          19      16       12
  middle       44      20       31
with(ml, do.call(rbind, tapply(write, prog, function(x) c(M = mean(x), SD = sd(x)))))
M	SD
academic	56.25714	7.943343
general	51.33333	9.397775
vocation	46.76000	9.318754
ml$prog2 <- relevel(ml$prog, ref = "academic")
test <- multinom(prog2 ~ ses + write, data = ml)
# weights:  15 (8 variable)
initial  value 219.722458 
iter  10 value 179.983731
final  value 179.981726 
converged

```{r}
summary(test)
```

Call:
multinom(formula = prog2 ~ ses + write, data = ml)

Coefficients:
         (Intercept)    seslow sesmiddle       write
general     1.689478 1.1628411 0.6295638 -0.05793086
vocation    4.235574 0.9827182 1.2740985 -0.11360389

Std. Errors:
         (Intercept)    seslow sesmiddle      write
general     1.226939 0.5142211 0.4650289 0.02141101
vocation    1.204690 0.5955688 0.5111119 0.02222000

Residual Deviance: 359.9635 
AIC: 375.9635 
### Test Statistics and P-values
z <- summary(test)$coefficients/summary(test)$standard.errors
z

p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
(Intercept)	seslow	sesmiddle	write
general	1.376987	2.261364	1.353816	-2.705658
vocation	3.515904	1.650050	2.492798	-5.112687
(Intercept)	seslow	sesmiddle	write
general	0.1685163893	0.02373673	0.1757949	6.816914e-03
vocation	0.0004382601	0.09893276	0.0126741	3.176088e-07


---------------------------------------------------------------------------
<hline>

### Interpretation
Model execution output shows some iteration history and includes the final negative log-likelihood 179.981726. This value is multiplied by two as shown in the model summary as the Residual Deviance.

The summary output has a block of coefficients and another block of standard errors. Each blocks has one row of values corresponding to one model equation. In the block of coefficients, we see that the first row is being compared to prog = “general” to our baseline prog = “academic” and the second row to prog = “vocation” to our baseline prog = “academic”.

A one-unit increase in write decreases the log odds of being in general program vs. academic program by 0.0579

A one-unit increase in write decreases the log odds of being in vocation program vs. academic program by 0.1136

The log odds of being in general program than in academic program will decrease by 1.163 if moving from ses=”low” to ses=”high”.

On the other hand, Log odds of being in general program than in academic program will decrease by 0.5332 if moving from ses="low" to ses=”middle”

The log odds of being in vocation program vs. in academic program will decrease by 0.983 if moving from ses="low" to ses=”high”

The log odds of being in vocation program vs. in academic program will increase by 0.291 if moving from ses="low" to ses="middle"

Now we’ll calculate Z score and p-Value for the variables in the model.

## extract the coefficients from the model and exponentiate
exp(coef(test))

head(pp <- fitted(test))
(Intercept)	seslow	sesmiddle	write
general	5.416653	3.199009	1.876792	0.9437152
vocation	69.101326	2.671709	3.575477	0.8926115
academic	general	vocation
1	0.1482721	0.3382509	0.5134769
2	0.1201988	0.1806335	0.6991678
3	0.4186768	0.2368137	0.3445095
4	0.1726839	0.3508433	0.4764728
5	0.1001206	0.1689428	0.7309367
6	0.3533583	0.2378047	0.4088370
dses <- data.frame(ses = c("low", "middle", "high"), write = mean(ml$write))

predict(test, newdata = dses, "probs")
academic	general	vocation
1	0.4396813	0.3581915	0.2021272
2	0.4777451	0.2283359	0.2939190
3	0.7009046	0.1784928	0.1206026
dwrite <- data.frame(ses = rep(c("low", "middle", "high"), each = 41), write = rep(c(30:70),
    3))

## store the predicted probabilities for each value of ses and write
pp.write <- cbind(dwrite, predict(test, newdata = dwrite, type = "probs", se = TRUE))

## calculate the mean probabilities within each level of ses
by(pp.write[, 3:5], pp.write$ses, colMeans)
pp.write$ses: high
 academic   general  vocation 
0.6164348 0.1808049 0.2027603 
------------------------------------------------------------ 
pp.write$ses: low
 academic   general  vocation 
0.3972955 0.3278180 0.2748864 
------------------------------------------------------------ 
pp.write$ses: middle
 academic   general  vocation 
0.4256172 0.2010877 0.3732951 
## melt data set to long for ggplot2
lpp <- melt(pp.write, id.vars = c("ses", "write"), value.name = "probability")
head(lpp)  # view first few rows

## plot predicted probabilities across write values for each level of ses
## facetted by program type
ggplot(lpp, aes(x = write, y = probability, colour = ses)) + geom_line() + facet_grid(variable ~
    ., scales = "free")
ses	write	variable	probability
low	30	academic	0.09843258
low	31	academic	0.10716517
low	32	academic	0.11650018
low	33	academic	0.12645441
low	34	academic	0.13704163
low	35	academic	0.14827211
png

## melt data set to long for ggplot2
lpp <- melt(pp.write, id.vars = c("ses", "write"), value.name = "probability")
head(lpp)  # view first few rows
 
##   ses write variable probability
## 1 low    30 academic     0.09844
## 2 low    31 academic     0.10717
## 3 low    32 academic     0.11650
## 4 low    33 academic     0.12646
## 5 low    34 academic     0.13705
## 6 low    35 academic     0.14828
 



## plot predicted probabilities across write values for each level of ses
## facetted by program type
ggplot(lpp, aes(x = write, y = probability, colour = ses)) + geom_line() + facet_grid(variable ~
    ., scales = "free")
