---
title: "OSUN R Users Community"
subtitle: "Kevin O'Brien"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

### Exploratory Data Analysis

BE THOROUGH

* Domain Knowledge
* Summary Statistics
* Data Visualization
* Outlier Detection
* Missing Data Analysis


#### Remark: 

Cluster Analysis (e.g. K-means) can be a very useful EDA procedure.

---

### Exploratory Data Analysis

#### Data Inspection

* {inspectdf}
* {janitor}

#### Data Visualization

* {WVPlots}
* {ggally}

#### tidyverse

* {broom} and {modelr} (succeeded by {tidymodels})

* <tt>dplyr::tally()</tt>
* <tt>dplyr::distinct()</tt>


---

### Useful Packages (a selection)

* {A3}
* {arsenal}
* {analytics}
* {gdata}
* {descriptr}
* {furniture}
* {rockchalk}
* {yardstick}


---

### {inspectdf}

#### inspectdf: Inspection, Comparison and Visualisation of Data Frames

inspectdf is collection of utilities for columnwise summary, comparison and visualisation of data frames. Functions are provided to summarise missingness, categorical levels, numeric distribution, correlation, column types and memory usage. 
The package has three aims: to speed up repetitive checking and exploratory tasks for data frames

---

### {inspectdf}

Key functions
* <tt>inspect_types()</tt>: summary of column types
* <tt>inspect_mem()</tt>: summary of memory usage of columns
* <tt>inspect_na()</tt>: columnwise prevalence of missing values
* <tt>inspect_cor()</tt>: correlation coefficients of numeric columns
* <tt>inspect_imb()</tt>: feature imbalance of categorical columns
* <tt>inspect_num()</tt>: summaries of numeric columns
* <tt>inspect_cat()</tt>: summaries of categorical columns

---
  
### {inspectdf}

```{r message=FALSE, warning=FALSE}
library(inspectdf)

# Load dplyr for starwars data & pipe
library(dplyr)
# Single dataframe summary
inspect_cat(starwars)

```

---

### {inspectdf}

```{r}
# Paired dataframe comparison
inspect_cat(starwars, starwars[1:20, ])
```

---

### {inspectdf}

```{r}
# Grouped dataframe summary
starwars %>% group_by(species) %>% inspect_cat()
```

---
### WVPlots: Common Plots for Analysis
#
Select data analysis plots, under a standardized calling interface implemented on top of 'ggplot2' and 'plotly'. Plots of interest include: 'ROC', gain curve, scatter plot with marginal distributions, conditioned scatter plot with marginal densities, box and stem with matching theoretical distribution, and density with matching theoretical distribution.



```{r}
set.seed(34903490)
x = rnorm(50)
y = 0.5*x^2 + 2*x + rnorm(length(x))
frm = data.frame(
  x = x,
  y = y,
  yC = y>=as.numeric(quantile(y,probs=0.8)),
  stringsAsFactors = FALSE)

frm$absY <- abs(frm$y)
frm$posY = frm$y > 0
```

---
### Scatterplot


Scatterplot with smoothing line through points.

```{r eval=FALSE, out.width = "100%" }
WVPlots::ScatterHist(frm, "x", "y", title="Example Fit")
```

---

```{r echo=FALSE, out.width = "80%" }
WVPlots::ScatterHist(frm, "x", "y", title="Example Fit")
```

