---
title: "forcats"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(forcats)

library(magrittr)

library(dplyr)

```

## R Markdown

* ``as_factor``  
* ``fct_anon``  
* ``fct_c``  
* ``fct_collapse``  
* ``fct_count`` 
* ``fct_drop``  
* ``fct_expand`` 
* ``fct_explicit_na``  
* ``fct_inorder``
* ``fct_lump`` 
* ``fct_other`` 
* ``fct_recode`` 
* ``fct_relabel`` 
* ``fct_relevel`` 
* ``fct_reorder`` 
* ``fct_rev``  
* ``fct_shift`` 
* ``fct_shuffle`` 
* ``fct_unify`` 
* ``fct_unique`` 
* ``gss_cat``  
* ``lvls``  
* ``lvls_union`` 






```{r}

x <- factor(c( "A", "D", "B", "A", "B", "A", "A", "B", "B", "A", "C"))

fct_recode(x, O = "C", O = "D") # O for other

```


```{r}

# Dropping Unused Levels

f <- factor(c("a", "b"), levels = c("a", "b", "c"))

f

fct_drop(f)

# Set only to restrict which levels to drop

fct_drop(f, only = "a")

fct_drop(f, only = "c")

```


```{r}

gss_cat

table(gss_cat$partyid)

fct_count(gss_cat$partyid)

partyid2 <- fct_collapse(gss_cat$partyid,
   missing = c("No answer", "Don't know"),
   other = "Other party",
   rep = c("Strong republican", "Not str republican"),
   ind = c("Ind,near rep", "Independent", "Ind,near dem"),
   dem = c("Not str democrat", "Strong democrat")
)

fct_count(partyid2)

```


```{r}

x <- factor(rep(LETTERS[1:9], times = c(40, 30, 7, 5, 2, 1, 1, 1, 1)))

fct_other(x, keep = c("A", "B"))

fct_other(x, drop = c("A", "B"))

```


```{r}

f <- factor(c("a", "b", "c"))

fct_shuffle(f)

fct_shuffle(f)

```


```{r}
set.seed(1337)

Fact <- LETTERS[rbinom(1000, 6, 0.25) + 1 ]

Fact <- factor(Fact, levels=c("A","B","C","D","E","F","G") )

Var1 <- sample(100:200, 1000, TRUE)

Var2 <- rexp(1000, 1.25) %>% round(2)

myDF <- data.frame(Fact, Var1, Var2)
```







```{r}
f <- factor(c("b", "b", "a", "c", "c", "c"))
f
fct_inorder(f)
fct_infreq(f)
fct_inorder(f, ordered = TRUE)


```

# fct_count Count entries in a factor
# fct_drop Drop unused levels
# fct_anon Anonymise factor levels
# fct_collapse :  Collapse factor levels into manually defined groups
# fct_c : Concatenate factors, combining levels
# fct_other : Replace levels with "other"
# fct_inorder 

# Reorder factors levels by first appearance or frequency
