Diagnostics Plots For Regression Analysis
========================================================



```{r}
attach(mtcars)
FIT1 <-  lm(mpg ~ cyl + wt)
summary(FIT1)

# What are the regression coefficients?
coef(FIT1)

# What are the p-values for the regression coefficients?
summary(FIT1)$coefficients



# What are the 95% confidence intervals for this coefficients?
confint(FIT1)

# What is the AIC value for this fitted model?
AIC(FIT1)


```

### Test Residuals for Normality

- compute the residuals of the fitted model using the `resid()` command.
- Test the hypothesis of normality using `shapiro.test`.
```{r}
Residuals = resid(FIT1)

shapiro.test(Residuals)
```
The p-value is 0.06341. We fail to reject the null hypothesis that the residuals are normally distributed.

#### Use `mfrow` argument to create grid of plots

`mfrow` is used to create a grid of plots. Simply specify the number of rows and columns of the grid. 

On each of the plots, which points are identified as being influential?
Which points, if any, are influential according to all four plots.
```{r fig.width=18, fig.height=20}
par(mfrow=c(2,2))
plot(FIT1,pch=18,col="red",cex=2)
#Put it back to Normal Mode
par(mfrow=c(1,1))
```

