---
title: "loddsratio"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 Calculate Generalized Log Odds Ratios for Frequency Tables

#### Description

Computes (log) odds ratios and their asymptotic variance covariance matrix foC (x strata) tables.
Odds ratios are calculated for two array dimensions, separately for each level of all stratifying
dimensions. See Friendly et al. (2011) for a sketch of a general theory.

#### Usage
<pre><code>

loddsratio(x, ...)
## Default S3 method:
loddsratio(x, strata = NULL, log = TRUE,
ref = NULL, correct = any(x == 0L), ...)
## S3 method for class 'formula'
loddsratio(formula, data = NULL, ...,
subset = NULL, na.action = NULL)
oddsratio(x, stratum = NULL, log = TRUE)
## S3 method for class 'loddsratio'
loddsratio 69
coef(object, log = object$log, ...)
## S3 method for class 'loddsratio'
vcov(object, log = object$log, ...)
## S3 method for class 'loddsratio'
print(x, log = x$log, ...)
## S3 method for class 'loddsratio'
confint(object, parm, level = 0.95, log = object$log, ...)
## S3 method for class 'loddsratio'
as.array(x, log=x$log, ...)
## S3 method for class 'loddsratio'
t(x)
## S3 method for class 'loddsratio'
aperm(a, perm, ...)

</code></pre>
#### Arguments

* ``x``: an object. For the default method a k-way matrix/table/array of frequencies. The
number of margins has to be at least 2.
strata, stratum
Numeric or character indicating the margins of a $k$-way table x (with $k$
greater than 2) that should be employed as strata. By default all dimensions
except the first two are used.
ref numeric or character. Reference categories for the (non-stratum) row and column
dimensions that should be employed for computing the odds ratios. By
default, odds ratios for profile contrasts (or sequential contrasts, i.e., successive
differences of adjacent categories) are used. See details below.
formula a formula specifying the variables used to create a contingency table from data.
A conditioning formula can be specified; the conditioning variables will then be
used as strata variables.
data either a data frame, or an object of class "table" or "ftable".
subset an optional vector specifying a subset of observations to be used.
na.action a function which indicates what should happen when the data contain NAs. Ignored
if data is a contingency table.
log logical. Should the results be displayed on a log scale or not? All internal computations
are always on the log-scale but the results are transformed by default
if log = TRUE.
correct logical or numeric. Should a continuity correction be applied before computing
odds ratios? If TRUE, 0.5 is added to all cells; if numeric (or an array conforming
to the data) that value is added to all cells. By default, this not employed unless
there are any zero cells in the table, but this correction is often recommended to
reduce bias when some frequencies are small (Fleiss, 1981).
a, object an object of class loddsratio as computed by loddsratio.
perm numeric or character vector specifying a permutation of strata.
... 
</code></pre>
#### Arguments
 passed to methods.
70 loddsratio
parm a specification of which parameters are to be given confidence intervals, either
a vector of numbers or a vector of names. If missing, all parameters are considered.
level the confidence level required for the confint method.
Details
For aC table, (log) odds ratios are formed for the set of (R-1) x (C-12 tables, corresponding
to some set of contrasts among the row and column variables. The ref argument allows these
to be specified in a general way.
ref = NULL (default) corresponds to "profile contrasts" (or sequential contrasts or successive differences)
for ordered categories, i.e., R1-R2, R2-R3, R3-R4, etc., and similarly for the column
categories. These are sometimes called "local odds ratios".
regives contrasts with the first category; ref = dim(x) gives contrasts with the last category;
ref = c(2, 4) or ref = list(2, 4) corresponds to the reference being the second category in
rows and the fourth in columns.
Combinations like ref = list(NULL, 3) are also possible, as are character vectors, e.g., ref = c("foo", "bar")
also works ("foo" pertaining again to the row reference and "bar" to column reference).
Note that all such parameterizations are equivalent, in that one can derive all other possible odds
ratios from any non-redundant set, but the interpretation of these 
#### Value 
 depends on the parameterization.
Note also that these reference level parameterizations only have meaning when the primary (nonstrata)
table dimensions are larger than 2x2. In the 2x2 case, the odds ratios are defined by the order
of levels of those variables in the table, so you can achieve a desired interpretation by manipulating
the table.
See the help page of plot.loddsratio for visualization methods.
Value
An object of class loddsratio, with the following components:
coefficients A named vector, of length (R-1) x (C-1) x prod(dim(x)[strata]) containing
the log odds ratios. Use the coef method to extract these from the object, and
the confint method for confidence intervals. For a two-way table, the names
for the log odds ratios are constructed in the form Ri:Rj/Ci:Cj using the table
names for rows and columns. For a stratified table, the names are constructed in
the form Ri:Rj/Ci:Cj|Lk.
vcov Variance covariance matrix of the log odds ratios.
dimnames Dimension names for the log odds ratios, considered as a table of size (R-1, C-
1, dim(x)[strata]). Use the dim and dimnames methods to extract these and
manipulate the log odds ratios in relation to the original table.
dim Corresponding dimension vector.
contrasts A matrix C, such that C %*% as.vector(log(x)) gives the log odds ratios.
Each row corresponds to one log odds ratio, and is all zero, except for 4 elements
of c(1, -1, -1, 1) for a give2 subtable.
log A logical, indicating the value of log in the original call.
loddsratio 71
Note
The method of calculation is an example of the use of the delta method described by Agresti (2013),
Section 16.1.6, giving estimates of log odds ratios and their asymptotic covariance matrix.
The coef method returns the coefficients component as a vector of length (R-1) x (C-1) x
prod(dim(x)[strata]). The dim and dimnames methods provide the proper attributes for treating
the coefficients vector as an (R-1) x (C-1) x strata array. as.matrix and as.array methods are
also provided for this purpose.
The confint method computes confidence intervals for the log odds ratios (or for odds ratios, with
log = FALSE). The coeftest method (summary is an alias) prints the asymptotic standard errors,
z tests (standardized log odds ratios), and the corresponding p 
#### Value 
.
Structural zeros: In addition to the options for zero cells provided by correct, the function allows
for structural zeros to be represented as NA in the data argument. NA in the data yields NA as the LOR
estimate, but does not affect other cells.
oddsratio is just an alias to loddsratio for backward compatibility.
Author(s)
Achim Zeileis, * Michael Friendly and David Meyer.
References
A. Agresti (2013), Categorical Data Analysis, 3rd Ed. New York: Wiley.
Fleiss, J. L. (1981). Statistical Methods for Rates and Proportions. 2nd Edition. New York: Wiley.
M. Friendly (2000), Visualizing Categorical Data. SAS Institute, Cary, NC.
Friendly, M., Turner, H" Firth, D., Zeileis, A. (2011). Advances in Visualizing Categorical Data
Using the vcd, gnm and vcdExtra Packages in R. Correspondence Analysis and Related Methods
(CARME 2011). http://www.datavis.ca/papers/adv-vcd-4up.pdf
See Also
plot.loddsratio for some plotting methods; confint for confidence intervals; coeftest for ztests
of significance

#### Examples
```{r}

## artificial example
set.seed(1)
x <- matrix(rpois(5 * 3, 7), ncol = 5, nrow = 3)
dimnames(x) <- list(Row = head(letters, 3), Col = tail(letters, 5))
x_lor <- loddsratio(x)
coef(x_lor)
x_lor
confint(x_lor)
summary(x_lor)
#cases
#data(CoalMiners, package = "vcd")

lor_CM <- loddsratio(CoalMiners)
lor_CM
coef(lor_CM)
confint(lor_CM)
confint(lor_CM, log = FALSE)
#k x 2
lor_Emp <-loddsratio(Employment)
lor_Emp
confint(lor_Emp)
## 4 way tables
data(Punishment, package = "vcd")
lor_pun <- loddsratio(Freq ~ memory + attitude | age + education, data = Punishment)
lor_pun
confint(lor_pun)
summary(lor_pun)
# fit linear model using WLS
lor_pun_df <- as.data.frame(lor_pun)
pun_mod1 <- lm(LOR ~ as.numeric(age) * as.numeric(education),
data = lor_pun_df, weight/ ASE^2)
anova(pun_mod1)
## illustrate ref levels
VA.fem <- xtabs(Freq ~ left + right, subset=gender=="female", data=VisualAcuity)
VA.fem
loddsratio(VA.fem) # profile contrasts
loddsratio(VA.fem, ref=1) # contrasts against level 1
loddsratio(VA.fem, ref=dim(VA.fem)) # contrasts against level 4
```


