---
title: "Ord_plot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```
### Ord Plots

#### Desription

Ord plots for diagnosing discrete distributions.

#### Usage
<pre><code>

Ord_plot(obj, legend = TRUE, estimate = TRUE, tol = 0.1, type = NULL,
xlim = NULL, ylim = NULL, xlab = "Number of occurrences",
ylab = "Frequency ratio", main = "Ord plot", gp = gpar(cex = 0.5),
lwd = c(2,2), lty=c(2,1), col=c("black", "red"),
name = "Ord_plot", newpage = TRUE, pop = TRUE,return_grob = FALSE, ...)

Ord_estimate(x, type = NULL, tol = 0.1)

</code></pre>
#### Arguments

* obj either a vector of counts, a 1-way table of frequencies of counts or a data frame
or matrix with frequencies in the first column and the corresponding counts in
the second column.
* legend logical. Should a legend be plotted?.
* estimate logical. Should the distribution and its parameters be estimated from the data?
See details.
* tol tolerance for estimating the distribution. See details.
* type a character string indicating the distribution, must be one of "poisson", "binomial",
"nbinomial" or "log-series" or NULL. In the latter case the distribution is estimated
from the data. See details.
xlim limits for the x axis.
ylim limits for the y axis.
xlab a label for the x axis.
ylab a label for the y axis.
* main a title for the plot.
* gp a "gpar" object controlling the grid graphical parameters of the points.
lwd, lty vectors of length 2, giving the line width and line type used for drawing the OLS
line and the WLS lines.
* col vector of length 2 giving the colors used for drawing the OLS and WLS lines.
* name name of the plotting viewport.
* newpage logical. Should grid.newpage be called before plotting?
* pop logical. Should the viewport created be popped?
* return_grob logical. Should a snapshot of the display be returned as a grid grob?
* ... further arguments
 passed to grid.points.
x a vector giving intercept and slope for the (fitted) line in the Ord plot.

#### Examples
```{r}

## Simulated data examples

dummy <- rnbinom(1000, size = 1.5, prob = 0.8)
Ord_plot(dummy)
## Real data examples

data("HorseKicks")
data("Federalist")
data("Butterfly")
data("WomenQueue")
## Not run:
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 2)))
pushViewport(viewport(layout.pos.col=1, layout.pos.row=1))

Ord_plot(HorseKicks, main = "Death by horse kicks", newpage = FALSE)
popViewport()
pushViewport(viewport(layout.pos.col=1, layout.pos.row=2))
Ord_plot(Federalist, main = "Instances of 'may' in Federalist papers", newpage = FALSE)
popViewport()
pushViewport(viewport(layout.pos.col=2, layout.pos.row=1))
Ord_plot(Butterfly, main = "Butterfly species collected in Malaya", newpage = FALSE)
popViewport()
pushViewport(viewport(layout.pos.col=2, layout.pos.row=2))
Ord_plot(WomenQueue, main = "Women in queues of length 10", newpage = FALSE)
popViewport(2)
## End(Not run)
## same
mplot(
Ord_plot(HorseKicks, return_grob = TRUE, main = "Death by horse kicks"),
Ord_plot(Federalist, return_grob = TRUE, main = "Instances of 'may' in Federalist papers"),
Ord_plot(Butterfly, return_grob = TRUE, main = "Butterfly species collected in Malaya"),
Ord_plot(WomenQueue, return_grob = TRUE, main = "Women in queues of length 10")
)
```
#### Value
A vector giving the intercept and slope of the weighted OLS line.

#### Details
The Ord plot plots the number of occurrences against a certain frequency ratio (see Friendly (2000)
for details) and should give a straight line if the data comes from a poisson, binomial, negative
binomial or log-series distribution. The intercept and slope of this straight line conveys information
about the underlying distribution.
Ord_plot fits a usual OLS line (black) and a weighted OLS line (red). From the coefficients of the
latter the distribution is estimated by Ord_estimate as described in Table 2.10 in Friendly (2000).

To judge whether a coefficient is positive or negative a tolerance given by tol is used. If none
of the distributions fits well, no parameters are estimated. Be careful with the conclusions from
``Ord_estimate`` as it implements just some simple heuristics!


Author(s)
Achim Zeileis <Achim.Zeileis@R-project.org>
References
J. K. Ord (1967), Graphical methods for a class of discrete distributions, Journal of the Royal
Statistical Society, A 130, 232-238.
* Michael Friendly (2000), Visualizing Categorical Data. SAS Institute, Cary, NC.
