---
title: "vcd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```


Pairs plot panel functions for off-diagonal cells
Off-diagonal Panel Functions for Table Pairs Plot

#### Desription

Off-diagonal panel functions for pairs.table.

#### Usage
<pre><code>

pairs_strucplot(panel = mosaic,
type = c("pairwise", "total", "conditional", "joint"),
legend = FALSE, margins = c(0, 0, 0, 0), labeling = NULL, ...)
pairs_assoc(...)
pairs_mosaic(...)
pairs_sieve(...)

</code></pre>
#### Arguments

panel function to be used for the plots in each cell, such as pairs_assoc, pairs_mosaic,
and pairs_sieve.
type character string specifying the type of independence model visualized in the
cells.
legend logical specifying whether a legend should be displayed in the cells or not.
margins margins inside each cell (see strucplot).
labeling labeling function or labeling-generating function (see strucplot).
... pairs_mosaic, pairs_assoc, and pairs_sieve: parameters passed to pairs_strucplot.
pairs_strucplot: other parameters passed to panel function.
Details
These functions really just wrap assoc, sieve, and mosaic by basically inhibiting labeling and
legend-drawing and setting the margins to 0.
Value
A function with 
</code></pre>
#### Arguments
:
x contingency table.
i, j cell coordinates.
Author(s)
David Meyer <David.Meyer@R-project.org>
pairs.table 87
References
Cohen, A. (1980), On the graphical display of the significant components in a two-way contingency
table. Communications in Statistics-Theory and Methods, A9, 1025-1041.
Friendly, M. (1992), Graphical methods for categorical data. SAS User Group International Conference
Proceedings, 17, 190-200. http://datavis.ca/papers/sugi/sugi17.pdf
See Also
pairs.table, pairs_text, pairs_barplot, assoc, mosaic

#### Examples
```{r}

data("UCBAdmissions")
data("PreSex")
pairs(PreSex)
pairs(UCBAdmissions)
pairs(UCBAdmissions, upper_panel_args = list(shade = FALSE))
pairs(UCBAdmissions, lower_panel = pairs_mosaic(type = "conditional"))
pairs(UCBAdmissions, upper_panel = pairs_assoc)
```

