---
title: "struc_assoc"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


struc_assoc Core-generating Function for Association Plots

#### Desription

Core-generating function for strucplot returning a function producing association plots.

#### Usage
<pre><code>

struc_assoc(compress = TRUE, xlim = NULL, ylim = NULL,
yspace = unit(0.5, "lines"), xscale = 0.9, gp_axis = gpar(lty = 3))

</code></pre>
#### Arguments

compress logical; if FALSE, the space between the rows (columns) are chosen such that
the total heights (widths) of the rows (column) are all equal. If TRUE, the space
between the rows and columns is fixed and hence the plot is more "compressed".
xlim eithe k matrix of doubles, k the number of total columns of the plot, or
a recycled vector from which such a matrix will be constructed. The columns
of xlim correspond to the columns of the association plot, the rows describe
the column ranges (minimums in the first row, maximums in the second row).
If xlim is NULL, the ranges are determined from the residuals according to
compress (if TRUE: widest range from each column, if FALSE: from the whole
association plot matrix).
ylim eithe k matrix of doubles, k the number of total rows of the plot, or a
recycled vector from which such a matrix will be constructed. The columns of
ylim correspond to the rows of the association plot, the rows describe the column
ranges (minimums in the first row, maximums in the second row). If ylim
is NULL, the ranges are determined from the residuals according to compress (if
TRUE: widest range from each row, if FALSE: from the whole association plot
matrix).
xscale scale factor resizing the tile's width, thus adding additional space between the
tiles.
yspace object of class "unit" specifying additional space separating the rows.
* `` gp_axis`` :  ar" specifying the visual aspects of the tiles' baseline.
Details
This function is usually called by strucplot (typically when called by assoc) and returns a function
used by strucplot to produce association plots.
Value
A function with 
</code></pre>
#### Arguments
:
residuals table of residuals.
observed not used by struc_assoc.
expected table of expected frequencies.
spacing object of class "unit" specifying the space between the tiles.
gp list of gpar objects used for the drawing the tiles.
split_vertical vector of logicals indicating, for each dimension of the table, the split direction.
Author(s)
David Meyer <David.Meyer@R-project.org>
struc_mosaic 123

#### Examples
```{r}

## UCB Admissions
data("UCBAdmissions")
ucb <- aperm(UCBAdmissions)
## association plot for conditional independence
strucplot(ucb, expecteDept * (Admit + Gender),
core = struc_assoc(ylim = c(-4, 4)), labeling_args = list(abbreviate = c(Admit = 3)))
```

#### References
Cohen, A. (1980), On the graphical display of the significant components in a two-way contingency
table. Communications in Statistics-Theory and Methods, A9, 1025-1041.
Friendly, M. (1992), Graphical methods for categorical data. SAS User Group International Conference
Proceedings, 17, 190-200. http://datavis.ca/papers/sugi/sugi17.pdf
Meyer, D., Zeileis, A., and Hornik, K. (2006), The strucplot framework: Visualizing multi-way contingency
tables with vcd. Journal of Statistical Software, 17(3), 1-48. URL http://www.jstatsoft.org/v17/i03/
and available as vignette("strucplot").
See Also
assoc, strucplot, structable
