---
title: "vcd"
author: "Kevin O'Brien"
date: "27 December 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


structable Structured Contingency Tables

#### Desription

This function produces a 'flat' representation of a high-dimensional contingency table constructed
by recursive splits (similar to the construction of mosaic displays).

#### Usage
<pre><code>

## S3 method for class 'formula'
structable(formula, data,
direction = NULL, split_vertical = NULL, ..., subset, na.action)
## Default S3 method:
structable(..., direction = NULL, split_vertical = FALSE)

</code></pre>
#### Arguments

formula a formula object with possibly both left and right hand sides specifying the column
and row variables of the flat table.
data a data frame, list or environment containing the variables to be cross-tabulated,
or an object inheriting from class table.
subset an optional vector specifying a subset of observations to be used. Ignored if
data is a contingency table.
na.action a function which indicates what should happen when the data contain NAs. Ignored
if data is a contingency table
120 structable
... R objects which can be interpreted as factors (including character strings), or a
list (or data frame) whose components can be so interpreted, or a contingency
table object of class "table" or "ftable".
split_vertical logical vector indicating, for each dimension, whether it should be split vertically
or not (default: FALSE). 
#### Value 
 are recycled as needed. If the argument is
of length 1, the value is alternated for all dimensions. Ignored if direction is
provided.
direction character vector alternatively specifying the splitting direction ("h" for horizontal
and "v" for vertical splits). 
#### Value 
 are recycled as needed. If the argument is
of length 1, the value is alternated for all dimensions.
Details
This function produces textual representations of mosaic displays, and thus 'flat' contingency tables.
The formula interface is quite similar to the one of ftable, but also accepts the mosaic-like formula
interface (empty left-hand side). Note that even if the ftable interface is used, the split_vertical
or direction argument is needed to specify the order of the horizontal and vertical splits. If
pretabulated data with a Freq column is used, than the left-hand side should be left empty-the
Freq column will be handled correctly.
"structable" objects can be subset using the [ and [[ operators, using either level indices or
names (see 
#### Examples
```{r}
). The corresponding replacement functions are available as well. In addition,
appropriate aperm, cbind, rbind, length, dim, and is.na methods do exist.
Value
An object of class "structable", inheriting from class "ftable", with the splitting in
</code></pre>
#### Format 
ion
("split_vertical") as additional attribute.
Author(s)
David Meyer <David.Meyer@R-project.org>
References
Meyer, D., Zeileis, A., and Hornik, K. (2006), The strucplot framework: Visualizing multi-way contingency
tables with vcd. Journal of Statistical Software, 17(3), 1-48. URL http://www.jstatsoft.org/v17/i03/
and available as vignette("strucplot").
See Also
strucplot, mosaic, ftable

#### Examples
```{r}

structable(Titanic)
structable(Titanic, split_vertical = c(TRUE, TRUE, FALSE, FALSE))
structable(Titanic, direction = c("h","h","v","v"))
structable(Sex + Class ~ Survived + Age, data = Titanic)
struc_assoc 121
## subsetting of structable objects
(hec <- structable(aperm(HairEyeColor)))
## The "[" operator treats structables as a block-matrix and selects parts of the matrix:
hec[1]
hec[2]
hec[1,c(2,4)]
hec["Male",c("Blue","Green")]
## replacement funcion:
tmp <- hec
(tmp[1,2:3] <- tmp[2,c(1,4)])
## In contrast, the "[[" operator treats structables as two-dimensional
## lists. Indexing conditions on specified levels and thus reduces the dimensionality:
## seek subtables conditioning on levels of the first dimension:
hec[[1]]
hec[[2]]
## Seek subtable from the first two dimensions, given the level "Male"
## of the first variable, and "Brown" from the second
## (the following two commands are equivalent):
hec[["Male"]][["Brown"]]
hec[[c("Male","Brown")]]
## Seeking subtables by conditioning on row and/or column variables:
hec[["Male","Hazel"]]
hec[[c("Male","Brown"),]]
hec[[c("Male","Brown"),"Hazel"]]
## a few other operations
t(hec)
dim(hec)
dimnames(hec)
as.matrix(hec)
length(hec)
cbind(hec[,1],hec[,3])
as.vector(hec) ## computed on the _multiway_ table
as.vector(unclass(hec))
