---
title: "vcd::table2d_summary"
author: "Kevin O'Brien"
date: "5 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
table2d_summary of a 2-way Table

#### Desription

Prints a 2-way contingency table along with percentages, marginal, and conditional distributions.

#### Usage
<pre><code>

table2d_summary(object, margins = TRUE, percentages = FALSE,
conditionals = c("none", "row", "column"), chisq.test = TRUE, ...)

</code></pre>
#### Arguments

objec c-contingency table
margins if TRUE, marginal distributions are computed.
percentages if TRUE, relative frequencies are computed.
conditionals if not "none", the conditional distributions, given the row/column factor, are
computed.
chisq.test if TRUE, a chi-squared test of independence is carried out.
... currently not used.
Value
Returns invisiblk table, k depending on the amount of choices (at most 3).
Author(s)
David Meyer <David.Meyer@R-project.org>
See Also
mar_table, prop.table, independence_table

#### Examples
```{r}

data("UCBAdmissions")
table2d_summary(margin.table(UCBAdmissions, 1:2))

```
