---
title: "vcd"
author: "Kevin O'Brien"
date: "5 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
WeldonDice Weldon's Dice Data

#### Desription

Data from Pearson (1900) about the frequency of 5s and 6s in throws of 12 dice. Weldon tossed the
dice 26,306 times and reported his results in a letter to Francis Galton on 1894-02-02.

#### Usage
<pre><code>

data("WeldonDice")

</code></pre>
#### Format 

A 1-way table giving the frequency ooin 26,306 throws of 12 dice where 10 indicates '10
or more' 5s or 6s. The variable and its levels are
WomenQueue 137
No Name Levels
1 n56 0, 1. , 10
Source
M. Friendly (2000), Visualizing Categorical Data, pages 20-21.
References
K. Pearson (1900), On the criterion that a given system of deviations from the probable in the case
of a correlated system of variables is such that it can be reasonably supposed to have arisen by
random sampling, Philosophical Magazine, 50 (5th series), 157-175.
M. Friendly (2000), Visualizing Categorical Data. SAS Institute, Cary, NC.

#### Examples
```{r}

data("WeldonDice")
gf <- goodfit(WeldonDice, type = "binomial")
summary(gf)
plot(gf)
```
