---
title: "vcd::WomenQueue"
author: "Kevin O'Brien"
date: "5 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 Women in Queues

#### Desription

Data from Jinkinson \& Slater (1981) and Hoaglin \& Tukey (1985) reporting the frequency distribution
of females in 100 queues of length 10 in a London Underground station.

#### Usage
<pre><code>

data("WomenQueue")

</code></pre>
#### Format 

A 1-way table giving the number of women in 100 queues of length 10. The variable and its levels
are
No Name Levels
1 nWomen 0, 1. , 10
Source
M. Friendly (2000), Visualizing Categorical Data, pages 19-20.
138 woolf_test
References
D. C. Hoaglin \& J.W. Tukey (1985), Checking the shape of discrete distributions. In D. C. Hoaglin,
F. Mosteller, J. W. Tukey (eds.), Exploring Data Tables, Trends and Shapes, chapter 9. John Wiley
\& Sons, New York.
R. A. Jinkinson \& M. Slater (1981), Critical discussion of a graphical method for identifying
discrete distributions, The Statistician, 30, 239-248.
M. Friendly (2000), Visualizing Categorical Data. SAS Institute, Cary, NC.

#### Examples
```{r}

data("WomenQueue")
gf <- goodfit(WomenQueue, type = "binomial")
summary(gf)
plot(gf)
```
