---
title: "Untitled"
output: html_document
---

US Counties Shapefile
=======================

* Each county or statistically equivalent entity is assigned a three-character numeric Federal Information Processing Series (FIPS) code based on alphabetical sequence that is unique within state and an eight-digit National Standard feature identifier.

<pre><code>
library(tigris)
</code></pre>


library(ggplot2)
data(midwest)
head(midwest)
table(midwest$state)
table(midwest$state,midwest$county)
table(midwest$county)


setwd("C:/Users/pc/Desktop/SF")

USAZip <- "http://biogeo.ucdavis.edu/data/diva/adm/USA_adm.zip"

download.file(USAZip, destfile="USAmap.zip")

unzip("USAmap.zip")

# Put this all in a folder called "USAmap"

USA1 <- sf::read_sf("USAmap/USA_adm1.shp")


#############################################

head(USA1)

# Alaska? 
# Two separate geometries for simplicity?
# We are not using it here.

str_sub(USA1$VARNAME_1, 1, 2)

Code <- str_sub(USA1$VARNAME_1, 1, 2)

table(Code)

# "Co?"


# USA1$NAME_1 : full names

cbind(USA1$NAME_1,Code)
############################################

# Probably quickest to fix these manually

# This approach doesnt work, but worth a look

str_sub(USA1$NAME_1, 1, 1) %>% toupper

str_sub(USA1$NAME_1, -1, -1) %>% toupper

paste(str_sub(USA1$NAME_1, 1, 1),str_sub(USA1$NAME_1, -1, -1)) %>% toupper

paste(str_sub(USA1$NAME_1, 1, 1),str_sub(USA1$NAME_1, -1, -1),sep="") %>% toupper

# Dreadful (in this instance) !!!

############################################

# Yeah Easiest to fix these manually, and keep as a lookup table.

stateCodes <- cbind(USA1$NAME_1,Code)

# data.entry(stateCodes)

#############################################

# USA Counties

USA2 <- sf::read_sf("USAmap/USA_adm2.shp")

data(midwest)

head(midwest)

midwest$county <- toupper(midwest$county)

head(midwest)

midwest$percwhite

midwest$percwhite %>% summary


* https://rdrr.io/cran/tigris/man/counties.html

Midwest demographics
=========================




Midwest demographics: Demographic information of midwest counties from 2000 US census

<pre><code>
library(ggplot2)
</code></pre>


https://ggplot2.tidyverse.org/reference/midwest.html


Demographic information of midwest counties from 5 US states


(June 2018: Make sure to use the development version of ggplot2 for this exercise.)


```{r}
# 1. For this exercise we will use the following R packages
library(ggplot2) # Development Version
library(dplyr)
library(magrittr)

```


```{r}
# 2. Load the data set and carry out some preliminary analysis
data(midwest)
  # Preliminary Analysis Functions
head(midwest)
```

```{r}
# dim(midwest)
# class(midwest)
# names(midwest)
# help(midwest)
```

```{r}

# 3. Import the USA counties shape file
UScounties <- sf::read_sf("USA_adm2.shp") 
  # some More Analysis
head(UScounties)
class(UScounties)
names(UScounties)
help(UScounties)
```


```{r}
# 4. Matching up the two data sets by State - Part 1
table(midwest$state)
table(UScounties$NAME_1)
```

```{r}
# 5. Get rid of non-relevant columns
table(UScounties$ID_0)
table(UScounties$ISO)
table(UScounties$TYPE_2)
table(UScounties$ENGTYPE_2)
table(UScounties$NL_NAME_2)
table(UScounties$VARNAME_2)
UScounties <- UScounties %>% select(-c(ID_0,ISO, NL_NAME_2))
```


```{r}
# 6. Subset it to midwest states using dplyr
MWcounties <- UScounties %>% filter(NAME_1 %in% c("Ohio","Michigan","Indiana","Wisconsin","Illinois"))
```


```{r}
# 7. Lets see how things match up
dim(midwest)
table(midwest$states)
dim(MWcounties)
table(MWcounties$NAME_1)
table(MWcounties$ENGTYPE_2)
```


```{r}
# 8. Lets Join the tables. Some work to do first
  # Add state column to MWcounties using a Lookup table
statecode <- data.frame(
   NAME_1 = c("Ohio","Michigan","Indiana","Wisconsin","Illinois"),
   state  = c("OH", "MI", "IN" ,"WI", "IL")
   )
   # Carry out a "SQL Join" operation
MWcounties <- left_join(MWcounties,statecode)
   # can get rid of statecode dataframe now 
   # Lets clean up the dataframe using dplyr::select
MWcounties <- MWcounties %>% select(1:NAME_1,state, NAME_2 , everything())
   # Also  - get rid of Water features
MWcounties <- MWcounties %>% filter(ENGTYPE_2 != Water body")
head(MWcounties)
  # So Far So Good
```


```{r}
# 9. Put the counties names in the same case ( arbitrarily - all upper case)
 
 # Lets check midwest data set
head(midwest$county)
head(MWcounties$NAME_2)
midwest$county    <- toupper(midwest$county)
MWcounties$NAME_2 <- toupper(MWcounties$NAME_2)
```


```{r}
# 10. Lets Join the tables
  # First Lets do some checks!!
full_join(midwest, MWcounties, by= c("state"="state", "county" = "NAME_2")) %>% dim()
inner_join(midwest, MWcounties, by= c("state"="state", "county" = "NAME_2")) %>% dim()
anti_join(midwest, MWcounties, by= c("state"="state", "county" = "NAME_2")) %>% dim()
anti_join(MWcounties, midwest,  by= c("state"="state", "NAME_2" = "county")) %>% dim()
  # How many problem cases have we?
  # May Have to fix them manually
```


```{r}
# 11. Sort both data sets by state and county name and then do some checks
midwest <- midwest %>% arrange( state,county)
MWcounties <- MWcountues %>% arrange(state,NAME_2)
mean(midwest$state == MWcounties$state)
sum(midwest$county == MWcounties$NAME_2)
sum(midwest$county != MWcounties$NAME_2)
 #  Alphabetial order is broken
```


```{r}
# 12. OK lets fix it with "gsub"
 # A Basic understanding of regular expressions is recommended
midwest$county <- gsub("ST ","SAINT ",midwest$county)
midwest$county <- gsub("DU","DU ",midwest$county)
midwest$county <- gsub("LA","LA ",midwest$county)
```
