---
title: "USA States Maps"
subtitle: "Working with Shapefiles"
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

### Load the Relevant R packages

```{r warning = FALSE, message = FALSE}
library(sf)
library(ggplot2)
library(tidyverse)
```


### US States Shapefile

The following code segment can be used to download a USA administrative areas shapefile as a zip file into the working directory, and then unzipped into a sub-directory.

<pre><code>
USAZip <- "http://biogeo.ucdavis.edu/data/diva/adm/USA_adm.zip"

download.file(USAZip, destfile="USAmap.zip")

# Put this all in a folder called "USAmap"

unzip("USAmap.zip", exdir = "USAmap")
</code></pre>

* Check the folder to see if the unzipping operation was successful. 
* If it was successful, we can delete the Zip file.

<pre><code>
list.files("USAmap")
file.remove("USAmap.zip")
</code></pre>

* We can now upload the shapefile into R
* ***USA_adm1*** refers to the tier 1 administrative units of the USA, i.e the states.
* Similarly ***USA_adm2*** refers to the tier 2 administrative units of the USA, i.e the counties within each state.

```{r}
UScounties <- sf::read_sf("USAmap/USA_adm2.shp")
```


### Inspecting the Data

```{r}
dim(UScounties)
```

```{r}
head(UScounties)
```

### Select Columns to Remove

```{r}
# 5. Get rid of non-relevant columns
table(UScounties$ID_0)
table(UScounties$ISO)
table(UScounties$TYPE_2)
table(UScounties$ENGTYPE_2)
table(UScounties$NL_NAME_2)
table(UScounties$VARNAME_2)
UScounties <- UScounties %>% select(-c(ID_0,ISO, NL_NAME_2))
```


### Counties in Midwest States

```{r}

MWcounties <- UScounties %>% 
  filter(NAME_1 %in% c("Ohio","Michigan",
                       "Indiana","Wisconsin","Illinois"))
```


```{r}
dim(MWcounties)
```


```{r}
table(MWcounties$NAME_1)
```


```{r}
table(MWcounties$ENGTYPE_2)
```

```{r}
# Also - get rid of Water features
MWcounties <- MWcounties %>% filter(ENGTYPE_2 != "Water body")
```

### Map

```{r}
ggplot(MWcounties) + geom_sf() + theme_classic()
```

```{r}
head(MWcounties)
```

```{r}
data("midwest")
head(midwest)
```


```{r}
List1 <- MWcounties %>% pull(NAME_2) %>% toupper()
List2 <- midwest %>% pull(county) %>% toupper()
```


```{r}
setdiff(List2,List1)
setdiff(List1,List2)
```
