---
title: "USA States Maps"
subtitle: "Working with Shapefiles"
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

### Load the Relevant R packages

```{r warning = FALSE, message = FALSE}
library(sf)
library(ggplot2)
library(tidyverse)
```


### US States Shapefile

The following code segment can be used to download a USA administrative areas shapefile as a zip file into the working directory, and then unzipped into a sub-directory.

<pre><code>
USAZip <- "http://biogeo.ucdavis.edu/data/diva/adm/USA_adm.zip"

download.file(USAZip, destfile="USAmap.zip")

# Put this all in a folder called "USAmap"

unzip("USAmap.zip", exdir = "USAmap")
</code></pre>

* Check the folder to see if the unzipping operation was successful. 
* If it was successful, we can delete the Zip file.

<pre><code>
list.files("USAmap")
file.remove("USAmap.zip")
</code></pre>

* We can now upload the shapefile into R
* ***USA_adm1*** refers to the tier 1 administrative units of the USA, i.e the states.
* Similarly ***USA_adm2*** refers to the tier 2 administrative units of the USA, i.e the counties within each state.

```{r}
USA1 <- sf::read_sf("USAmap/USA_adm1.shp")
```

### Inspecting the Data

* The names of the states are recorded in the ***Names_1*** Variable.
* Alaska is listed twice (There is a good reason for that).

```{r}
head(USA1)
```

### USA State Map

* Notice the peculiar consequence of Alaska crossing the international meridian.

```{r}
ggplot(USA1) + geom_sf() + theme_bw()
```

### Southern States

* Let's subset the dataset to feature only a selection of states from the South of the USA.

```{r}
SOUTH <- USA1 %>% filter(NAME_1 %in% c("Alabama","Mississippi","Georgia","Florida"))
ggplot(SOUTH)+ geom_sf() + theme_bw()
```
