
\##*{Exercise:  The Air Quality Data Set}

A data frame with 154 observations on 6 variables.


* [Ozone]numericOzone (ppb)
* [Solar.R]numericSolar R (lang)
* [Wind]numericWind (mph)
* [Temp]numericTemperature (degrees F)
* [Month]numericMonth (1--12)
* [Day]numericDay of month (1--31)

==================================== %

```{r}

tail(airquality)
help(airquality)

```

==================================

\###{Exercises}
\begin{enumerate}
*  For each variable - how many missing values are there?
*  How may complete cases are there? 
*  What is the variance of each of the continuous variables?
*  How many complete cases are there (i.e. no missing values)
*  If you dont include the "wind" variable, how many complete cases are there?
\end{enumerate}
==================================
\###{Complete Cases}
For a data frame of $n$ cases, the command \texttt{complete.cases()} returns an $n$ element logical vector.

Logical vectors ( \texttt{TRUE} and \texttt{FALSE} ) can be converted into the corresponding numeric values (1 and 0 respectively) by using the \texttt{as.numeric()} command.


```{r}
> X <- c(T,T,F,F,T)
>
> as.numeric(X)
[1] 1 1 0 0 1
>
```


