

% Load any packages needed for this document

\###*{Autocorrelation} 

*  Adjacent residuals should not be correlated with each other (\textbf{autocorrelation}). 
*  If you can use one residual to predict the next residual, there is some predictive information present that is not captured by the predictors. 
*  Typically, this situation involves time-ordered observations. 


\begin{figure}[h!]
\centering
\includegraphics[width=0.7\linewidth]{autocorrelation1}
\caption{(disregard the titles)}

\end{figure}

*  For example, if a residual is more likely to be followed by another residual that has the same sign, adjacent residuals are positively correlated. 
*  You can include a variable that captures the relevant time-related information, or use a time series analysis.

 

\###*{Durbin-Watson Test for Autocorrelated Errors}
The \textbf{\textit{Durbin-Watson} }procedure is commonly used to to test for autocorrelation of residuals. To perform this test, we use the \texttt{durbinWatsonTest()} from the car R package. All you have to do is to specify the name of the fitted mode.


```{r}

 FitMod <- lm(mpg~wt+cyl,data=mtcars)

# library(car)
durbinWatsonTest(FitMod)

```


*  The null hypothesis can simply be stated as "There is no autocorrelation present in the residuals".. 
*  The \texttt{R} code output provides a $p-$value to base a determination on.



```{r}
> durbinWatsonTest(FitMod)
 lag Autocorrelation D-W Statistic p-value
   1       0.1302185      1.671096   0.252
 Alternative hypothesis: rho != 0
```


% http://polisci.msu.edu/jacoby/icpsr/regress3/lectures/week3/11.Outliers.pdf


