---
title: "Cox regression in R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load data

```{r}
## Load survival package
library(survival)
## List datasets in survival package
## data(package = "survival")
```
```{r}
## Load lung data
data(lung)

## Show first 6 rows
head(lung)
```


### NCCTG Lung Cancer Data

#### Description:
Survival in patients with advanced lung cancer from the North
     Central Cancer Treatment Group.  Performance scores rate how well
     the patient can perform usual daily activities.

inst:       Institution code
time:       Survival time in days
status:     censoring status 1=censored, 2=dead
age:        Age in years
sex:        Male=1 Female=2
ph.ecog:    ECOG performance score (0=good 5=dead)
ph.karno:   Karnofsky performance score (bad=0-good=100) rated by physician
pat.karno:  Karnofsky performance score as rated by patient
meal.cal:   Calories consumed at meals
wt.loss:    Weight loss in last six months

### Create a survival object

```{r}
## Add survival object. status == 2 is death
lung$SurvObj <- with(lung, Surv(time, status == 2))

### Check data
head(lung)
```

Kaplan-Meier estimator without grouping
```{r}
## Kaplan-Meier estimator. The "log-log" confidence interval is preferred.
km.as.one <- survfit(SurvObj ~ 1, data = lung, conf.type = "log-log")
km.by.sex <- survfit(SurvObj ~ sex, data = lung, conf.type = "log-log")

## Show object
km.as.one
```


## See survival estimates at given time (lots of outputs)
```{r}
summary(km.as.one)
```

```{r}
summary(km.by.sex)
```

```{r}
## Plotting without any specification
plot(km.as.one)
```

```{r}
plot(km.by.sex)
```

```{r}
## Without
plot(km.as.one, mark.time = FALSE)
```

```{r}
plot(km.by.sex, mark.time = F)
```

### Cox regression using coxph

```{r}
## Fit Cox regression: age, sex, Karnofsky performance score, wt loss
res.cox1 <- coxph(SurvObj ~ age + sex + ph.karno + wt.loss, data =  lung)
res.cox1
```

```{r}
## Check for violation of proportional hazard (constant HR over time)
(res.zph1 <- cox.zph(res.cox1))

```

```{r}
## Displays a graph of the scaled Schoenfeld residuals, along with a smooth curve.
plot(res.zph1)
```

```{r}

## residuals(res.cox1, type = "scaledsch")
## c("martingale", "deviance", "score", "schoenfeld", "dfbeta", "dfbetas", "scaledsch","partial")

## Check Karnofsky performance score (only 6 discrete values)
table(lung$ph.karno)
```

```{r}

## Solution 1: Stratify
res.cox1.strata <- coxph(SurvObj ~ age + sex + strata(ph.karno) + wt.loss, data =  lung)
cox.zph(res.cox1.strata)
```

```{r}
## Solution 2: Time-varying effect
## Event status variable necessary
lung$event <- (lung$status == 2)
```

```{r}
## Erase the survival object (Leaving it in data frames breaks the conversion)
#lung2 <- lung2 %>% select(-SurvObj)
```

```{r}
## Counting process format creation
lung.split <- survSplit(data    = lung,
                        cut     = c(200,400), # vector of timepoints to cut at
                        end     = "time",  # character string with name of event time variable
                        event   = "event", # character string with name of censoring indicator
                        start   = "start", # character string with name of start time variable (created)
                        id      = "id",    # character string with name of new id variable to create
                        zero    = 0,       # If start doesn't already exist, used as start
                        episode = NULL     # character string with name of new episode variable (optional)
)
```

```{r}                
## Make id numeric
lung.split$id <- as.numeric(lung.split$id)
```

```{r}
## Reorder by id, then start time
library(doBy)
lung.split <- orderBy( ~ id + start, lung.split)
```

```{r}
## Recreate SurbObj
lung.split$SurvObj <- with(lung.split, Surv(time = (start), time2 = time, event = event))
```

```{r}
## Check
head(lung.split, 10)
```

```{r}
## Time-varying effect of baseline variable by including interaction with interval
res.cox1.strata <- coxph(SurvObj ~ age + sex + ph.karno + ph.karno:factor(start) + wt.loss + cluster(id),
                         data =  lung.split)
summary(res.cox1.strata)
```

```{r}
rms::survplot() function
## Load rms package
library(rms)
```

```{r}
## without specification
survplot(km.by.sex)
plot of chunk unnamed-chunk-6
```


```{r}
## Minimum decent
survplot(km.by.sex,
         conf = "none")
```


```{r}
## Full-fledged grammer
survplot(fit  = km.by.sex,
         conf = c("none","bands","bars")[1],
         xlab = "", ylab = "Survival",
         ## xlim(0,100),
         label.curves = TRUE,                     # label curves directly
         ## label.curves = list(keys = "lines"),  # legend instead of direct label
         levels.only  = FALSE,                    # show only levels, no label
         abbrev.label = FALSE,                    # if label used, abbreviate
         ## fun = function(x) {1 - x},            # Cumulative probability plot         
         loglog   = FALSE,                        # log(-log Survival) plot
         logt     = FALSE,                        # log time
         time.inc = 100,                          # time increment
         dots     = TRUE,                         # dot grid
         n.risk   = TRUE,                         # show number at risk
         ## srt.n.risk = 0, sep.n.risk = 0.056, adj.n.risk = 1,
         ## y.n.risk = 0, cex.n.risk = 0.6
         )
```


```{r}


## Without dots
survplot(fit  = km.by.sex,
         conf = c("none","bands","bars")[1],
         xlab = "", ylab = "Survival",
         ## xlim(0,100),
         label.curves = TRUE,                     # label curves directly
         ## label.curves = list(keys = "lines"),  # legend instead of direct label
         levels.only  = FALSE,                    # show only levels, no label
         abbrev.label = FALSE,                    # if label used, abbreviate
         ## fun = function(x) {1 - x},            # Cumulative probability plot         
         loglog   = FALSE,                        # log(-log Survival) plot
         logt     = FALSE,                        # log time
         time.inc = 100,                          # time increment
         dots     = FALSE,                         # dot grid
         n.risk   = TRUE,                         # show number at risk
         ## srt.n.risk = 0, sep.n.risk = 0.056, adj.n.risk = 1,
         ## y.n.risk = 0, cex.n.risk = 0.6
         )
```


```{r}


## Different time incremente
survplot(fit  = km.by.sex,
         conf = c("none","bands","bars")[1],
         xlab = "", ylab = "Survival",
         ## xlim(0,100),
         label.curves = TRUE,                     # label curves directly
         ## label.curves = list(keys = "lines"),  # legend instead of direct label
         levels.only  = FALSE,                    # show only levels, no label
         abbrev.label = FALSE,                    # if label used, abbreviate
         ## fun = function(x) {1 - x},            # Cumulative probability plot         
         loglog   = FALSE,                        # log(-log Survival) plot
         logt     = FALSE,                        # log time
         time.inc = 300,                           # time increment
         dots     = FALSE,                        # dot grid
         n.risk   = TRUE,                         # show number at risk
         ## srt.n.risk = 0, sep.n.risk = 0.056, adj.n.risk = 1,
         ## y.n.risk = 0, cex.n.risk = 0.6
         )
```


```{r}



## Plot cumulative probability F(t) = 1 - S(t)
survplot(fit  = km.by.sex,
         conf = c("none","bands","bars")[1],
         xlab = "", ylab = "Cumulative Incidence",
         ## xlim(0,100),
         label.curves = TRUE,                     # label curves directly
         ## label.curves = list(keys = "lines"),  # legend instead of direct label
         levels.only  = FALSE,                    # show only levels, no label
         abbrev.label = FALSE,                    # if label used, abbreviate
         fun = function(x) {1 - x},             # Cumulative probability plot         
         loglog   = FALSE,                        # log(-log Survival) plot
         logt     = FALSE,                        # log time
         time.inc = 100,                          # time increment
         dots     = FALSE,                        # dot grid
         n.risk   = TRUE,                         # show number at risk
         ## srt.n.risk = 0, sep.n.risk = 0.056, adj.n.risk = 1,
         ## y.n.risk = 0, cex.n.risk = 0.6
         )
```


```{r}


## Change position of number at risk
survplot(fit  = km.by.sex,
         conf = c("none","bands","bars")[1],
         xlab = "", ylab = "Survival",
         ## xlim(0,100),
         label.curves = TRUE,                     # label curves directly
         ## label.curves = list(keys = "lines"),  # legend instead of direct label
         levels.only  = FALSE,                    # show only levels, no label
         abbrev.label = FALSE,                    # if label used, abbreviate
         ## fun = function(x) {1 - x},            # Cumulative probability plot         
         loglog   = FALSE,                        # log(-log Survival) plot
         logt     = FALSE,                        # log time
         time.inc = 100,                          # time increment
         dots     = FALSE,                        # dot grid
         n.risk   = TRUE,                         # show number at risk
         ## srt.n.risk = 0, sep.n.risk = 0.056, adj.n.risk = 1,
         y.n.risk = -0.2, cex.n.risk = 0.6
         )
```





References
Fox Cox PH model: http://cran.r-project.org/doc/contrib/Fox-Companion/appendix-cox-regression.pdf