---
title: "New York Air Quality"
author: "R Workshop"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  rmdformats::readthedown:
    toc_depth: 3
    use_bookdown: TRUE
    #code_folding: hide
    fig_caption: TRUE


html_document:
  fig_caption: yes
  theme: flatly #sandstone #spacelab #
  highlight: pygments
  number_sections: TRUE
  toc: TRUE
  toc_width: 40
  toc_depth: 2
  toc_float:
    smooth_scroll: FALSE


---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
```
  
```{r import_libraries, echo=FALSE, message=FALSE}
# library(conflicted)
library(tidyverse)
# library(scales)
# library(cowplot)
# library(magrittr)

# library(rlang)
# library(stringr)
# library(glue)
# library(purrr)
# library(furrr)
# library(arules)
# library(arulesViz)
# library(DT)
# library(tidygraph)
# library(rfm)
#source("lib_utils.R")
# resolve_conflicts(
#   c("magrittr", "rlang", "dplyr", "readr", "purrr", "ggplot2", "arules",
#     "Matrix", "DT")
# )

knitr::opts_chunk$set(
  tidy       = FALSE,
  cache      = FALSE,
  warning    = FALSE,
  message    = FALSE,
  fig.height =     8,
  fig.width  =    12
)

options(
  width = 90L,
  warn  = 1
)

#theme_set(theme_cowplot())
set.seed(1234)
#plan(multisession)
```



New York Air Quality Measurements
==============================================

## Introduction

### Description
Daily air quality measurements in New York, May to September 1973.


## Exercise:  The Air Quality Data Set

A data frame with 154 observations on 6 variables.


* [Ozone]numericOzone (ppb)
* [Solar.R]numericSolar R (lang)
* [Wind]numericWind (mph)
* [Temp]numericTemperature (degrees F)
* [Month]numericMonth (1--12)
* [Day]numericDay of month (1--31)





Usage
airquality

#### Format
A data frame with 153 observations on 6 variables.

[,1]OzonenumericOzone (ppb)
[,2]Solar.RnumericSolar R (lang)
[,3]WindnumericWind (mph)
[,4]TempnumericTemperature (degrees F)
[,5]MonthnumericMonth (1--12)
[,6]DaynumericDay of month (1--31)

#### Details

* Daily readings of the following air quality values for May 1, 1973 (a Tuesday) to September 30, 1973.

* Ozone: Mean ozone in parts per billion from 1300 to 1500 hours at Roosevelt Island

* Solar.R: Solar radiation in Langleys in the frequency band 4000–7700 Angstroms from 0800 to 1200 hours at Central Park

* Wind: Average wind speed in miles per hour at 0700 and 1000 hours at LaGuardia Airport

* Temp: Maximum daily temperature in degrees Fahrenheit at La Guardia Airport.




## Data Summary


```{r airquality}
summary(airquality)
```

## Including Plots

```{r}


require(graphics)
pairs(airquality, panel = panel.smooth, main = "airquality data")
```
You can also embed plots, for example:

```{r pressure, echo=FALSE}
kable(airquality)
```


```{r}

tail(airquality)
help(airquality)

```


###{Exercises}

*  For each variable - how many missing values are there?
*  How may complete cases are there? 
*  What is the variance of each of the continuous variables?
*  How many complete cases are there (i.e. no missing values)
*  If you dont include the "wind" variable, how many complete cases are there?

### Complete Cases 
For a data frame of $n$ cases, the command \texttt{complete.cases()} returns an $n$ element logical vector.

Logical vectors ( \texttt{TRUE} and \texttt{FALSE} ) can be converted into the corresponding numeric values (1 and 0 respectively) by using the \texttt{as.numeric()} command.


```{r}
 X <- c(T,T,F,F,T)

 as.numeric(X)
```



Appendices
=====================================================

## Source

### Source 
The data were obtained from the New York State Department of Conservation (ozone data) and the National Weather Service (meteorological data).

### References
Chambers, J. M., Cleveland, W. S., Kleiner, B. and Tukey, P. A. (1983) Graphical Methods for Data Analysis. Belmont, CA: Wadsworth.

