---
title: "The Cheddar Dataset"
subtitle: "Advanced Regression Models"
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(faraway)
library(MASS)       # For stepAIC
library(modelr)     # For rmse(), mae(), etc.
library(broom)      # For tidy(), augment()
```

---

## Data Overview: Cheddar Taste Study

The `cheddar` dataset from the **{faraway}** package contains measurements from a study on cheddar cheese flavor, conducted in the LaTrobe Valley of Victoria, Australia. Researchers collected chemical composition and subjective taste scores across several cheese samples.

### Variables

- `Taste`: Subjective taste rating (response variable)
- `Acetic`: Log concentration of acetic acid
- `H2S`: Log concentration of hydrogen sulfide
- `Lactic`: Concentration of lactic acid

```{r}
head(cheddar)
dim(cheddar)
```

---

## Simple Linear Models

```{r}
FitA <- lm(Taste ~ Acetic, data = cheddar)
FitB <- lm(Taste ~ H2S, data = cheddar)
FitC <- lm(Taste ~ Lactic, data = cheddar)

summary(FitA)
```

```{r}
# Compare models via AIC and ANOVA
AIC(FitA)
anova(FitA)
```

---

## Multiple Linear Regression Models

```{r}
Fit1 <- lm(Taste ~ Acetic + H2S, data = cheddar)
Fit2 <- lm(Taste ~ Acetic + Lactic, data = cheddar)
Fit3 <- lm(Taste ~ H2S + Lactic, data = cheddar)
FitAll <- lm(Taste ~ Acetic + H2S + Lactic, data = cheddar)
```

---

## Model Selection: Stepwise Regression

```{r}
# Stepwise selection from full model
step_model <- stepAIC(FitAll, direction = "both", trace = FALSE)
summary(step_model)
```

---

## Model Performance Metrics

```{r}
# Evaluate performance on original data
rmse(step_model, cheddar)
mae(step_model, cheddar)
rsquare(step_model, cheddar)

# If you intended qae() for quantile error—consider implementing or clarifying this function
```

---

## Information Criteria

```{r}
AIC(step_model)
BIC(step_model)
```

---

## Model Tidying with {broom}

```{r}
# Coefficients and model summary
tidy(step_model)

# Augmented data (fitted values, residuals, etc.)
augment(step_model, cheddar)
```
