---
title: "Poisson Regression — Crabs Data Set"
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(glm2)
library(knitr)
library(magrittr)
```

## Overview

In this tutorial, we apply **Poisson regression** to explore the relationship between the number of satellites (male crabs near a female horseshoe crab) and various characteristics of the female crab, such as **carapace width**, **color**, and **spine condition**.

---

## Data Source

This dataset comes from a study on nesting horseshoe crabs by J. Brockmann (Ethology, 1996). Each observation corresponds to a female crab with associated explanatory variables and the response: number of satellite males.

```{r}
data(crabs)
head(crabs) %>% kable()
```

```{r}
summary(crabs[, 1:4]) %>% kable()
```

### Variables

- `Satellites`: Number of male crabs near the female (response)
- `Width`: Carapace width in centimeters
- `Dark`: Binary indicator for dark coloration
- `GoodSpine`: Binary indicator for good spine condition

---

## Poisson Regression: Satellites vs. Width

Let’s model the number of satellites using `Width` as the sole predictor:

```{r}
crabs.pois <- glm2(Satellites ~ Width, data = crabs, family = "poisson")
summary(crabs.pois)
```

### Interpretation

The fitted model is:

\[
\log(\mathbb{E}[\text{Satellites}]) = \beta_0 + \beta_1 \cdot \text{Width}
\]

So:

- \( e^{\beta_1} \) represents the **multiplicative change** in the expected number of satellites for a **1 cm increase** in width.

```{r}
exp(coef(crabs.pois)["Width"])
```

---

## Predictions

Let’s calculate expected number of satellites for specific carapace widths.

```{r}
# Reuse coefficients from model
b0 <- coef(crabs.pois)[1]
b1 <- coef(crabs.pois)[2]

# Prediction for width = 22 cm
exp(b0 + b1 * 22)

# Prediction for width = 25 cm
exp(b0 + b1 * 25)
```

---

## Visualizing Fit

```{r}
plot(crabs$Width, crabs$Satellites,
     pch = 16, col = "darkred",
     xlab = "Carapace Width (cm)",
     ylab = "Number of Satellites",
     main = "Observed vs. Predicted Values")

points(crabs$Width, fitted(crabs.pois),
       col = "darkgreen", pch = 19)
legend("topleft", legend = c("Observed", "Predicted"),
       pch = c(16, 19), col = c("darkred", "darkgreen"))
```

---

## Recap

- Poisson regression is ideal for modeling **count data** like the number of satellite males.
- **Exponentiating the slope** gives the multiplicative change per unit increase in the predictor.
- Predictions can be made directly from the model using `exp(b0 + b1 * x)`.

---
