---
title: "Quasi-Poisson Regression"
subtitle: Advanced Regression Models
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


### Quasi-Poisson Regression

The Quasi-Poisson Regression is a generalization of the Poisson regression and is used when modeling an overdispersed count variable.

The Poisson model assumes that the variance is equal to the mean, which is not always a fair assumption. When the variance is greater than the mean, a Quasi-Poisson model, which assumes that the variance is a linear function of the mean, is more appropriate.

```{r}
## quasipoisson. compare with example(glm)
counts <- c(18,17,15,20,10,20,25,13,12)
outcome <- gl(3,1,9)
treatment <- gl(3,3)
d.AD <- data.frame(treatment, outcome, counts)
```

```{r}
d.AD
```

```{r}
var(counts)/mean(counts)
```

----------------

#### Poisson Model

```{r}
glm.pois <- glm(counts ~ outcome + treatment, family = poisson())
summary(glm.pois)
```


----------------

#### Quasi-Poisson Model


```{r}
glm.qpois <- glm(counts ~ outcome + treatment, family = quasipoisson())
summary(glm.qpois)
```

-----------------------



### Anova


```{r}
anova(glm.qpois, test = "F")
```


------------


```{r}
## for Poisson results use
anova(glm.qpois, dispersion = 1, test = "Chisq")

```

-----------




```{r}
summary(glm.qpois, dispersion = 1)
```



