---
title: "MethComp Data Sets"
author: "KevinOBrien"
date: "18 September 2018"
output: html_document
---

## ``glucose`` Glucose measurements by different methods

### Description
* 74 persons in 5 centres in Finland had blood glucose measured by 11 different methods, based on 4 different types of blood.
* Each person had blood sampled at 0, 30, 60 and 120 min after a 75g glucose load.

### Format
A data frame with 1302 observations on the following 6 variables.

* ``meth`` Method of measurement. A factor with 11 levels: n.plas1 n.plas2 h.cap h.blood h.plas
h.serum m.plas m.serum o.cap s.serum k.plas.
* ``type`` Type of blood sample. A factor with 4 levels: blood plasma serum capil
* ``item`` Person id.
* ``time`` Time of blood sampling. Minutes since glucose load.
* ``cent`` Center of sampling. Except for the two first methods, n.plas1 and n.plas2, samples were
analyzed at the centres too
* ``y`` Glucose measurement in mmol/l.

### Load useful packages for analysis

```{r setup, include=FALSE}
library(MethComp)
library(nlme)
library(dplyr)
library(tidyr)
library(lme4)
library(forcats)
```
* Analysis under Bland and Altman
* Analysis under Roy's Method
 - Implementation of the four models
 - Expression of the Matrices
 - Implementation of the three hypothesis tests        
* Analysis under BXC
 - Computation of Limits of Agreement


```{r}
data(glucose)
head(glucose)
```

```{r,echo=FALSE}

data(glucose)
summary(glucose)
glucose$item <- factor(glucose$item)
table(glucose$time)
glucose <- glucose %>% mutate(repl = as.factor(time),
                              repl = fct_recode(repl,"1"="0","2"="30","3"="60","4"="120"))
```


```
glucose$repl <- factor(glucose$repl)


```

```{r}
dat <- glucose %>% filter(meth %in% c("n.plas1","n.plas2"))

glucose.roy1 = lme(y ~ meth-1, data = dat,
              random = list(item=pdSymm(~ meth-1)), 
              weights=varIdent(form=~1|meth),
              correlation = corSymm(form=~1 | item/repl), 
              method="ML")
glucose.roy2 = lme(y ~ meth-1, data = dat,
              random = list(item=pdCompSymm(~ meth-1)), 
              correlation = corSymm(form=~1 | item/repl), 
              method="ML")
glucose.roy3 = lme(y ~ meth-1, data = dat,
              random = list(item=pdSymm(~ meth-1)),
              weights=varIdent(form=~1|meth), 
              correlation = corCompSymm(form=~1 | item/repl), method="ML")
glucose.roy4 = lme(y ~ meth-1, data = dat,
              random = list(item=pdCompSymm(~ meth-1)), 
              correlation = corCompSymm(form=~1 | item/repl), 
              method="ML")
#getSigma(glucose.roy1)
#getOmega(glucose.roy1)
#roy.SDV(glucose.roy1)
```
## Analysis using BXC
```{r}
glucose.bxc1 = lme( y ~ meth + item, random = list( item = pdIdent( ~ meth-1 ) ), 
               weights = varIdent( form = ~1 | meth ), data=dat)
glucose.bxc2 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ), 
               weights = varIdent( form = ~1 | meth ), data=dat )
glucose.bxc3 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ),
               weights = varIdent( form = ~1 | meth ),
               correlation = corCompSymm(form=~1 | item/repl ),data=dat)

```