---
title: "MethComp Data Sets"
author: "KevinOBrien"
date: "18 September 2018"
output: html_document
---

```{r setup, include=FALSE}
library(MethComp)
library(nlme)
library(lme4)
library(magrittr)
library(tidyr)
library(broom)
library(dplyr)
```

## Fat 
This data set comprises fat content of human milk determined by measurement of glycerol released by enzymic hydrolysis of triglycerides (Trig) and measurement by the Standard Gerber method (Gerber). 
Units are (g/100 ml).

### Limits of Agreement
For the `Fat' data set, the inter-method bias is shown to be $0.045$. The limits of agreement are $(-0.23 , 0.32)$

For Carstensen's `fat' data, the limits of agreement computed using Roy's method are consistent with the estimates given by Carstensen et al (2008); $ = (-0.224,  0.314).$


A linear mixed effects model is formulated, and implementation through several software packages is demonstrated.
All of the necessary terms are presented in the computer output. The limits of agreement are therefore,
\begin{equation}
0.0449  \pm 1.96 \times  \sqrt{2 \times 0.0596^2 + 0.0772^2 + 0.0724^2} = (-0.220,  0.309).
\end{equation}

All of these terms are given or determinable in computer output. The limits of agreement can therefore be evaluated using
\begin{equation}
\bar{y_{A}}-\bar{y_{B}} \pm 1.96 \times \sqrt{ \sigma^2_{A} + \sigma^2_{B}  - 2(\sigma_{AB})}.
\end{equation}
```{r}
data(fat)
head(fat)
```

```{r}
data(fat)

fat <- fat %>% gather("Meth","y",Sub:Vic)
fat <- fat %>% set_colnames(c("item","obs","repl","meth","y"))
fat %>% filter(item == 1)
```

```{r}
# Analysis under Bland and Altman
# Analysis under Roy's Method
#       - Implementation of the four models
#       - Expression of the Matrices
#       - Implementation of the three hypothesis tests        
# Analysis under BXC
#       - Computation of Limits of Agreement
#################################################################
#
# Load useful packages for analysis

```


```{r}

dat=fat

fat.roy1 = lme(y ~ meth-1, data = dat,
              random = list(item=pdSymm(~ meth-1)), 
              weights=varIdent(form=~1|meth),
              correlation = corSymm(form=~1 | item/repl), 
              method="ML")
fat.roy2 = lme(y ~ meth-1, data = dat,
              random = list(item=pdCompSymm(~ meth-1)), 
              correlation = corSymm(form=~1 | item/repl), 
              method="ML")
fat.roy3 = lme(y ~ meth-1, data = dat,
              random = list(item=pdSymm(~ meth-1)),
              weights=varIdent(form=~1|meth), 
              correlation = corCompSymm(form=~1 | item/repl), method="ML")
fat.roy4 = lme(y ~ meth-1, data = dat,
              random = list(item=pdCompSymm(~ meth-1)), 
              correlation = corCompSymm(form=~1 | item/repl), 
              method="ML")
#getSigma(fat.roy1)
#getOmega(fat.roy1)
#roy.SDV(fat.roy1)
```
## Analysis using BXC
```{r}
fat.bxc1 = lme( y ~ meth + item, random = list( item = pdIdent( ~ meth-1 ) ), 
               weights = varIdent( form = ~1 | meth ), data=dat)
fat.bxc2 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ), 
               weights = varIdent( form = ~1 | meth ), data=dat )
fat.bxc3 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ),
               weights = varIdent( form = ~1 | meth ),
               correlation = corCompSymm(form=~1 | item/repl ),data=dat)

```

