---
title: "MethComp Data Sets"
author: "KevinOBrien"
date: "18 September 2018"
output: html_document
---

Oximetry Data
=============================
Carstensen *et al* (2008) introduces a second data set; the oximetry study. This study done at the Royal Children's Hospital in Melbourne to assess the agreement between co-oximetry and pulse oximetry in small babies.

In most cases, measurements were taken by both method at three different times. In some cases there are either one or two pairs of measurements, hence the data is unbalanced. 

Carstensen *et al* (2008) describes many of the children as being very sick, and with very low oxygen saturations levels. Therefore it must be assumed that a biological change can occur in interim periods, and measurements are not true replicates.

Carstensen *et al* (2008) demonstrate the necessity of accounting for linked replicated by comparing the limits of agreement from the `oximetry' data set using a model with the additional term, and one without. When the interaction is accounted for the limits of agreement are (-9.62,14.56). When the interaction is not accounted for, the limts of agreement are (-11.88,16.83). It is shown that the failure to include this additional term results in an over-estimation of the standard deviations of differences.

```{r setup, include=FALSE}
## Load useful packages for analysis
library(MethComp)
library(nlme)
library(lme4)
library(magrittr)
library(tidyr)
library(broom)
library(dplyr)
```

```{r,echo=FALSE}
data(ox)
summary(ox)
ox$item <- factor(ox$item)
ox$repl <- factor(ox$repl)
dat=ox
```


* Analysis under Bland and Altman
* Analysis under Roy's Method
 - Implementation of the four models
 - Expression of the Matrices
 - Implementation of the three hypothesis tests        
* Analysis under BXC
 - Computation of Limits of Agreement

```{r}
data(ox)
ox2 <- ox  %>% spread(repl,y) %>% set_colnames(c("meth","item","repl1","repl2","repl3")) 
ox2 <- data.frame(ox2,y.means = (ox2 %>% select(starts_with("repl")) %>% rowMeans(na.rm=TRUE) ))
var.test(y.means ~ meth,data=ox2)
t.test(y ~ meth,data=ox)
```

```{r}
ox.roy1 = lme(y ~ meth-1, data = dat,
              random = list(item=pdSymm(~ meth-1)), 
              weights=varIdent(form=~1|meth),
              correlation = corSymm(form=~1 | item/repl), 
              method="ML")
ox.roy2 = lme(y ~ meth-1, data = dat,
              random = list(item=pdCompSymm(~ meth-1)), 
              correlation = corSymm(form=~1 | item/repl), 
              method="ML")
ox.roy3 = lme(y ~ meth-1, data = dat,
              random = list(item=pdSymm(~ meth-1)),
              weights=varIdent(form=~1|meth), 
              correlation = corCompSymm(form=~1 | item/repl), method="ML")
ox.roy4 = lme(y ~ meth-1, data = dat,
              random = list(item=pdCompSymm(~ meth-1)), 
              correlation = corCompSymm(form=~1 | item/repl), 
              method="ML")
#getSigma(ox.roy1)
#getOmega(ox.roy1)
#roy.SDV(ox.roy1)
```
## Analysis using BXC
```{r}
ox.bxc1 = lme( y ~ meth + item, random = list( item = pdIdent( ~ meth-1 ) ), 
               weights = varIdent( form = ~1 | meth ), data=dat)
ox.bxc2 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ), 
               weights = varIdent( form = ~1 | meth ), data=dat )
ox.bxc3 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ),
               weights = varIdent( form = ~1 | meth ),
               correlation = corCompSymm(form=~1 | item/repl ),data=dat)

```