---
title: "MethComp Data Sets"
author: "KevinOBrien"
date: "20 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


``cardiac`` : Measurement of cardiac output by two different methods.

### Description
For each subject cardiac output is measured repeatedly (three to six times) by impedance cardiography
(IC) and radionuclide ventriculography (RV).

### Format
A data frame with 120 observations on the following 4 variables.
* ``meth`` a factor with levels IC RV
* ``item`` a numeric vector giving the item number.
* ``repl`` a numeric vector with replicate number.
* ``y`` the measuremnts of cardiac output.



```{r setup, include=FALSE}
## Load useful packages for analysis
library(MethComp)
library(nlme)
library(lme4)
library(magrittr)
library(tidyr)
library(broom)
library(dplyr)
```

```{r,echo=FALSE}
data(cardiac)
summary(cardiac)
cardiac$item <- factor(cardiac$item)
cardiac$repl <- factor(cardiac$repl)
dat=cardiac
```


* Analysis under Bland and Altman
* Analysis under Roy's Method
- Implementation of the four models
- Expression of the Matrices
- Implementation of the three hypothesis tests        
* Analysis under BXC
- Computation of Limits of Agreement

```{r}
data(cardiac)
cardiac2 <- cardiac  %>% spread(repl,y,fill=NA) %>% set_colnames(c("meth","item","repl1","repl2","repl3","repl4","repl5","repl6")) 
head(cardiac2)
```


```{r}
cardiac2 <- data.frame(cardiac2,y.means = (cardiac2 %>% select(starts_with("repl")) %>% rowMeans(na.rm=TRUE) ))
var.test(y.means ~ meth,data=cardiac2)
t.test(y ~ meth,data=cardiac)
```

```{r}
cardiac.roy1 <- lme(y ~ meth-1, data = dat,
  random = list(item=pdSymm(~ meth-1)), 
  weights=varIdent(form=~1|meth),
  correlation = corSymm(form=~1 | item/repl), 
  method="ML")
```
```{r}
cardiac.roy2 <- lme(y ~ meth-1, data = dat,
  random = list(item=pdCompSymm(~ meth-1)), 
  correlation = corSymm(form=~1 | item/repl), 
  method="ML")
```

```{r}
cardiac.roy3 <- lme(y ~ meth-1, data = dat,
  random = list(item=pdSymm(~ meth-1)),
  weights=varIdent(form=~1|meth), 
  correlation = corCompSymm(form=~1 | item/repl), method="ML")
```
```{r}
cardiac.roy4 <- lme(y ~ meth-1, data = dat,
  random = list(item=pdCompSymm(~ meth-1)), 
  correlation = corCompSymm(form=~1 | item/repl), 
  method="ML")

```
## Analysis using BXC
```{r}
cardiac.bxc1 = lme( y ~ meth + item, random = list( item = pdIdent( ~ meth-1 ) ), 
weights = varIdent( form = ~1 | meth ), data=dat)
cardiac.bxc2 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ), 
weights = varIdent( form = ~1 | meth ), data=dat )
cardiac.bxc3 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ),
weights = varIdent( form = ~1 | meth ),
correlation = corCompSymm(form=~1 | item/repl ),data=dat)

```

### Details
It is not entirely clear from the source whether the replicates are exchangeable within (method,item)
or whether they represent pairs of measurements. From the description it looks as if replicates are
linked between methods, but in the paper they are treated as if they were not.

### Source
The dataset is adapted from table 4 in: JM Bland and DG Altman: Measuring agreement in method
comparison studies. Statistical Methods in Medical Research, 8:136-160, 1999. Originally supplied
to Bland \& Altman by Dr LS Bowling, see: Bowling LS, Sageman WS, O'Connor SM, Cole
R, Amundson DE. Lack of agreement between measurement of ejection fraction by impedance
cardiography versus radionuclide ventriculography. Critical Care Medicine 1993; 21: 1523-27.
