---
title: "MethComp Data Sets"
author: "KevinOBrien"
date: "18 September 2018"
output: html_document
---

```{r setup, include=FALSE}
library(MethComp)
library(nlme)
library(lme4)
library(magrittr)
library(tidyr)
library(broom)
library(dplyr)
```


### PEFR and Cardiac
	
Two further data sets applied to both methodologies are the``Cardiac" and ``PEFR" , which are both contained on Carstensen's MethComp package. This data is from Bland and Altman (1986): two measurements of peak expiratory flow rate (PEFR) are compared. One of these measurements uses a ``Large" meter and the other a ``Mini" meter.
	
Two measurements were made with a Wright peak flow meter and two with a mini Wright meter, in random order.  All measurements were taken by the same observer, using the same two instruments. (These data were collected to demonstrate the statistical method and provide no evidence on the comparability of these two instruments.)
	


```{r}
data(PEFR)
head(PEFR)
```


```{r}
PEFR2 <- PEFR  %>% spread(repl,y) %>% set_colnames(c("meth","item","repl1","repl2")) 
PEFR2 <- data.frame(PEFR2,y.means = (PEFR2 %>% select(starts_with("repl")) %>% rowMeans(na.rm=TRUE) ))
```

```{r}

var.test(y.means ~ meth,data=PEFR2)
t.test(y ~ meth,data=PEFR)



```


```{r}

library(nlme)

WrightOne = c(494, 395, 516, 434, 476, 557, 413, 442, 650,  433, 417, 656, 267, 478, 178, 423, 427)
WrightTwo = c(490, 397, 512, 401, 470, 611, 415, 431, 638,  429, 420, 633, 275, 492, 165, 372, 421)
MiniOne = c(512, 430, 520, 428, 500, 600, 364, 380, 658,  445, 432, 626, 260, 477, 259, 350, 451)
MiniTwo = c(525, 415, 508, 444, 500, 625, 460, 390, 642,  432, 420, 605, 227, 467, 268, 370, 443)

PEFR=c(MiniOne,MiniTwo,WrightOne,WrightTwo)
Subject= rep(1:17,4)
repl= c(rep(1,17),rep(2,17),rep(1,17),rep(2,17))
meth=c(rep("M",34),rep("W",34))

Flow = groupedData( resp ~ method | subject ,
        data = data.frame( resp = PEFR, subject = Subject,
        method = meth,
        obs = repl),
    labels = list(resp = "PEFR", 
    method = "Measurement Device"),
    order.groups = FALSE )

H.o = lme(resp ~ method-1, data = Flow,
  random = list(subject=pdSymm(~ method-1)),
  weights=varIdent(form=~1|method), 
  correlation = corSymm(form=~1 | subject/obs), 
  method="ML")

######################################## Flow1
PEFR1=c(WrightOne,WrightTwo,MiniOne,MiniTwo)
Subject= rep(1:17,4)
repl= c(rep(1,17),rep(2,17),rep(1,17),rep(2,17))
meth1=c(rep("W",34),rep("M",34))

Flow1 = groupedData( resp ~ method | subject ,
        data = data.frame( resp = PEFR1, subject = Subject,
        method = meth1,
        obs = repl),
    labels = list(resp = "PEFR", method = "Measurement Device"),
    order.groups = FALSE )

H.oa = lme(resp ~ method-1, data = Flow1,
  random = list(subject=pdSymm(~ method-1)),
  weights=varIdent(form=~1|method), 
  correlation = corSymm(form=~1 | subject/obs), 
  method="ML")
######################################## Flow1(ordered)
Flow2=orderdata(Flow1)

H.ob = lme(resp ~ method-1, data = Flow2,
  random = list(subject=pdSymm(~ method-1)),
  weights=varIdent(form=~1|method), 
  correlation = corSymm(form=~1 | subject/obs), 
  method="ML")

########################################
getD(H.o)
getD(H.oa)
getD(H.ob) # accords with H.o

getSigma(H.o)
getSigma(H.oa)
getSigma(H.ob)
########################################

```