---
title: "MethComp Data Sets"
author: "KevinOBrien"
date: "18 September 2018"
output: html_document
---

Enzyme: Enzyme activity data
================================

### Description
* Three measurement of enzyme activity on 24 patients. 
* The measurements is of the enzymes sucrase and alkaline phosphatase. 
* The interest is to compare the 'homogenate' and 'pellet' methods.

Format
A data frame with 72 observations on the following 3 variables.
* ``meth`` a factor with levels SucHom SucPel Alkphos, representing three different measurements, i.e.
homogenate and pellet values of sucrase, as well as homogenate values of alkaline.
* ``item`` a numeric vector, the person ID for the 24 patients
* ``y`` a numeric vector, the measurements on the enzyme activity.



```{r setup, include=FALSE}
library(MethComp)
library(nlme)
library(lme4)
library(magrittr)
library(tidyr)
library(broom)
library(dplyr)
```
```{r}
data(Enzyme)
Enzyme %>% filter(item == 1)
```

```{r}
Enzyme <- Meth( Enzyme )
summary( Enzyme )

```



```{r}

summary( Enzyme )
plot( Enzyme )
```




```{r}

dat=Enzyme

Enzyme.roy1 = lme(y ~ meth-1, data = dat,
              random = list(item=pdSymm(~ meth-1)), 
              weights=varIdent(form=~1|meth),
              correlation = corSymm(form=~1 | item/repl), 
              method="ML")
Enzyme.roy2 = lme(y ~ meth-1, data = dat,
              random = list(item=pdCompSymm(~ meth-1)), 
              correlation = corSymm(form=~1 | item/repl), 
              method="ML")
```
```{r}
Enzyme.roy3 = lme(y ~ meth-1, data = dat,
              random = list(item=pdSymm(~ meth-1)),
              weights=varIdent(form=~1|meth), 
              correlation = corCompSymm(form=~1 | item/repl), method="ML")
Enzyme.roy4 = lme(y ~ meth-1, data = dat,
              random = list(item=pdCompSymm(~ meth-1)), 
              correlation = corCompSymm(form=~1 | item/repl), 
              method="ML")

```
## Analysis using BXC
```{r}
Enzyme.bxc1 = lme( y ~ meth + item, random = list( item = pdIdent( ~ meth-1 ) ), 
               weights = varIdent( form = ~1 | meth ), data=dat)
Enzyme.bxc2 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ), 
               weights = varIdent( form = ~1 | meth ), data=dat )
Enzyme.bxc3 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ),
               weights = varIdent( form = ~1 | meth ),
               correlation = corCompSymm(form=~1 | item/repl ),data=dat)

```

```{r}
 tidy(Enzyme.roy) # conditional Residuals 
```



### Source
R. L. Carter; Restricted Maximum Likelihood Estimation of Bias and Reliability in the Comparison
of Several Measuring Methods; Biometrics, Dec., 1981, Vol. 37, No. 4, pp. 733-741.