---
title: "MethComp Data Sets"
author: "KevinOBrien"
date: "18 September 2018"
output: html_document
---

```{r setup, include=FALSE}
library(MethComp)
library(nlme)
library(lme4)
library(magrittr)
library(tidyr)
library(broom)
library(dplyr)
```
VitCap Merits of two instruments designed to measure certain aspects of human
lung function (Vital Capacity)
  Description
Measurement on certain aspects of human lung capacity for 72 patients on 4 instrument-operative
combination, i.e. two different instruments and two different users, a skilled one and a new one.

### Format
A data frame with 288 observations on the following 5 variables.

* meth a factor with levels StNew, StSkil, ExpNew and ExpSkil, representing the instrument by user
combinations. See below.
* item a numeric vector, the person ID, i.e. the 72 patients
* y a numeric vector, the measurements, i.e. vital capacity.
* user a factor with levels New Skil, for the new user and the skilled user
* instrument a factor with levels Exp and St, for the experimental instrument and the standard one.

Examples
```{r}
data(VitCap)
Vcap <- Meth( VitCap )
str( Vcap )
plot( Vcap )
```

```{r}
data(VitCap)

VitCap <- VitCap %>% gather("Meth","y",Sub:Vic)
VitCap <- VitCap %>% set_colnames(c("item","obs","repl","meth","y"))
VitCap %>% filter(item == 1)
```

* Analysis under Bland and Altman
* Analysis under Roy's Method
 - Implementation of the four models
 - Expression of the Matrices
 - Implementation of the three hypothesis tests        
* Analysis under BXC
 - Computation of Limits of Agreement



```{r}

dat=VitCap

VitCap.roy1 = lme(y ~ meth-1, data = dat,
              random = list(item=pdSymm(~ meth-1)), 
              weights=varIdent(form=~1|meth),
              correlation = corSymm(form=~1 | item/repl), 
              method="ML")
VitCap.roy2 = lme(y ~ meth-1, data = dat,
              random = list(item=pdCompSymm(~ meth-1)), 
              correlation = corSymm(form=~1 | item/repl), 
              method="ML")
```
```{r}

VitCap.roy3 = lme(y ~ meth-1, data = dat,
              random = list(item=pdSymm(~ meth-1)),
              weights=varIdent(form=~1|meth), 
              correlation = corCompSymm(form=~1 | item/repl), method="ML")
VitCap.roy4 = lme(y ~ meth-1, data = dat,
              random = list(item=pdCompSymm(~ meth-1)), 
              correlation = corCompSymm(form=~1 | item/repl), 
              method="ML")
#getSigma(VitCap.roy1)
#getOmega(VitCap.roy1)
#roy.SDV(VitCap.roy1)
```
## Analysis using BXC
```{r}
VitCap.bxc1 = lme( y ~ meth + item, random = list( item = pdIdent( ~ meth-1 ) ), 
               weights = varIdent( form = ~1 | meth ), data=dat)
VitCap.bxc2 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ), 
               weights = varIdent( form = ~1 | meth ), data=dat )
VitCap.bxc3 = lme( y ~ meth + item, random=list( item = pdIdent( ~ meth-1 ), repl = ~1 ),
               weights = varIdent( form = ~1 | meth ),
               correlation = corCompSymm(form=~1 | item/repl ),data=dat)

```

```{r}
 tidy(VitCap.roy) # conditional Residuals 
```

```{r}
getLOA(VitCap.roy) 
```
### Source
V. D. Barnett, Simultaneous Pairwise Linear Structural Relationships, Biometrics, Mar. 1969, Vol.
25, No. 1, pp. 129-142.