---
title: "MethComp Data Sets"
author: "KevinOBrien"
date: "18 September 2018"
output: html_document
---

```{r setup, include=FALSE}
library(MethComp)
library(nlme)
library(lme4)
library(magrittr)
library(tidyr)
library(broom)
library(dplyr)
```

```{r}
data(sbp)
```





sbp Systolic blood pressure measured by three different methods.
Description
For each subject (item) there are three replicate measurements by three methods (two observers, J
                                                                                 and R and the automatic machine, S). The replicates are linked within (method,item).
Usage
data(sbp)
Format
A data frame with 765 observations on the following 4 variables:
  meth Methods, a factor with levels J(observer 1), R(observer 2) and S(machine)
item Person id, numeric.
repl Replicate number, a numeric vector
y Systolic blood pressure masurement, a numeric vector
Source
The dataset is adapted from table 1 in: JM Bland and DG Altman: Measuring agreement in method
comparison studies. Statistical Methods in Medical Research, 8:136-160, 1999. Originally supplied
to Bland \& Altman by E. O'Brien, see: Altman DG, Bland JM. The analysis of blood pressure data.
In O'Brien E, O'Malley K eds. Blood pressure measurement. Amsterdam: Elsevier, 1991: 287-314.