---
title: "DRR"
author: "KevinO'Brien"
date: "22 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DRR)
```

## Dimensionality Reduction via Regression

#### Description

``drr`` Implements Dimensionality Reduction via Regression using Kernel Ridge Regression.

####  Usage
<pre><code>
drr(X, ndim = ncol(X), lambda = c(0, 10^(-3:2)), kernel = "rbfdot",
  kernel.pars = list(sigma = 10^(-3:4)), pca = TRUE, pca.center = TRUE,
  pca.scale = FALSE, fastcv = FALSE, cv.folds = 5, fastcv.test = NULL,
  fastkrr.nblocks = 4, verbose = TRUE)
</code></pre>

#### Arguments

* ``X``	
input data, a matrix.
* ``ndim``	
the number of output dimensions and regression functions to be estimated, see details for inversion.
* ``lambda``	
the penalty term for the Kernel Ridge Regression.
* ``kernel``	
a kernel function or string, see kernel-class for details.
* ``kernel.pars``	
a list with parameters for the kernel. each parameter can be a vector, crossvalidation will choose the best combination.
* ``pca``	
logical, do a preprocessing using pca.
* ``pca.center``	
logical, center data before applying pca.
* ``pca.scale``	
logical, scale data before applying pca.
* ``fastcv``	
if TRUE uses fastCV, if FALSE uses CV for crossvalidation.
* ``cv.folds``	
if using normal crossvalidation, the number of folds to be used.
* ``fastcv.test``	
an optional separate test data set to be used for fastCV, handed over as option test to fastCV.
* ``fastkrr.nblocks``	
the number of blocks used for fast KRR, higher numbers are faster to compute but may introduce numerical inaccurracies, see constructFastKRRLearner for details.
* ``verbose``	
logical, should the crossvalidation report back.

#### Details

Parameter combination will be formed and cross-validation used to select the best combination. Cross-validation uses CV or fastCV.

Pre-treatment of the data using a PCA and scaling is made ?? = Vx. the representation in reduced dimensions is

\[y_i = \alpha_ - f_i(\alpha__1, ., \alpha__{i-1})\]

then the final DRR representation is:

\[r = (\alpha__1, y_2, y_3, .,y_d)\]

DRR is invertible by

\[\alpha__i = y_i + f_i(\alpha__1,\alpha__2, ., \alpha_{i-1})\]

If less dimensions are estimated, there will be less inverse functions and calculating the inverse will be inaccurate.

#### Value

A list the following items:

* "fitted.data" The data in reduced dimensions.

* "pca.means" The means used to center the original data.

* "pca.scale" The standard deviations used to scale the original data.

* "pca.rotation" The rotation matrix of the PCA.

* "models" A list of models used to estimate each dimension.

* "apply" A function to fit new data to the estimated model.

* "inverse" A function to untransform data.

#### References

Laparra, V., Malo, J., Camps-Valls, G., 2015. Dimensionality Reduction via Regression in Hyperspectral Imagery. IEEE Journal of Selected Topics in Signal Processing 9, 1026-1036. doi:10.1109/JSTSP.2015.2417833

### Examples
```{r}
tt <- seq(0,4*pi, length.out = 200)
helix <- cbind(
  x = 3 * cos(tt) + rnorm(length(tt), sd = seq(0.1, 1.4, length.out = length(tt))),
  y = 3 * sin(tt) + rnorm(length(tt), sd = seq(0.1, 1.4, length.out = length(tt))),
  z= 2 * tt      + rnorm(length(tt), sd = seq(0.1, 1.4, length.out = length(tt)))
)

```
```{r}
helix <- helix[sample(nrow(helix)),] # shuffling data is important!!
system.time(
drr.fit  <- drr(helix, ndim = 3, cv.folds = 4,
                lambda = 10^(-2:1),
                kernel.pars = list(sigma = 10^(0:3)),
                fastkrr.nblocks = 2, verbose = TRUE,
                fastcv = FALSE)
)

```
```{r}
## Not run: 
library(rgl)
plot3d(helix)
points3d(drr.fit$inverse(drr.fit$fitted.data[,1,drop = FALSE]), col = 'blue')
points3d(drr.fit$inverse(drr.fit$fitted.data[,1:2]),             col = 'red')

plot3d(drr.fit$fitted.data)
pad <- -3
fd <- drr.fit$fitted.data
xx <- seq(min(fd[,1]),       max(fd[,1]),       length.out = 25)
yy <- seq(min(fd[,2]) - pad, max(fd[,2]) + pad, length.out = 5)
zz <- seq(min(fd[,3]) - pad, max(fd[,3]) + pad, length.out = 5)
```

```{r}
dd <- as.matrix(expand.grid(xx, yy, zz))
plot3d(helix)
for(y in yy) for(x in xx)
  rgl.linestrips(drr.fit$inverse(cbind(x, y, zz)), col = 'blue')
for(y in yy) for(z in zz)
  rgl.linestrips(drr.fit$inverse(cbind(xx, y, z)), col = 'blue')
for(x in xx) for(z in zz)
  rgl.linestrips(drr.fit$inverse(cbind(x, yy, z)), col = 'blue')
```
