`geom_jitter`
========================================================
This geom is for points, jittered to reduce overplotting.

### un-jittered plot
```{r fig.width=10, fig.height=4}
library(ggplot2)
p <- ggplot(mpg, aes(displ, hwy))
p + geom_point()
```
#### Jittering added to plot
```{r fig.width=10, fig.height=4}

p + geom_point(position = "jitter")
```
### Add aesthetic mappings
```{r fig.width=10, fig.height=4}

p + geom_jitter(aes(colour = cyl))
```
### Varying parameters
```{r fig.width=10, fig.height=4}

p + geom_jitter(position = position_jitter(width = .5))
p + geom_jitter(position = position_jitter(height = .5))
```
### Using `qplot`
```{r fig.width=10, fig.height=4}

qplot(displ, hwy, data = mpg, geom = "jitter")
qplot(class, hwy, data = mpg, geom = "jitter")
qplot(class, hwy, data = mpg, geom = c("boxplot", "jitter"))
qplot(class, hwy, data = mpg, geom = c("jitter", "boxplot"))

```
###  `position_jitter` 
This is used to adjust amount of jittering.
```{r fig.width=5, fig.height=3}
table(mtcars$am, mtcars$vs)
qplot(am, vs, data = mtcars) #not informative
```
Default amount of jittering will generally be too much for
small datasets:
```{r fig.width=10, fig.height=4}
qplot(am, vs, data = mtcars, position = "jitter")
```
Control the amount as follows
```{r fig.width=10, fig.height=4}
qplot(am, vs, data = mtcars, position = position_jitter(w = 0.1, h = 0.1))
# Much Better
```

The default works better for large datasets, where it will
take up as much space as a boxplot or a bar
```{r fig.width=10, fig.height=4}
qplot(class, hwy, data = mpg, geom = c("boxplot", "jitter"))
```

### `position_jitter`with ggplot

```{r fig.width=10, fig.height=5}

ggplot(mtcars, aes(x = am, y = vs)) + geom_point(position = "jitter")
ggplot(mtcars, aes(x = am, y = vs)) + geom_point(position = position_jitter(w = 0.1, h = 0.1))
```


