---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---



- Graphs outcomevspredictor variables holding the rest constant (also called partial-regression plots)


```{r}

library(car)
reg1 <-lm(prestige ~ education + income + type, data = Prestige)
avPlots(reg1, id.n=2, id.cex=0.7)

# Help identify the effect(or influence) of an observation on the regression coefficient of the predictor variable
```
```{r}
library(car)
reg1 <-lm(prestige ~ education + income + type, data = Prestige)
influenceIndexPlot(reg1, id.n=3)
```

- Cook's distance measures how much an observation influences the overall model or predicted values

- Studentizidedresiduals are the residuals divided by their estimated standard deviation as a way to standardized

- Bonferronitest to identify outliers

- Hat-points identify influential observations (have a high impact on the predictor variables)

NOTE: If an observation is an outlier and influential (high leverage) then that observation can change the fit of the linear model, it is advisable to remove it. To remove a case(s) type

```{r}
# reg1a <-update(prestige.reg4, subset=rownames(Prestige) != "general.managers")
#reg1b <-update(prestige.reg4, subset= !(rownames(Prestige) %in% c("general.managers","medical.technicians")))
```
```{r}
library(car)
reg1 <-lm(prestige ~ education + income + type, data = Prestige)
influencePlot(reg1, id.n=3)
# Creates a bubble-plot combining the display of Studentizedresiduals, hat-values, and Cook's distance (represented in the circles).
```
```{r}

library(car)
reg1 <-lm(prestige ~ education + income + type, data = Prestige)
qqPlot(reg1)
# Look for the tails, points should be close to the line or within the confidence intervals.
```
