---
title: The {car} R package - The Prestige Data set
subtitle: Appraising and Validating Linear Models
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r echo=FALSE,warning=FALSE,message=FALSE}
library(car)
reg1 <-lm(prestige ~ education + income + type, data = Prestige)
```


```{r,warning=FALSE,message=FALSE}
avPlots(reg1)
```

*  id.n–id most influential observation
* id.cex –font size for id.
* Graphs outcomevspredictor variables holding the rest constant (also called partial-regression plots)
* Help identify the effect(or influence) of an observation on the regression coefficient of the predictor variable

---------------------------------------------------------

```{r,warning=FALSE,message=FALSE}
influenceIndexPlot(reg1)
```

* Cook's distance measures how much an observation influences the overall model or predicted values
* Studentizidedresiduals are the residuals divided by their estimated standard deviation as a way to standardized
* Bonferronitest to identify outliers
*  Hat-points identify influential observations (have a high impact on the predictor variables)
NOTE: If an observation is an outlier and influential (high leverage) then that observation can change the fit of the linear model, it is advisable to remove it. To remove a case(s) type


```{r}
# reg1a <-update(prestige.reg4, subset=rownames(Prestige) != "general.managers")
# reg1b <-update(prestige.reg4, subset= !(rownames(Prestige) %in% c("general.managers","medical.technicians")))
```

---------------------------------------------------------


```{r}
influencePlot(reg1)
```

* Creates a bubble-plot combining the display of Studentizedresiduals, hat-values, and Cook's distance (represented in the circles).

---------------------------------------------------------

```{r}
qqPlot(reg1)
```

Look for the tails, points should be close to the line or within the confidence intervals.

---------------------------------------------------------
