
### The stackloss data set
%Used in "rlm" help file
Brownlee's Stack Loss Plant Data contains operational data of a plant for the oxidation of ammonia to nitric acid.

The variables are: 

* 	Air.Flow	 Flow of cooling air
* 	Water.Temp	 Cooling Water Inlet Temperature
* 	Acid.Conc.	 Concentration of acid [per 1000, minus 500]
* 	stack.loss	 Stack loss


```{r}
fit.sl = lm(stack.loss ~ ., data = stackloss)

#attach(stackloss)
#plot(Acid.Conc. , stack.loss, pch = 18, col="red")

par(mfrow=c(2,2))
plot(fit.sl)
par(mfrow=c(1,1))
```

-----------------------------------

### Fitting a robust model (<tt>rlm()</tt>}

The rlm command in the MASS package command implements several versions of robust regression. 

```{r}
library(MASS)

summary(rlm(stack.loss ~ ., data = stackloss))
```




-------------------------------------

### Using Other <i>Psi</i> Operators


*  huber
*  bisquare
*  hampel

Fitting is done by ***iterated re-weighted least squares (IWLS).***

Psi functions are supplied for the Huber, Hampel and Tukey bisquare proposals as <tt>psi.huber</tt>, <tt>psi.hampel</tt> and <tt>psi.bisquare</tt>. Huber's corresponds to a convex optimization problem and gives a unique solution (up to collinearity). The other two will have multiple local minima, and a good starting point is desirable.



-----------------------

### Huber


```{r}
summary(rlm(stack.loss ~ ., data = stackloss))
```

```{r}
summary(rlm(stack.loss ~ ., data = stackloss,  psi = psi.huber))
```


-------------------------------------

### Bisquare


```{r}
 rlm(stack.loss ~ ., stackloss, psi = psi.bisquare)
```


-------------------------------------

### Hampel


```{r}
 rlm(stack.loss ~ ., stackloss, psi = psi.hampel, init = "lts")
```




