---
title: "Poisson Regression - Crabs Data Set"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(magrittr)
```

### Prussian Cavalry Example


In 1898, a Russian economist and statistician, Ladislaus Josephovich Bortkiewicz, published an interesting findings about the probability distribution of Prussian soldiers accidentally killed by horse-kick. The data was derived from ten army corps who were observed over 20 years. There were a total of 200 observations and 122 soldiers were killed by horse kick over that 20 years. In average, the number of death is

$$ \lambda = \frac{122}{200} = 0.61$$
By using lambda value 0.61, Bortkiewicz applied Poisson formula to predict the probability of number of death, $x$, with x = 0, 1, 2, 3, 4, 5, 6:

```{r}
dpois(0:6,lambda=0.61) %>% round(4)
```

### Simulation Exercise

```{r}
set.seed(12345)
Cavalry <- rpois(200,lambda=0.61)
```


```{r}
Cavalry
```


### Mean and Variance 

*  Should be numerically close.
*  With a large sample, the convergence would be closer.

```{r}
mean(Cavalry);var(Cavalry)
```



### Dispersion Parameter


$$ \theta = \frac{Var(X)}{E(X)} \approx 1$$


```{r}
Theta <-  var(Cavalry)/mean(Cavalry)

Theta %>% round(3)
```


### Simulation Exercise


```{r}
N <-  100000
Theta <- numeric(N)

for ( i in 1:N){
   X=rpois(200,lambda=5);
   Theta[i] = var(X)/mean(X)
}
```

### 95% Tolerance Interval of Theta Values

```{r}
hist(Theta, breaks = 100, col=c("lightblue","lightpink"))
```

```{r}
quantile(Theta, c(0.025,0.975))  %>% round(3)
```



