---
title: "Poisson Regression - Crabs Data Set"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(magrittr)
```

The crabs data set
============================

The crabs data set is derived from Agresti (2007, Table 3.2, pp.76-77). It
gives 4 variables for each of 173 female horseshoe crabs.
    
* ***Satellites*** number of male partners in addition to the female's primary
partner
* ***Width*** width of the female in centimeters
* ***Dark*** a binary factor indicating whether the female has dark coloring
(yes or no)
* ***GoodSpine*** a binary factor indicating whether the female has good
spine condition (yes or no)

```{r warning=FALSE,message=FALSE}
library(glm2)

data(crabs)
head(crabs) %>% kable()
```

```{r}
summary(crabs[,1:4])  %>% kable()
```

#### Question 

Fit a Poisson regression model with the number of Satellites as the outcome
and the width of the female as the covariate. 

What is the multiplicative
change in the expected number of crabs for each additional centimeter of
width?



```{r}
crabs.pois <- glm2( Satellites ~ Width, data=crabs, family="poisson")

summary(crabs.pois)
```


```{r}
exp(-3.30476)*exp(0.164*25)

```




```{r}

plot(crabs$Width, crabs$Satellites,
     pch=16, col="darkred")
points(crabs$Width, crabs.pois$fitted.values, 
     col="green", type="p", lwd=3)

```

### Question
What is the expected number of Satellites for a female of width 22cm?

Given a set of parameters $\{\beta_0,\beta_1, \ldots, \beta_n\}$ and an input vector $x$ (i.e $\{x_1,x_2, \ldots x_n\}$), the mean of the predicted Poisson distribution is given by
$$\operatorname{E}(Y|x)=e^{\beta_0+\beta_1x_1 + \ldots + \beta_nx_n}\,
$$
In the case of one predictor variable, we can say
$$\operatorname{E}(Y|x)=e^{\beta_0+\beta_1x}=e^{\beta_0}\times e^{\beta_1x}\,$$

***Remark***: The expected value does not have to be an integer.





```{r}
exp(-3.30476)*exp(0.16405*22)
```
