
Assumption of Constant Variance
============================================

#### Homoscedasticity

* ***Homoscedascity*** is the technical term to describe the variance of the residuals being constant across the range of predicted values. 

* ***Heteroscedascity*** is the converse scenario : the variance differs along the range of values.



Suppose you plot the individual residuals against the predicted value, the variance of the residuals predicted value should be constant (Diagnostic Plot 1). 



```{r include=FALSE,echo=FALSE}

fit <- Fit_4
```


Using R
=====================================
```{r}	

# Evaluate homoscedasticity

# non-constant error variance test

library(car)
ncvTest(fit)
```

component+residual plots
==================================

* component+residual plots, also called partial-residual plots, for linear and generalized linear models.
* a graphical technique that attempts to show the relationship between a given independent variable and the response variable given that other independent variables are also in the model.
* Ceres plots are a generalization of component+residual (partial residual) plots that are less prone to leakage of nonlinearity among the predictors. 


Nonlinearity
=============================================================

#### component + residual plot 

```{r out.width = "80%" }	
	

# Evaluate Nonlinearity

# component + residual plot 


crPlots(fit)
```


Nonlinearity
=============================================================

#### ceres plot

```{r out.width = "80%" }
# Ceres plots 

ceresPlots(fit)
```

Nonlinearity
=============================================================


Test for Autocorrelated Errors

```{r}
# Test for Autocorrelated Errors

durbinWatsonTest(fit)

```


avplot & av plots
==================================
* The avplot command is short for added-variable plot and draws exactly that. 
* The other name for added variable plot is a partial-regression leverage plot. 
