---
title: "``Anova`` Anova Tables for Various Statistical Models"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```

Anova Anova Tables for Various Statistical Models

####Description

Calculates type-II or type-III analysis-of-variance tables for model objects produced by lm, glm,
multinom (in the nnet package), polr (in the MASS package), coxph (in the survival package),
coxme (in the coxme pckage), svyglm (in the survey package), rlm (in the MASS package), lmer
in the lme4 package, lme in the nlme package, and (by the default method) for most models with
a linear predictor and asymptotically normal coefficients (see details below). For linear models,
4 Anova
F-tests are calculated; for generalized linear models, likelihood-ratio chisquare, Wald chisquare,
or F-tests are calculated; for multinomial logit and proportional-odds logit models, likelihood-ratio
tests are calculated. Various test statistics are provided for multivariate linear models produced
by lm or manova. Partial-likelihood-ratio tests or Wald tests are provided for Cox models. Wald
chi-square tests are provided for fixed effects in linear and generalized linear mixed-effects models.
Wald chi-square or F tests are provided in the default case.

####Usage
<pre><code>

Anova(mod, ...)
Manova(mod, ...)
## S3 method for class 'lm'
Anova(mod, error, type=c("II","III", 2, 3),
white.adjust=c(FALSE, TRUE, "hc3", "hc0", "hc1", "hc2", "hc4"),
vcov.=NULL, singular.ok, ...)
## S3 method for class 'aov'
Anova(mod, ...)
## S3 method for class 'glm'
Anova(mod, type=c("II","III", 2, 3),
test.statistic=c("LR", "Wald", "F"),
error, error.estimate=c("pearson", "dispersion", "deviance"),
singular.ok, ...)
## S3 method for class 'multinom'
Anova(mod, type = c("II","III", 2, 3), ...)
## S3 method for class 'polr'
Anova(mod, type = c("II","III", 2, 3), ...)
## S3 method for class 'mlm'
Anova(mod, type=c("II","III", 2, 3), SSPE, error.df,
idata, idesign, icontrasts=c("contr.sum", "contr.poly"), imatrix,
test.statistic=c("Pillai", "Wilks", "Hotelling-Lawley", "Roy"),...)
## S3 method for class 'manova'
Anova(mod, ...)
## S3 method for class 'mlm'
Manova(mod, ...)
## S3 method for class 'Anova.mlm'
print(x, ...)
## S3 method for class 'Anova.mlm'
summary(object, test.statistic, univariate=object$repeated,
Anova 5
multivariate=TRUE, p.adjust.method, ...)
## S3 method for class 'summary.Anova.mlm'
print(x, digits = getOption("digits"),
SSP=TRUE, SSPE=SSP, ... )
## S3 method for class 'univaov'
print(x, digits = max(getOption("digits") - 2L, 3L),
style=c("wide", "long"),
by=c("response", "term"),
...)
## S3 method for class 'univaov'
as.data.frame(x, row.names, optional, by=c("response", "term"), ...)
## S3 method for class 'coxph'
Anova(mod, type=c("II", "III", 2, 3),
test.statistic=c("LR", "Wald"), ...)
## S3 method for class 'coxme'
Anova(mod, type=c("II", "III", 2, 3),
test.statistic=c("Wald", "LR"), ...)
## S3 method for class 'lme'
Anova(mod, type=c("II","III", 2, 3),
vcov.=vcov(mod, complete=FALSE), singular.ok, ...)
## S3 method for class 'mer'
Anova(mod, type=c("II", "III", 2, 3),
test.statistic=c("Chisq", "F"), vcov.=vcov(mod, complete=FALSE), singular.ok, ...)
## S3 method for class 'merMod'
Anova(mod, type=c("II", "III", 2, 3),
test.statistic=c("Chisq", "F"), vcov.=vcov(mod, complete=FALSE), singular.ok, ...)
## S3 method for class 'svyglm'
Anova(mod, ...)
## S3 method for class 'rlm'
Anova(mod, ...)
## Default S3 method:
Anova(mod, type=c("II", "III", 2, 3),
test.statistic=c("Chisq", "F"), vcov.=vcov(mod, complete=FALSE),
singular.ok, ...)

</code></pre>

####Arguments

mod lm, aov, glm, multinom, polr mlm, coxph, coxme, lme, mer, merMod, svyglm,
6 Anova
rlm, or other suitable model object.
error for a linear model, an lm model object from which the error sum of squares
and degrees of freedom are to be calculated. For F-tests for a generalized linear
model, a glm object from which the dispersion is to be estimated. If not
specified, mod is used.
type type of test, "II", "III", 2, or 3. Roman numerals are equivalent to the corresponding
Arabic numerals.
singular.ok defaults to TRUE for type-II tests, and FALSE for type-III tests where the tests for
models with aliased coefficients will not be straightforwardly interpretable; if
FALSE, a model with aliased coefficients produces an error.
test.statistic for a generalized linear model, whether to calculate "LR" (likelihood-ratio),
"Wald", or "F" tests; for a Cox or Cox mixed-effects model, whether to calculate
"LR" (partial-likelihood ratio) or "Wald" tests; in the default case or for linear
mixed models fit by lmer, whether to calculate Wald "Chisq" or Kenward-
Roger "F" tests with Satterthwaite degrees of freedom (warning: the KR F-tests
can be very time-consuming). For a multivariate linear model, the multivariate
test statistic to compute - one of "Pillai", "Wilks", "Hotelling-Lawley",
or "Roy", with "Pillai" as the default. The summary method for Anova.mlm
objects permits the specification of more than one multivariate test statistic, and
the default is to report all four.
error.estimate for F-tests for a generalized linear model, base the dispersion estimate on the
Pearson residuals ("pearson", the default); use the dispersion estimate in the
model object ("dispersion"); or base the dispersion estimate on the residual
deviance ("deviance"). For binomial or Poisson GLMs, where the dispersion
is fixed to 1, setting error.estimate="dispersion" is changed to "pearson",
with a warning.
white.adjust if not FALSE, the default, tests use a heteroscedasticity-corrected coefficient covariance
matrix; the various values of the argument specify different corrections.
See the documentation for hccm for details. If white.adjust=TRUE then the
"hc3" correction is selected.
SSPE For Anova for a multivariate linear model, the error sum-of-squares-and-products
matrix; if missing, will be computed from the residuals of the model; for the
print method for the summary of an Anova of a multivariate linear model,
whether or not to print the error SSP matrix (defaults to TRUE).
SSP if TRUE (the default), print the sum-of-squares and cross-products matrix for the
hypothesis and the response-transformation matrix.
error.df The degrees of freedom for error; if missing, will be taken from the model.
i* ``data  `` : an optional data frame giving a factor or factors defining the intra-subject model
for multivariate repeated-measures data. See details
 for an explanation of the
intra-subject design and for further explanation of the other arguments relating
to intra-subject factors.
idesign a one-sided model formula using the "data" in idata and specifying the intrasubject
design.
icontrasts names of contrast-generating functions to be applied by default to factors and
ordered factors, respectively, in the within-subject "data"; the contrasts must
Anova 7
produce an intra-subject model matrix in which different terms are orthogonal.
The default is c("contr.sum", "contr.poly").
imatrix as an alternative to specifying idata, idesign, and (optionally) icontrasts,
the model matrix for the within-subject design can be given directly in the form
of list of named elements. Each element gives the columns of the within-subject
model matrix for a term to be tested, and must have as many rows as there are
responses; the columns of the within-subject model matrix for different terms
must be mutually orthogonal.
x, object object of class "Anova.mlm" to print or summarize.
multivariate, univariate
compute and print multivariate and univariate tests for a repeated-measures ANOVA
or multivariate linear model; the default is TRUE for both for repeated measures
and TRUE for multivariate for a multivariate linear model.
p.adjust.method
if given for a multivariate linear model when univariate tests are requested, the
univariate tests are corrected for simultaneous inference by term; if specified,
should be one of the methods recognized by p.adjust or TRUE, in which case
the default (Holm) adjustment is used.
digits minimum number of significant digits to print.
style for printing univariate tests if requested for a multivariate linear model; one of
"wide", the default, or "long".
by if univariate tests are printed in "long" style, they can be ordered by "response",
the default, or by "term".
row.names, optional
not used.
* `` vcov.``:  in the default method, an optional coefficient-covariance matrix or function
to compute a covariance matrix, computed by default by applying the generic
vcov function to the model object. A similar argument may be supplied to the
lm method, and the default (NULL) is to ignore the argument; if both vcov. and
white.adjust are supplied to the lm method, the latter is used.
... do not use.



####Examples

```{r}

## Two-Way Anova
mod <- lm(conformity ~ fcategory*partner.status, data=Moore,
contrasts=list(fcategory=contr.sum, partner.status=contr.sum))
Anova(mod)
## One-Way MANOVA
## See ?Pottery for a description of the data set used in this example.
summary(Anova(lm(cbind(Al, Fe, Mg, Ca, Na) ~ Site, data=Pottery)))
## MANOVA for a randomized block design (example courtesy of Michael Friendly:
## See ?Soils for description of the data set)
soils.mod <- lm(cbind(pH,N,Dens,P,Ca,Mg,K,Na,Conduc) ~ Block + Contour*Depth,
data=Soils)
Manova(soils.mod)
summary(Anova(soils.mod), univariate=TRUE, multivariate=FALSE,
p.adjust.method=TRUE)
```


```{r}

## a multivariate linear model for repeated-measures data
## See ?OBrienKaiser for a description of the data set used in this example.
phase <- factor(rep(c("pretest", "posttest", "followup"), c(5, 5, 5)),
levels=c("pretest", "posttest", "followup"))
hour <- ordered(rep(1:5, 3))
idata <- data.frame(phase, hour)
idata
mod.ok <- lm(cbind(pre.1, pre.2, pre.3, pre.4, pre.5,
post.1, post.2, post.3, post.4, post.5,
fup.1, fup.2, fup.3, fup.4, fup.5) ~ treatment*gender,
data=OBrienKaiser)
(av.ok <- Anova(mod.ok, idata=idata, idesign=~phase*hour))
summary(av.ok, multivariate=FALSE)
```


```{r}

## A "doubly multivariate" design with two distinct repeated-measures variables
## (example courtesy of Michael Friendly)
## See ?WeightLoss for a description of the dataset.
imatrix <- matrix(c(
1,0,-1, 1, 0, 0,
1,0, 0,-2, 0, 0,
1,0, 1, 1, 0, 0,
0,1, 0, 0,-1, 1,
0,1, 0, 0, 0,-2,
0,1, 0, 0, 1, 1), 6, 6, byrow=TRUE)
colnames(imatrix) <- c("WL", "SE", "WL.L", "WL.Q", "SE.L", "SE.Q")
rownames(imatrix) <- colnames(WeightLoss)[-1]
(imatrix <- list(measure=imatrix[,1:2], month=imatrix[,3:6]))
contrasts(WeightLoss$group) <- matrix(c(-2,1,1, 0,-1,1), ncol=2)
(wl.mod<-lm(cbind(wl1, wl2, wl3, se1, se2, se3)~group, data=WeightLoss))
Anova(wl.mod, imatrix=imatrix, test="Roy")

```


```{r}
## mixed-effects models examples:
## Not run:
library(nlme)
example(lme)
Anova(fm2)
## End(Not run)
## Not run:
library(lme4)
example(glmer)
Anova(gm1)
## End(Not run)
```
####Details

The designations "type-II" and "type-III" are borrowed from SAS, but the definitions used here do
not correspond precisely to those employed by SAS. Type-II tests are calculated according to the
principle of marginality, testing each term after all others, except ignoring the term's higher-order
relatives; so-called type-III tests violate marginality, testing each term in the model after all of the
others. This definition of Type-II tests corresponds to the tests produced by SAS for analysis-ofvariance
models, where all of the predictors are factors, but not more generally (i.e., when there
are quantitative predictors). Be very careful in formulating the model for type-III tests, or the
hypotheses tested will not make sense.
As implemented here, type-II Wald tests are a generalization of the linear hypotheses used to generate
these tests in linear models.
For tests for linear models, multivariate linear models, andWald tests for generalized linear models,
Cox models, mixed-effects models, generalized linear models fit to survey data, and in the default
case, Anova finds the test statistics without refitting the model. The svyglm method simply calls the
default method and therefore can take the same arguments.
The standard R anova function calculates sequential ("type-I") tests. These rarely test interesting
hypotheses in unbalanced designs.
A MANOVA for a multivariate linear model (i.e., an object of class "mlm" or "manova") can optionally
include an intra-subject repeated-measures design. If the intra-subject design is absent (the
default), the multivariate tests concern all of the response variables. To specify a repeated-measures
design, a data frame is provided defining the repeated-measures factor or factors via idata, with
default contrasts given by the icontrasts argument. An intra-subject model-matrix is generated
from the formula specified by the idesign argument; columns of the model matrix corresponding to
different terms in the intra-subject model must be orthogonal (as is insured by the default contrasts).
Note that the contrasts given in icontrasts can be overridden by assigning specific contrasts to the
factors in idata. As an alternative, the within-subjects model matrix can be specified directly via
the imatrix argument. Manova is essentially a synonym for Anova for multivariate linear models.
If univariate tests are requested for the summary of a multivariate linear model, the object returned
contains a univaov component of "univaov"; print and as.data.frame methods are provided
for the "univaov" class.
For the default method to work, the model object must contain a standard terms element, and must
respond to the vcov, coef, and model.matrix functions. If any of these requirements is missing,
then it may be possible to supply it reasonably simply (e.g., by writing a missing vcov method for
the class of the model object).

#### Value
An object of class "anova", or "Anova.mlm", which usually is printed. For objects of class
"Anova.mlm", there is also a summary method, which provides much more detail than the print
method about theMANOVA, including traditional mixed-model univariate F-tests with Greenhouse-
Geisser and Huynh-Feldt corrections.
Warning
Be careful of type-III tests.




#### #### Author(s)
John Fox <jfox@mcmaster.ca>; the code for the Mauchly test and Greenhouse-Geisser and Huynh-
Feldt corrections for non-spericity in repeated-measures ANOVA are adapted from the functions
stats:::stats:::mauchly.test.SSD and stats:::sphericity by R Core; summary.Anova.mlm
and print.summary.Anova.mlm incorporates code contributed by Gabriel Baud-Bovy.

####References

* Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.
* Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.
* Hand, D. J., and Taylor, C. C. (1987) Multivariate Analysis of Variance and Repeated Measures: A Practical Approach for Behavioural Scientists. Chapman and Hall.
* O'Brien, R. G., and Kaiser, M. K. (1985) MANOVA method for analyzing repeated measures designs: An extensive primer. Psychological Bulletin 97, 316-333.

Anova 9
See Also
linearHypothesis, anova anova.lm, anova.glm, anova.mlm, anova.coxph, svyglm.