---
title: "``avPlots``: Added-Variable Plots"
author: "Kevin O Brien"
date: "22 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```


####Description

These functions construct added-variable, also called partial-regression, plots for linear and generalized
linear models.

####Usage
<pre><code>

avPlots(model, terms=~., intercept=FALSE, layout=NULL, ask, main, ...)
avp(...)
avPlot(model, ...)
## S3 method for class 'lm'
avPlot(model, variable,
id=TRUE, col = carPalette()[1], col.lines = carPalette()[2],
xlab, ylab, pch = 1, lwd = 2,
main=paste("Added-Variable Plot:", variable),
grid=TRUE,
ellipse=FALSE,
marginal.scale=FALSE, ...)
## S3 method for class 'glm'
avPlot(model, variable,
id=TRUE,
col = carPalette()[1], col.lines = carPalette()[2],
xlab, ylab, pch = 1, lwd = 2, type=c("Wang", "Weisberg"),
main=paste("Added-Variable Plot:", variable), grid=TRUE,
ellipse=FALSE, ...)

</code></pre>

####Arguments

* `` model ``:   model object produced by lm or glm.
* `` terms ``:    A one-sided formula that specifies a subset of the predictors. One added-variable
plot is drawn for each term. For example, the specification terms = ~.-X3
would plot against all terms except for X3. If this argument is a quoted name of
one of the terms, the added-variable plot is drawn for that term only.
* intercept Include the intercept in the plots; default is FALSE.
* `` variable``:    A quoted string giving the name of a regressor in the model matrix for the horizontal axis.
* ``layout`` :  If set to a value like c(1, 1) or c(4, 3), the layout of the graph will have
this many rows and columns. If not set, the program will select an appropriate layout. If the number of graphs exceed nine, you must select the layout yourself,
or you will get a maximum of nine per page. If layout=NA, the function does not set the layout and the user can use the par function to control the layout, for example to have plots from two models in the same graphics window.
* `` main ``:  The title of the plot; if missing, one will be supplied.
ask If TRUE, ask the user before drawing the next plot; if FALSE don't ask.
... avPlots passes these arguments to avPlot. avPlot passes them to plot. id controls point identification; if FALSE, no points are identified; can be a list of named arguments to the showLabels function; TRUE, the default, is equivalent to list(method=list(abs(residuals(model, type="pearson")), "x"), n=2, cex=1, col=carPalette()[which identifies the 2 points with the largest residuals and the 2 points with the most extreme horizontal values (i.e., largest partial leverage).
* `` col  ``:  color for points; the default is the second entry in the current car palette (see
carPalette and par).
* col.lines color for the fitted line.
pch plotting character for points; default is 1 (a circle, see par).
lwd line width; default is 2 (see par).
xlab x-axis label. If omitted a label will be constructed.
ylab y-axis label. If omitted a label will be constructed.
type if "Wang" use the method of Wang (1985); if "Weisberg" use the method in the
Arc software associated with Cook and Weisberg (1999).
grid If TRUE, the default, a light-gray background grid is put on the graph.
12 avPlots
ellipse controls plotting data-concentration ellipses. If FALSE (the default), no ellipses
are plotted. Can be a list of named values giving levels, a vector of one or
more bivariate-normal probability-contour levels at which to plot the ellipses;
and robust, a logical value determing whether to use the cov.trob function in
the MASS package to calculate the center and covariance matrix for the data
ellipses. TRUE is equivalent to list(levels=c(.5, .95), robust=TRUE).
marginal.scale Consider an added-variable plot of Y versus X given Z. If this argument is FALSE
then the limits on the horizontal axis are determined by the range of the residuals
from the regression of X on Z and the limits on the vertical axis are determined
by the range of the residuals from the regressnio of Y on Z. If the argument
is TRUE, then the limits on the horizontal axis are determined by the range of
X minus it mean, and on the vertical axis by the range of Y minus its means;
adjustment is made if necessary to include outliers. This scaling allows visualization
of the correlations between Y and Z and between X and Z. For example,
if the X and Z are highly correlated, then the points will be concentrated on the
middle of the plot.


####Examples

```{r}

avPlots(lm(prestige ~ income + education + type, data=Duncan))
avPlots(glm(partic != "not.work" ~ hincome + children,
data=Womenlf, family=binomial), id=FALSE)
m1 <- lm(partic ~ tfr + menwage + womwage + debt + parttime, Bfox)
```

```{r}
par(mfrow=c(1,3))
# marginal plot, ignoring other predictors:
with(Bfox, dataEllipse(womwage, partic, levels=0.5))
abline(lm(partic ~ womwage, Bfox), col="red", lwd=2)
# AV plot, adjusting for others:
avPlots(m1, ~ womwage, ellipse=list(levels=0.5))
# AV plot, adjusting and scaling as in marginal plot
avPlots(m1, ~ womwage, marginal.scale=TRUE, ellipse=list(levels=0.5))
```

####Details

The function intended for direct use is avPlots (for which avp is an abbreviation).
Value
These functions are used for their side effect id producing plots, but also invisibly return the coordinates
of the plotted points.
#### Author(s)
John Fox <jfox@mcmaster.ca>, Sanford Weisberg <sandy@umn.edu>

####References

Cook, R. D. and Weisberg, S. (1999) Applied Regression, Including Computing and Graphics.
Wiley.
* Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.
Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.
Wang, P C. (1985) Adding a variable in generalized linear models. Technometrics 27, 273-276.
* Weisberg, S. (2014) Applied Linear Regression, Fourth Edition, Wiley.
See Also
residualPlots, crPlots, ceresPlots, link{dataEllipse}, showLabels, dataEllipse.