---
title: "``durbinWatsonTest`` Durbin-Watson Test for Autocorrelated Errors"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```

####Description

Computes residual autocorrelations and generalized Durbin-Watson statistics and their bootstrapped
p-values. (``dwt`` is an abbreviation for ``durbinWatsonTest``.)

####Usage
<pre><code>

durbinWatsonTest(model, ...)
dwt(...)
## S3 method for class 'lm'
durbinWatsonTest(model, max.lag=1, simulate=TRUE, reps=1000,
method=c("resample","normal"),
alternative=c("two.sided", "positive", "negative"), ...)
## Default S3 method:
durbinWatsonTest(model, max.lag=1, ...)
## S3 method for class 'durbinWatsonTest'
print(x, ...)

</code></pre>

####Arguments

* `` model `` :  a linear-model object, or a vector of residuals from a linear model.
* max.lag maximum lag to which to compute residual autocorrelations and Durbin-Watson statistics.
* simulate if TRUE p-values will be estimated by bootstrapping.
* reps number of bootstrap replications.
* method bootstrap method: "resample" to resample from the observed residuals; "normal" to sample normally distributed errors with 0 mean and standard deviation equal to the standard error of the regression.
* alternative sign of autocorrelation in alternative hypothesis; specify only if max.lag = 1;
if max.lag > 1, then alternative is taken to be "two.sided".
* ... arguments to be passed down.
* x durbinWatsonTest object.

####Examples

```{r}

durbinWatsonTest(lm(fconvict ~ tfr + partic + degrees + mconvict, data=Hartnagel))
```

#### Value
Returns an object of type "durbinWatsonTest".

#### Note
p-values are available only from the lm method.

#### Author(s)
John Fox <jfox@mcmaster.ca>

####References

Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.
