---
title: "``panel.car``:   Panel Function for Coplotss"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```


#### Description

a panel function for use with coplot that plots points, a lowess line, and a regression line.

#### Usage
<pre><code>

panel.car(x, y, col, pch, cex=1, span=0.5, lwd=2,
reg.line=lm, lowess.line=TRUE, ...)

</code></pre>

#### Arguments

x vector giving horizontal coordinates.
y vector giving vertical coordinates.
col point color.
pch plotting character for points.
cex character expansion factor for points.
span span for lowess smoother.
lwd line width, default is 2.
reg.line function to compute coefficients of regression line, or FALSE for no line.
lowess.line if TRUE plot lowess smooth.
... other arguments to pass to functions lines and regLine.
Value
NULL. This function is used for its side effect: producing a panel in a coplot.

#### Author(s)

John Fox <jfox@mcmaster.ca>
See Also
coplot, regLine

####Examples

```{r}

coplot(prestige ~ income|education, panel=panel.car,
col="red", data=Prestige)

```
