---
title: "Citation History of Bland-Altman 1986"
subtitle: "Method Comparison Studies"
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE,warning=FALSE}
library(tidyverse)
library(rvest)
#library(textreadr)
library(janitor)
library(knitr)


```

"Statistical methods for assessing agreement between two methods of clinical measurement. (Lancet 327, 307–310 (1986)"
by Bland, J. M. & Altman, D. G. is one of the most cited papers of all time. This analysis tracks the citation history by since the paper's publication in the Lancet in 1986.

The search for Citations of Bland-Altman 1986 paper is from the following URL.

"http://scholar.google.com/scholar?hl=en&as_sdt=2005&sciodt=0%2C5&cites=7296362022254018043&scipsc=&as_ylo=2020&as_yhi=2020"


### Using textreadr

N.B. This package is no longer on CRAN
```{r eval=FALSE}
Citations_Data <- data.frame()
YEAR_RANGE <- 1986:2023

for(i in 1:length(YEAR_RANGE)){
  
    YEAR <- YEAR_RANGE[i]
    FILE <- paste0("http://scholar.google.com/scholar?hl=en&as_sdt=2005&sciodt=0%2C5&cites=7296362022254018043&scipsc=&as_ylo=",YEAR,"&as_yhi=",YEAR)

  
    LineList <- textreadr::read_html(FILE)
    LineList <- grep("About",LineList,value=T)
    LineList <- grep("results",LineList,value=T)
    LineList <- gsub("About ","",LineList)
    LineList <- gsub("results \\(","",LineList)
    LineList <- gsub(",","",LineList) %>% str_trim() %>% as.numeric()
    Citations_Data <- Citations_Data %>% bind_rows(
  data.frame(YEAR=YEAR,Citations = LineList)
)

}
```

### Using {rvest}


```{r eval=FALSE}
for(i in 1:length(YEAR_RANGE)){
  
  YEAR <- YEAR_RANGE[i]
  FILE <- paste0("http://scholar.google.com/scholar?hl=en&as_sdt=2005&sciodt=0%2C5&cites=7296362022254018043&scipsc=&as_ylo=",YEAR,"&as_yhi=",YEAR)
  
  LineList <- rvest::read_html(FILE) %>% 
    html_nodes("div.gs_ab_mdw") %>% 
    html_text()
  
  
  LineList <- LineList[2]
  LineList <- strsplit(LineList,"results")[[1]][1]
  LineList <- gsub("About ","",LineList) %>% str_trim() 
  Citations_Data <- Citations_Data %>% bind_rows(
    data.frame(YEAR=YEAR,Citations = LineList)
  )
}
```

### Making some adjustments to the Data

```{r eval=FALSE}
Bland_Altman_Citation <- Citations_Data
Bland_Altman_Citation <- Bland_Altman_Citation %>% 
  mutate(Citations=as.numeric(Citations)) %>% mutate(Year_Complete =case_when(
    YEAR == 1986 ~ "No",
    YEAR %in% 1987:2022 ~ "Yes",
    YEAR == 2023 ~ "No"
  ))
Bland_Altman_Citation %>% write.csv("Bland_Altman_Citations.csv",row.names = FALSE)
```

### Using a saved copy of the data

Repeated use of the programme may result in 429 Gateway errors.

```{r message=FALSE,warning=FALSE}
Bland_Altman_Citation <- read_csv("Bland_Altman_Citations.csv")
Bland_Altman_Citation %>% adorn_totals("row") %>% kable()
```

```{r}
Bland_Altman_Citation %>% ggplot( aes(x=as.factor(YEAR),y=Citations,fill=Year_Complete)) + 
  geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```