---
title: "Kolmogorov-Smirnov Test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Kolmogorov-Smirnov Test

For a single sample of data, the Kolmogorov-Smirnov test is used to test whether or not the sample of data is consistent with a specified distribution function. 


When there are two samples of data, it is used to test whether or not these two samples may reasonably be assumed to come from the same distribution.
The null and alternative hypotheses are as follows:

$H_0$: The two data sets are drawn from the same distribution

$H_1$: The data sets are not drawn from the same distribution


### Example

Consider two sample data sets X and Y that are both normally distributed with similar means and variances.

```{r}
X=rnorm(16,mean=20,sd=5)
Y=rnorm(18,mean=21,sd=4)
ks.test(X,Y)
```

#### Remark:

It doesn’t not suffice that both datasets are from the same distribution. 
They must have the same value for the defining parameters. 
Consider the case of data sets; X and Z. 
Both are normally distributed, but with different mean values.

```{r}
X=rnorm(16,mean=20,sd=5)
Z=rnorm(16,mean=14,sd=5)
ks.test(X,Z)
```



