
{gvlma}
========================

* The {gvlma} package is a comprehensive, automatic testing suite for many of the assumptions of general linear models. 
* It does both statistical tests and diagnostic plots using an extremely simple implementation for powerful results.


{gvlma}
========================

The package is an implementation of a paper by Pena & Slate called ***Global Validation of Linear Model Assumptions*** and allows you to quickly check for:

* ***Linearity*** - the Global Stat tests for the null hypothesis that our model is a linear combination of its predictors.

* ***Homoscedasticity*** - the respective stat tests for the null that the residial variance is relatively constant over the range of values.

* ***Normality*** - skewness and kurtosis tests help you understand if the residuals fits a normal distribution. 

* If the null is rejected you probably need to transform your data in some way (like a log transform). 
This can also be assessed by looking at the normal probability plot it generates.	

{gvlma}
========================

```{r echo=FALSE,include=FALSE}
model <- Fit_4
```

```{r}
  library("gvlma")

  # model <- lm(y ~ x, data)
  
  
  summary(gvlma(model))
```


{gvlma}
========================
* The diagnostic plots also let you understand the relation between your data and these assumptions visually. 
* Other useful capabilities are the link function test which is used for understanding whether the underlying data is categorical or continuous.


{gvlma}
========================
class:smaller-95



```{r out.width = "80%" }

  plot(gvlma(model))

```





