## ``agreementplot`` 

Bangdiwala's Observer Agreement Chart

#### Desription

Representation of a kk confusion matrix, where the observed and expected diagonal elements are represented by superposed black and white rectangles, respectively. The function also computes a statistic measuring the strength of agreement (relation of respective area sums).

#### Usage
<pre><code>

## Default S3 method:
agreementplot(x, reverse_y = TRUE, main = NULL,
weights = c(11/(ncol(x) - 1)^2), margins = par("mar"),
newpage = TRUE, pop = TRUE,
xlab = names(dimnames(x))[2],
ylab = names(dimnames(x))[1],
xlab_rot = 0, xlab_just = "center",
ylab_rot = 90, ylab_just = "center",
fill_col = function(j) gray((1 - (weights[j]) ^ 2) ^ 0.5),
line_col = "red", xscale = TRUE, yscale = TRUE,
return_grob = FALSE,
prefix = "", ...)
## S3 method for class 'formula'
agreementplot(formula, data = NULL, ..., subset)

</code></pre>
#### Arguments

x a confusion matrix, i.e., a table with equal-sized dimensions.
reverse_y if TRUE, the y axis is reversed (i.e., the rectangles' positions correspond to the contingency table).
main user-specified main title.
weights vector of weights for successive larger observed areas, used in the agreement strength statistic, and also for the shading. The first element should be 1.
margins vector of margins (see par).
newpage logical; if TRUE, the plot is drawn on a new page.
* pop logical; if TRUE, all newly generated viewports are popped after plotting.
* return_grob logical. Should a snapshot of the display be returned as a grid grob?
* xlab, ylab labels of x- and y-axis.
* xlab_rot, ylab_rot
* rotation angle for the category labels.
* xlab_just, ylab_just
justification for the category labels.
fill_col a function, giving the fill colors used for exact and partial agreement
* line_col color used for the diagonal reference line
* formula a formula, such ax. For details, see xtabs.
* data a data frame (or list), or a contingency table from which the variables in formula should be taken.
* subset an optional vector specifying a subset of the rows in the data frame to be used for plotting.
* xscale, yscale logicals indicating whether the marginals should be added on the x-axis/y-axis, respectively.
* prefix character string used as prefix for the viewport name
* ... further graphics parameters (see par).


#### Examples
```{r}

data("SexualFun")
agreementplot(t(SexualFun))
data("MSPatients")
## Not run:
## best visualized using a resized device, e.g. using:
## get(getOption("device"))(width = 12)
pushViewport(viewport(layout = grid.layout(ncol = 2)))
pushViewport(viewport(layout.pos.col = 1))
agreementplot(t(MSPatients[,,1]), main = "Winnipeg Patients",
newpage = FALSE)
popViewport()
pushViewport(viewport(layout.pos.col = 2))
agreementplot(t(MSPatients[,,2]), main = "New Orleans Patients",
newpage = FALSE)
popViewport(2)
dev.off()
## End(Not run)
## alternatively, use cotabplot:
cotabplot(MSPatients, panel = cotab_agreementplot)
```

























Details
Weights can be specified to allow for partial agreement, taking into account contributions from offdiagonal
cells. Partial agreement is typically represented in the display by lighter shading, as given
by fill_col(j), corresponding to weights[j].
agreementplot 5
A weight vector of length 1 means strict agreement only, each additional element increases the
maximum number of disagreement steps.
cotabplot can be used for stratified analyses (see 
#### Examples
```{r}
).
Value
Invisibly returned, a list with components
Bangdiwala the unweighted agreement strength statistic.
Bangdiwala_Weighted
the weighted statistic.
weights the weight vector used.
Author(s)
David Meyer <David.Meyer@R-project.org>
References
Bangdiwala, S. I. (1988). The Agreement Chart. Department of Biostatistics, University of North
Carolina at Chapel Hill, Institute of Statistics Mimeo Series No. 1859, http://www.stat.ncsu.
edu/in
</code></pre>
#### Format 
ion/library/mimeo.archive/ISMS_1988_1859.pdf
Bangdiwala, S. I., Ana S. Haedo, Marcela L. Natal, and Andres Villaveces. The agreement chart as
an alternative to the receiver-operating characteristic curve for diagnostic tests. Journal of Clinical
Epidemiology, 61 (9), 866-874.
* Michael Friendly (2000), Visualizing Categorical Data. SAS Institute, Cary, NC.
