---
title: "vcd::cd_plot "
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```

Conditional Density Plots

#### Desription

Computes and plots conditional densities describing how the distribution of a categorical variable y changes over a numerical variable x.

#### Usage
<pre><code>

cd_plot(x, ...)
## Default S3 method:
cd_plot(x, y,
plot = TRUE, ylab_tol = 0.05,
bw = "nrd0"512, from = NULL, to = NULL,
main = "", xlab = NULL, ylab = NULL, margins = c(5.1, 4.1, 4.1, 3.1),
gp = gpar(), name = "cd_plot", newpage = TRUE, pop = TRUE, return_grob = FALSE, ...)
## S3 method for class 'formula'
cd_plot(formula, data = list(),
plot = TRUE, ylab_tol = 0.05,
bw = "nrd0"512, from = NULL, to = NULL,
main = "", xlab = NULL, ylab = NULL, margins = c(5.1, 4.1, 4.1, 3.1),
gp = gpar(), name = "cd_plot", newpage = TRUE, pop = TRUE, return_grob = FALSE, ...)

</code></pre>
#### Arguments

* ``x``: an object, the default method expects either a single numerical variable.
y a "factor" interpreted to be the dependent variable
formula a "formula" of typx with a single dependent "factor" and a single
numerical explanatory variable.
data an optional data frame.
plot logical. Should the computed conditional densities be plotted?
ylab_tol convenience tolerance parameter for y-axis annotation. If the distance between
two labels drops under this threshold, they are plotted equidistantly.
bw, n, from, to, ...: Aguments
 passed to density
* main, xlab, ylab
 character strings for annotation
* margins margins when calling plotViewport
* gp a "gpar" object controlling the grid graphical parameters of the rectangles.
It should specify in particular a vector of fill colors of the same length as
levels(y). The default is to call gray.colors.
* name name of the plotting viewport.
* newpage logical. Should grid.newpage be called before plotting?

* return_grob logical. Should a snapshot of the display be returned as a grid grob?
* pop logical. Should the viewport created be popped?

#### Examples
```{r}

## Arthritis data
data("Arthritis")
cd_plot(Improved ~ Age, data = Arthritis)
cd_plot(Improved ~ Age, data = Arthritis, bw = 3)
cd_plot(Improved ~ Age, data = Arthritis, bw = "SJ")
## compare with spinogram
spine(Improved ~ Age, data = Arthritis, breaks = 3)
```

```{r}
## Space shuttle data
data("SpaceShuttle")
cd_plot(Fail ~ Temperature, data = SpaceShuttle, bw = 2)
## scatter plot with conditional density
cdens <- cd_plot(Fail ~ Temperature, data = SpaceShuttle, bw = 2, plot = FALSE)
plot(I(-1 * (as.numeric(Fail) - 2)) ~ jitter(Temperature, factor = 2), data = SpaceShuttle,
xlab = "Temperature", ylab = "Failure")
lines(53:81, cdens[[1]](53:81), col = 2)
```

Details
cd_plot computes the conditional densities of x given the levels of y weighted by the marginal
distribution of y. The densities are derived cumulatively over the levels of y.
This visualization technique is similar to spinograms (see spine) but they do not discretize the
explanatory variable, but rather use a smoothing approach. Furthermore, the original x axis and not
a distorted x axis (as for spinograms) is used. This typically results in conditional densities that are
based on very few observations in the margins: hence, the estimates are less reliable there.
Value
The conditional density functions (cumulative over the levels of y) are returned invisibly.
Author(s)
Achim Zeileis <Achim.Zeileis@R-project.org>
References
Hofmann, H., Theus, M. (2005), Interactive graphics for visualizing conditional distributions, Unpublished
Manuscript.
See Also
spine, density
