---
title: "vcd::grid_barplot "
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```

#### Desription

Bar plots of 1-way tables in grid.

#### Usage
<pre><code>

grid_barplot(height, width = 0.8, offset = 0,
names = NULL, xlim = NULL, ylim = NULL, xlab = "", ylab = "", main = "",
gp = gpar(fill = "lightgray"), name = "grid_barplot",
newpage = TRUE, pop = FALSE, return_grob = FALSE)

</code></pre>
#### Arguments

* height either a vector or a 1-way table of frequencies.
* width width of the bars (recycled if needed to the number of bars).
* offset offset of the bars (recycled if needed to the number of bars).
* names a vector of names for the bars, if set to NULL the names of height are used.
* xlim limits for the x axis.
* ylim limits for the y axis.
* xlab a label for the x axis.
* ylab a label for the y axis.
* main a title for the plot.
* gp a "gpar" object controlling the grid graphical parameters of the rectangles.
* name name of the plotting viewport.
* newpage logical. Should grid.newpage be called before plotting?
* pop logical. Should the viewport created be popped?
* return_grob logical. Shall the plot be returned as a grob object?

#### Examples
```{r}

grid_barplot(sample(1:6), names = letters[1:6])
```
Details
grid_barplot mimics (some of) the features of barplot, but currently it only supports 1-way
tables.
Author(s)
Achim Zeileis <Achim.Zeileis@R-project.org>
