
rootogram Rootograms

#### Description

Rootograms of observed and fitted 
#### Value 
.

#### Usage
<pre><code>

## Default S3 method:
rootogram(x, fitted, names = NULL, scale = c("sqrt", "raw"),
type = c("hanging", "standing", "deviation"),
shade = FALSE, legend = TRUE,
legend_args = list(x = 00.2, height = 0.6), df = NULL,
rect_gp = NULL, rect_gp_args = list(),
lines_gp = gpar(col = "red", lwd = 2),
points_gp = gpar(col = "red"), pch = 19,
xlab = NULL, ylab = NULL, ylim = NULL,
main = NULL, sub = NULL,
margins = unit(0, "lines"),
title_margins = NULL, legend_width = NULL,
main_gp = gpar(fontsize = 20),
sub_gp = gpar(fontsize = 15),
name = "rootogram", prefix = "",
keep_aspect_ratio = FALSE,
newpage = TRUE, pop = TRUE,
return_grob = FALSE, ...)

</code></pre>
#### Arguments

x either a vector or a 1-way table of frequencies.
fitted a vector of fitted frequencies.
names a vector of names passed to grid_barplot, if set to NULL the names of x are used.
* ``scale``:  a character string indicating whether the value should be plotted on the raw or square root scale.
type a character string indicating if the bars for the observed frequencies should be hanging or standing or indicate the deviation between observed and fitted frequencies.
shade logical specifying whether rect_gp should be set to colors corresponding to the pearson residuals, i.e., if a residual-based shading should be applied to the bars.
legend either a legend-generating function, or a legend function (see details and legends), or a logical. If legend is NULL or TRUE and gp is a function, legend defaults to legend_resbased.
legend_args list of arguments for the legend-generating function, if specified.
100 rootogram
df degrees of freedom passed to the shading functions used for inference.
rect_gp a "gpar" object controlling the grid graphical parameters of the rectangles, shading function or a corresponding generating function (see shadings). If unspecified and no shading is applied, defaults to light grey fill color for the
bars.
* ``rect_gp_args``: list of arguments for the shading-generating function, if specified for rect_gp. lines_gp a "gpar" object controlling the grid graphical parameters of the lines.
points_gp a "gpar" object controlling the grid graphical parameters of the points.
pch plotting character for the points.
* ``xlab``:  a label for the x axis.
* ``ylab``: a label for the y axis.
ylim limits for the y axis.
main either a logical, or a character string used for plotting the main title. If main is a logical and TRUE, the name of the object supplied as x is used.
sub a character string used for plotting the subtitle. If sub is a logical and TRUE and main is unspecified, the name of the object supplied as x is used.
margins either an object of class "unit" of length 4, or a numeric vector of length 4.
The elements are recycled as needed. The four components specify the top, right, bottom, and left margin of the plot, respectively. When a numeric vector
is supplied, the numbers are interpreted as "lines" units. In addition, the
unit or numeric vector may have named arguments
 ('top', 'right', 'bottom',
and 'left'), in which case the non-named arguments
 specify the default 
#### Value 

(recycled as needed), overloaded by the named arguments
.
title_margins either an object of class "unit" of length 2, or a numeric vector of length 2. The elements are recycled as needed. The two components specify the top and bottom
title margin of the plot, respectively. The default for each specified title are 2 lines (and 0 else), except when a legend is plotted and keep_aspect_ratio is TRUE: in this case, the default value 
 of both margins are set as to align the heights of legend and actual plot. When a numeric vector is supplied, the numbers are interpreted as "lines" units. In addition, the unit or numeric vector may
have named Arguments  ('top' and 'bottom'), in which case the non-named argument
specify the default value (recycled as needed), overloaded by the named arguments.
legend_width An object of class "unit" of length 1 specifying the width of the legend (if any).
Default: 5 lines.
main_gp, sub_gp
object of class "gpar" containing the graphical parameters used for the main (sub) title, if specified.
name name of the plotting viewport.
keep_aspect_ratio
logical indicating whether the aspect ratio should be fixed or not.
prefix optional character string used as a prefix for the generated viewport and grob names.

newpage logical. Should grid.newpage be called before plotting?
pop logical. Should the viewport created be popped?
return_grob logical. Should a snapshot of the display be returned as a grid grob?
``...``: further arguments passed to grid_barplot.
Details
The observed frequencies are displayed as bars and the fitted frequencies as a line. By default a sqrt scale is used to make the smaller frequencies more visible.
Author(s)
Achim Zeileis <Achim.Zeileis@R-project.org>, David Meyer <David.Meyer@R-project.org>
References
J. W. Tukey (1977), Exploratory Data Analysis. Addison Wesley, Reading, MA.
M. Friendly (2000), Visualizing Categorical Data. SAS Institute, Cary, NC.
See Also
grid_barplot


#### Simulated data Examples
```{r}

dummy <- rnbinom(200, size = 1.5, prob = 0.8)
observed <- table(dummy)
fitted1 <- dnbinom(as.numeric(names(observed)),
size = 1.5, prob = 0.8) * sum(observed)
fitted2 <- dnbinom(as.numeric(names(observed)),
size = 2, prob = 0.6) * sum(observed)
rootogram(observed, fitted1)
rootogram(observed, fitted2)
```

#### Real data Examples
```{r}

data("HorseKicks")
HK.fit <- goodfit(HorseKicks)
summary(HK.fit)
plot(HK.fit)
## or equivalently
rootogram(HK.fit)
data("Federalist")
F.fit <- goodfit(Federalist, type = "nbinomial")
summary(F.fit)
plot(F.fit)
## (Pearson) residual-based shading
data("Federalist")
Fed_fit0 <- goodfit(Federalist, type = "poisson")
plot(Fed_fit0, shade = TRUE)
```
