---
title: "``vcd::ternaryplot``"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```

ternaryplot Ternary Diagram

#### Desription

Visualizes compositional, 3-dimensional data in an equilateral triangle.

#### Usage
<pre><code>

ternaryplot(x, scale = 1, dimnames = NULL,
dimnames_position = c("corner","edge","none"),
dimnames_color = "black", dimnames_rot = c(-60, 60, 0),
id = NULL, id_color = "black",
id_just = c("center", "center"),
coordinates = FALSE, grid = TRUE, grid_color = "gray",
labels = c("inside", "outside", "none"),
labels_color = "darkgray", labels_rot = c(120, -120, 0),
border = "black", bg = "white",
pch = 19, cex = 1, prop_size = FALSE, col = "red",
main = "ternary plot", newpage = TRUE, pop = TRUE,
return_grob = FALSE, ...)

</code></pre>

#### Arguments

* x a matrix with three columns.
* scale row sums scale to be used.
* dimnames dimension labels (defaults to the column names of x).
* dimnames_position, dimnames_color position and color of dimension labels.
dimnames_rot Numeric vector of length 3, specifying the angle of the dimension labels.
id optional labels to be plotted below the plot symbols. coordinates and id are
mutual exclusive.
* id_color color of these labels.
* id_just character vector of length 1 or 2 indicating the justification of these labels.
* coordinates if TRUE, the coordinates of the points are plotted below them. coordinates and id are mutual exclusive.
grid if TRUE, a grid is plotted. May optionally be a string indicating the line type
(default: "dotted").
* grid_color grid color.
* labels, labels_color
position and color of the grid labels.
* labels_rot Numeric vector of length 3, specifying the angle of the grid labels.
border color of the triangle border.
ternaryplot 129
* bg triangle background.
* pch plotting character. Defaults to filled dots.
* cex a numerical value giving the amount by which plotting text and symbols should be scaled relative to the default. Ignored for the symbol size if prop_size is not FALSE.
* ``prop_size``: if TRUE, the symbol size is plotted proportional to the row sum of the three
variables, i.e., represents the weight of the observation.
col plotting color.
main main title.
newpage if TRUE, the plot will appear on a new graphics page.
pop logical; if TRUE, all newly generated viewports are popped after plotting.
return_grob logical. Should a snapshot of the display be returned as a grid grob?
... additional graphics parameters (see par)

#### Examples
```{r}

data("Arthritis")
## Build table by crossing Treatment and Sex
tab <- as.table(xtabs(~ I(Sex:Treatment) + Improved, data = Arthritis))
## Mark groups
col <- c("red", "red", "blue", "blue")
pch <- c(1, 19, 1, 19)
```

```{r}
## plot
ternaryplot(
tab,
col = col,
pch = pch,
prop_size = TRUE,
bg = "lightgray",
grid_color = "white",
labels_color = "white",
main = "Arthritis Treatment Data"
)
## legend
grid_legend(0.8, 0.7, pch, col, rownames(tab), title = "GROUP")
```

```{r}
## Titanic
data("Lifeboats")
attach(Lifeboats)
ternaryplot(
Lifeboats[,4:6],
pch = ifelse(side == "Port", 1, 19),
col = ifelse(side == "Port", "red", "blue"),
id = ifelse(men / total > 0.1, as.character(boat), NA),
prop_size = 2,
dimnames_position = "edge",
main = "Lifeboats on Titanic"
)
grid_legend(0.8, 0.9, c(1, 19),
c("red", "blue"), c("Port", "Starboard"),
title = "SIDE")
```

```{r}
## Hitters
data("Hitters")
attach(Hitters)
colors <- c("black","red","green","blue","red","black","blue")
pch <- substr(levels(Positions), 1, 1)
ternaryplot(
Hitters[,2:4],
pch = as.character(Positions),
col = colors[as.numeric(Positions)],
main = "Baseball Hitters Data"
)
grid_legend(0.8, 0.9, pch, colors, levels(Positions),
title = "POSITION(S)")
```

#### Details
A points' coordinates are found by computing the gravity center of mass points using the data entries
as weights. Thus, the coordinates of a point P(a; b; c)= 1, are: P(b + c=2; c
p
3=2).
Author(s)
David Meyer <David.Meyer@R-project.org>
References
M. Friendly (2000), Visualizing Categorical Data. SAS Institute, Cary, NC.
