---
title: "assoc"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```

assoc Extended Association Plots

#### Desription

Produce an association plot indicating deviations from a specified independence model in a possibly
high-dimensional contingency table.

#### Usage
<pre><code>

## Default S3 method:
assoc(x, row_vars = NULL, col_vars = NULL, compress = TRUE,
xlim = NULL, ylim = NULL,
spacing = spacing_conditional(sp = 0), spacing_args = list(),
split_vertical = NULL, keep_aspect_ratio = FALSE,
xscale = 0.9, yspace = unit(0.5, "lines"), main = NULL, sub = NULL,
..., residuals_type = "Pearson", gp_axis = gpar(lty = 3))
## S3 method for class 'formula'
assoc(formula, data = NULL, ..., subset = NULL, na.action = NULL, main = NULL, sub = NULL)

</code></pre>
#### Arguments

* ``x``: a contingency table in array form with optional category labels specified in the
dimnames(x) attribute, or an object inheriting from the "ftable" class (such as
"structable" objects).
* row_vars a vector of integers giving the indices, or a character vector giving the names of
the variables to be used for the rows of the association plot.
* col_vars a vector of integers giving the indices, or a character vector giving the names of
the variables to be used for the columns of the association plot.
* compress logical; if FALSE, the space between the rows (columns) are chosen such that
the total heights (widths) of the rows (columns) are all equal. If TRUE, the space between rows and columns is fixed and hence the plot is more “compressed”.
* xli k matrix of doubles, k number of total columns of the plot. The columns
of xlim correspond to the columns of the association plot, the rows describe
the column ranges (minimums in the first row, maximums in the second row).
If xlim is NULL, the ranges are determined from the residuals according to
compress (if TRUE: widest range from each column, if FALSE: from the whole
association plot matrix).
yli k matrix of doubles, k number of total rows of the plot. The columns of
ylim correspond to the rows of the association plot, the rows describe the column
ranges (minimums in the first row, maximums in the second row). If ylim
is NULL, the ranges are determined from the residuals according to compress (if
TRUE: widest range from each row, if FALSE: from the whole association plot
matrix).
spacing a spacing object, a spacing function, or a corresponding generating function
(see strucplot for more inFormat 
ion). The default is the spacing-generating
function spacing_conditional that is (by default) called with the argument
list spacing_args (see spacings for more details).
spacing_args list of Arguments
 for the spacing-generating function, if specified (see strucplot
for more inFormat 
ion).
split_vertical vector of logicals of length k, where k is the number of margins of x (default:
FALSE). 
Value 
 are recycled as needed. A TRUE component indicates that the
corresponding dimension is folded into the columns, FALSE folds the dimension
into the rows.

* keep_aspect_ratio
logical indicating whether the aspect ratio should be fixed or not.
residuals_type a character string indicating the type of residuals to be computed. Currently,
only Pearson residuals are supported.
xscale scale factor resizing the tile’s width, thus adding additional space between the
tiles.
yspace object of class "unit" specifying additional space separating the rows.
* `` gp_axis`` :   object of class "gpar" specifying the visual aspects of the tiles’ baseline.
formula a formula object with possibly both left and right hand sides specifying the column
and row variables of the flat table.
* data a data frame, list or environment containing the variables to be cross-tabulated,
or an object inheriting from class table.
subset an optional vector specifying a subset of observations to be used. Ignored if
data is a contingency table.
na.action an optional function which indicates what should happen when the data contain
NAs. Ignored if data is a contingency table.
main, sub either a logical, or a character string used for plotting the main (sub) title. If
logical and TRUE, the name of the data object is used.
... other parameters passed to strucplot

#### Details
Association plots have been suggested by Cohen (1980) and extended by Friendly (1992) and provide
a means for visualizing the residuals of an independence model for a contingency table.
assoc is a generic function and currently has a default method and a formula interface. Both are
high-level interfaces to the strucplot function, and produce (extended) association plots. Most
of the functionality is described there, such as specification of the independence model, labeling,
legend, spacing, shading, and other graphical parameters.
For a contingency table, the signed contribution to Pearson’s 2 for cell fi: kg is
dij:::k =
(fij:::k 􀀀 eij:::k)
p
eij:::k
where fij:::k and eij:::k are the observed and expected counts corresponding to the cell. In the
association plot, each cell is represented by a rectangle that has (signed) height proportional to dij:::k
assoc 9
and width proportional to p
eij:::k, so that the area of the box is proportional to the difference in
observed and expected frequencies. The rectangles in each row are positioned relative to a baseline
indicating independence (dij:::k = 0). If the observed frequency of a cell is greater than the expected
one, the box rises above the baseline, and falls below otherwise.
Additionally, the residuals can be colored depending on a specified shading scheme (see Meyer et
al., 2003). Package vcd offers a range of residual-based shadings (see the shadings help page).
Some of them allow, e.g., the visualization of test statistics.
Unlike the assocplot function in the graphics package, this function allows the visualization of
contingency tables with more than two dimensions. Similar to the construction of ‘flat’ tables (like
objects of class "ftable" or "structable"), the dimensions are folded into rows and columns.
The layout is very flexible: the specification of shading, labeling, spacing, and legend is modularized
(see strucplot for details).
Value
The "structable" visualized is returned invisibly.
Author(s)
David Meyer <David.Meyer@R-project.org>
References
Cohen, A. (1980), On the graphical display of the significant components in a two-way contingency
table. Communications in Statistics—Theory and Methods, A9, 1025–1041.
Friendly, M. (1992), Graphical methods for categorical data. SAS User Group International Conference
Proceedings, 17, 190–200. http://datavis.ca/papers/sugi/sugi17.pdf
Meyer, D., Zeileis, A., Hornik, K. (2003), Visualizing independence using extended association
plots. Proceedings of the 3rd International Workshop on Distributed Statistical Computing, K.
Hornik, F. Leisch, A. Zeileis (eds.), ISSN 1609-395X. https://www.R-project.org/conferences/
DSC-2003/Proceedings/
Meyer, D., Zeileis, A., and Hornik, K. (2006), The strucplot framework: Visualizing multi-way contingency
tables with vcd. Journal of Statistical Software, 17(3), 1-48. URL http://www.jstatsoft.org/v17/i03/
and available as vignette("strucplot").
See Also
mosaic, strucplot, structable

#### Examples
```{r}

data("HairEyeColor")
## Aggregate over sex:
(x <- margin.table(HairEyeColor, c(1, 2)))
## Ordinary assocplot:
assoc(x)
## and with residual-based shading (of independence)
10 assocstats
assoc(x, main = "Relation between hair and eye color", shade = TRUE)
## Aggregate over Eye color:
(x <- margin.table(HairEyeColor, c(1, 3)))
chisq.test(x)
assoc(x, main = "Relation between hair color and sex", shade = TRUE)
# Visualize multi-way table
assoc(aperm(HairEyeColor), expecte(Hair + Eye) * Sex,
labeling_args = list(just_labels = c(Eye = "left"),
offset_labels = c(right = -0.5),
offset_varnames = c(right = 1.2),
rot_labels = c(right = 0),
tl_varnames = c(Eye = TRUE))
)
assoc(aperm(UCBAdmissions), expecte(Admit + Gender) * Dept, compress = FALSE,
labeling_args = list(abbreviate = c(Gender = TRUE), rot_labels = 0)
)
