---
title: "``vcd::assocstats``"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```

## Association Statistics

#### Desription

Computes the Pearson chi-Squared test, the Likelihood Ratio chi-Squared test, the phi coefficient, the contingency coefficient and Cramer's V for possibly stratified contingency tables.

#### Usage

<pre><code>

assocstats(x)

</code></pre>

#### Arguments

* ``x``: a contingency table, with possibly more than 2 dimensions. In this case, all dimensions except the first two ones are considered as strata.

#### Examples
```{r}

data("Arthritis")
tab <- xtabs(~Improved + Treatment, data = Arthritis)
```

```{r}
tab
```


```{r}
summary(assocstats(tab))
```


```{r}
assocstats(UCBAdmissions)
```


#### Value
In case of a 2-dimensional table, a list with components:
* ``chisq_test``: 3 table with the chi-squared statistics.
* ``phi``: The absolute value of the phi coefficient (only defined fo2 tables).
* ``cont``: The contingency coefficient.
* ``cramer``: Cramer's V.

In case of higher-dimensional tables, a list of the above mentioned structure, each list component representing one stratum defined by the combinations of all levels of the stratum dimensions.

#### Author(s)
David Meyer <David.Meyer@R-project.org>

#### References
* Michael Friendly (2000), Visualizing Categorical Data. SAS Institute, Cary, NC. Fleiss, J. L. (1981). Statistical methods for rates and proportions (2nd ed). New York: Wiley
