---
title: "coindep_test "
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```
Test for (Conditional) Independence

#### Desription

Performs a test of (conditional) independence of 2 margins in a contingency table by simulation from the marginal distribution of the input table under (conditional) independence.

#### Usage
<pre><code>

coindep_test(x, margin = NULL1000,
indepfun = function(x) max(abs(x)), aggfun = max,
alternative = c("greater", "less"),
pearson = TRUE)

</code></pre>
#### Arguments

* ``x``: a contingency table.
* margin margin index(es) or corresponding name(s) of the conditioning variables. Each resulting conditional table has to be a 2-way table.
* ``n`` : number of (conditional) independence tables to be drawn.
* ``indepfun``: aggregation function capturing independence in (each conditional) 2-way table.
* aggfun aggregation function aggregating the test statistics computed by indepfun.
* alternative a character string specifying the alternative hypothesis; must be either "greater"
(default) or "less" (and may be abbreviated.)
* ``pearson``: logical. Should the table of Pearson residuals under independence be computed and passed to indepfun (default) or the raw table of observed frequencies?


#### Examples
```{r}

library(MASS)
TeaTasting <- matrix(c(3, 1, 1, 3), nr = 2,
dimnames = list(Guess = c("Milk", "Tea"),
Truth = c("Milk", "Tea"))
)
## compute maximum statistic
coindep_test(TeaTasting)
```

```{r}
## compute Chi-squared statistic
coindep_test(TeaTasting, indepfun = function(x) sum(x^2))
```

```{r}
## use unconditional asymptotic distribution
chisq.test(TeaTasting, correct = FALSE)
```

```{r}
chisq.test(TeaTasting)
data("UCBAdmissions")
## double maximum statistic
coindep_test(UCBAdmissions, margin = "Dept")
```

```{r}
## maximum of Chi-squared statistics
coindep_test(UCBAdmissions, margin = "Dept", indepfun = function(x) sum(x^2))
```

```{r}
## Pearson Chi-squared statistic
coindep_test(UCBAdmissions, margin = "Dept", indepfun = function(x) sum(x^2), aggfun = sum)
```

```{r}
## use unconditional asymptotic distribution
loglm(~ Dept * (Gender + Admit), data = UCBAdmissions)
```


#### Details
If margin is NULL this computes a simple independence statistic in a 2-way table. Alternatively, margin can give several conditioning variables and then conditional independence in the resulting conditional table is tested.
By default, this uses a (double) maximum statistic of Pearson residuals. By changing indepfun or aggfun a (maximum of) Pearson Chi-squared statistic(s) can be computed or just the usual Pearson Chi-squared statistics and so on. Other statistics can be computed by changing pearson to FALSE.
The function uses r2dtable to simulate the distribution of the test statistic under the null.

#### Value
A list of class "coindep_test" inheriting from "htest" with following components:
* statistic the value of the test statistic.
* p.value the p value for the test.
* method a character string indicating the type of the test.
* data.name a character string giving the name(s) of the data.
* observed observed table of frequencies
* expctd expected table of frequencies
* residuals corresponding Pearson residuals
* margin the margin used
* dist a vector of size n with simulated value of the distribution of the statistic under
the null.
* qdist the corresponding quantile function (for computing critical Value ).
pdist the corresponding distribution function (for computing p value ).

#### Author(s)
Achim Zeileis <Achim.Zeileis@R-project.org>

#### See Also
chisq.test, fisher.test, r2dtable
