---
title: "``vcd::cotab_panel``"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```

cotab_panel Panel-generating Functions for Contingency Table Coplots

#### Desription

Panel-generating functions visualizing contingency tables that can be passed to cotabplot.

#### Usage
<pre><code>
cotab_mosaic(x = NULL, condvars = NULL, ...)
cotab_assoc(x = NULL, condvars = NULL, ylim = NULL, ...)
cotab_sieve(x = NULL, condvars = NULL, ...)
cotab_loddsratio(x = NULL, condvars = NULL, ...)
cotab_agreementplot(x = NULL, condvars = NULL, ...)
cotab_fourfold(x = NULL, condvars = NULL, ...)
cotab_coindep(x, condvars,
  test = c("doublemax", "maxchisq", "sumchisq"),
  level = NULL1000, interpolate = c(2, 4),
  h = NULLNULLNULL, lty = 1,
  type = c("mosaic", "assoc"), legend = FALSE, ylim = NULL, ...)

</code></pre>
#### Arguments

* ``x``: a contingency tables in array form.
* ``condvars``: margin name(s) of the conditioning variables.
* ``ylim``: y-axis limits for assoc plot. By default this is computed from x.
* ``test``: character indicating which type of statistic should be used for assessing conditional independence.
* ``level``,``n``,``h``,``c``,``l``,``lty``,``interpolate``: variables controlling the HCL shading of the residuals, see shadings for more details.
* ``type``: character indicating which type of plot should be produced.
* ``legend``: logical. Should a legend be produced in each panel?
* ``...``: further arguments passed to the plotting function (such as mosaic or assoc or sieve respectively).

#### Examples
```{r}

data("UCBAdmissions")
cotabplot(~ Admit + Gender | Dept, data = UCBAdmissions)
cotabplot(~ Admit + Gender | Dept, data = UCBAdmissions, panel = cotab_assoc)
cotabplot(~ Admit + Gender | Dept, data = UCBAdmissions, panel = cotab_fourfold)
ucb <- cotab_coindep(UCBAdmissions, condvars = "Dept", type = "assoc",
n = 5000, margins = c(3, 1, 1, 3))
cotabplot(~ Admit + Gender | Dept, data = UCBAdmissions, panel = ucb)
```

#### Details
These functions of class "panel_generator" are panel-generating functions for use with cotabplot,
i.e., they return functions with the interface
panel(x, condlevels)
required for cotabplot. The functions produced by cotab_mosaic, cotab_assoc and cotab_sieve essentially only call co_table to produce the conditioned table and then call mosaic, assoc or sieve respectively with the arguments  specified.
The function cotab_coindep is similar but additionally chooses an appropriate residual-based shading visualizing the associated conditional independence model. The conditional independence test is carried out via coindep_test and the shading is set up via ``shading_hcl``.
A Desription
 of the underlying ideas is given in Zeileis, Meyer, Hornik (2005).

#### Author(s)
Achim Zeileis <Achim.Zeileis@R-project.org>
References
Meyer, D., Zeileis, A., and Hornik, K. (2006), The strucplot framework: Visualizing multi-way contingency
tables with vcd. Journal of Statistical Software, 17(3), 1-48. URL http://www.jstatsoft.org/v17/i03/
and available as vignette("strucplot").
Zeileis, A., Meyer, D., Hornik K. (2007), Residual-based shadings for visualizing (conditional)
independence, Journal of Computational and Graphical Statistics, 16, 507–525.

See Also
cotabplot, mosaic, assoc, sieve, co_table, coindep_test, shading_hcl
30 co_table
