
---
title: "grid_legend"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```

Legend Function for grid Graphics

#### Desription

This function can be used to add legends to grid-based plots.

#### Usage
<pre><code>

grid_legend(x, y, pch = NA, col = par('col'), labels, frame = TRUE,
hgap = unit(0.8, "lines"), vgap = unit(0.8, "lines"), default_units = "lines", gp = gpar(), draw = TRUE,
title = NULL, just = 'center', lwd = NA, lty = NA, size = 1,
gp_title = NULL, gp_labels = NULL, gp_frame = gpar(fill = "transparent"),
inset = c(0, 0))

</code></pre>
#### Arguments

x character string "topright", "topleft", "bottomright", "bottomleft", "top",
"bottom", "left", "right", "center" or x coordinate of the legend.
y y coordinates of the legend.
pch integer vector of plotting symbols, if any.
* col character vector of colors for the symbols.
* labels character vector of labels corresponding to the symbols.
* frame logical indicating whether the legend should have a border or not.
* hgap object of class "unit" specifying the space between symbols and labels.
* vgap object of class "unit" specifying the space between the lines.
* default_units character string indicating the default unit.
* gp object of class "gpar" used for the legend.
* draw logical indicating whether the legend be drawn or not.
* title character string indicating the plot’s title.
* just justification of the legend relative to its (x, y) location. see ?viewport for more details.
* lwd positive number to set the line width. if specified lines are drawn.
* lty line type. if specified lines are drawn.
* size size of the group symbols (in char units).
* gp_title object of class "gpar" used for the title.
* gp_labels object of class "gpar" used for the labels.
* gp_frame object of class "gpar" used for the frame.
* inset numeric vector of length 2 specifying the inset of the legend in npc units, relative to the specified x and y coordinates.


#### Examples
```{r}

data("Lifeboats")
attach(Lifeboats)
ternaryplot(Lifeboats[,4:6],
pch = ifelse(side == "Port", 1, 19),
col = ifelse(side == "Port", "red", "blue"),
id = ifelse(men / total > 0.1, as.character(boat), NA),
prop_size = 2,
dimnames_position = "edge",
main = "Lifeboats on Titanic")
grid_legend(0.8, 0.9, c(1, 19), c("red", "blue"),
c("Port", "Starboard"), title = "SIDE")
grid.newpage()
pushViewport(viewport(height = .9, width = .9 ))
grid.rect(gp = gpar(lwd = 2, lty = 2))
grid_legend(x = unit(.05,'npc'),
y = unit(.05,'npc'),
just = c(0,0),
pch = c(1,2,3),
col = c(1,2,3),
lwd=NA,
lty=NA,
labels = c("b",'r','g'),
title = NULL,
gp=gpar(lwd=2, cex=1),
hgap = unit(.8, "lines"),
vgap = unit(.9, "lines"))
grid_legend(x = unit(1,'npc'),
y = unit(1,'npc'),
just = c(1,1),
pch = NA,
col = c(1,2,3,4),
lwd=c(1,1,1,3),
lty=c(1,2,1,3),
labels = c("black",'red','green','blue'),
gp_labels = list(gpar(col = 1), gpar(col = 2), gpar(col = 3), gpar(col = 4)),
46 Hitters
title = NULL,
gp=gpar(lwd=2, cex=1),
hgap = unit(.8, "lines"),
vgap = unit(.9, "lines"))
grid_legend(x = 'topleft',
pch = c(1,NA,2,NA),
col = c(1,2,3,4),
lwd=NA,
lty=c(NA,2,NA,3),
labels = c("black",'red','green','blue'),
title = 'Some LONG Title',
gp_title = gpar(col = 3),
gp_frame = gpar(col = 4, lty = 2, fill = "transparent"),
gp_labels = gpar(col = 6),
gp=gpar(lwd=2, cex=2, col = 1),
hgap = unit(.8, "lines"),
vgap = unit(.9, "lines"))
grid_legend(x = .7,
y = .7,
pch = c(1,NA,2,NA),
col = c(1,2,3,4),
lwd=1,
lty=c(NA,2,NA,3),
labels = c("black",'red','green','blue'),
title = 'short T',
gp=gpar(lwd=1, cex=.7,col = 1),
hgap = unit(.8, "lines"),
vgap = unit(.9, "lines"))
grid_legend(x = 'bottomright',
pch = c(1,NA,2,NA),
col = c(2),
lwd=NA,
lty=c(NA,2,NA,3),
labels = c("black",'red','green','blue'),
title = NULL,
gp=gpar(lwd=2, cex=1,col = 1),
hgap = unit(.8, "lines"),
vgap = unit(.9, "lines"))


grid_legend 45
Value
Invisibly, the legend as a "grob" object.
Author(s)
David Meyer <David.Meyer@R-project.org> Florian Gerber <florian.gerber@math.uzh.ch>
See Also
legend
