---
title: "vcd::labeling_border "
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```
Labeling Functions for Strucplots

#### Desription

These functions generate labeling functions used for strucplots.

#### Usage
<pre><code>

labeling_border(labels = TRUE, varnames = labels,
set_labels = NULL, set_varnames = NULL,
tl_labels = NULL, alternate_labels = FALSE, tl_varnames = NULL,
gp_labels = gpar(fontsize = 12),
gp_varnames = gpar(fontsize = 12, fontface = 2),
rot_labels = c(0, 90, 0, 90), rot_varnames = c(0, 90, 0, 90),
pos_labels = "center", pos_varnames = "center",
just_labels = "center", just_varnames = pos_varnames,
boxes = FALSE, fill_boxes = FALSE,
offset_labels = c(0, 0, 0, 0), offset_varnames = offset_labels,
labbl_varnames = NULL, labels_varnames = FALSE, sep = ": ",
abbreviate_labs = FALSE, rep = TRUE, clip = FALSE, ...)
labeling_Value 
(value_type = c("observed", "expected", "residuals"),
suppress = NULL, digits = 1, clip_cells = FALSE, ...)
labeling_residuals(suppress = NULL, digits = 1, clip_cells = FALSE, ...)
labeling_conditional(...)
labeling_left(rep = FALSE, pos_varnames = "left",
pos_labels = "left", just_labels = "left", ...)
labeling_left2(tl_labels = TRUE, clip = TRUE, pos_varnames = "left",
pos_labels = "left", just_labels = "left", ...)
labeling_cboxed(tl_labels = TRUE, boxes = TRUE, clip = TRUE,
pos_labels = "center", ...)
labeling_lboxed(tl_labels = FALSE, boxes = TRUE, clip = TRUE,
pos_labels = "left", just_labels = "left",
labbl_varnames = FALSE, ...)
labeling_doubledecker(lab_pos = c("bottom", "top"), dep_varname = TRUE,
boxes = NULL, clip = NULL, labbl_varnames = FALSE,
rot_labels = rep.int(0, 4),
pos_labels = c("left", "center", "left", "center"),
just_labels = c("left", "left", "left", "center"),
varnames = NULL, gp_varnames = gpar(fontsize = 12, fontface = 2),
offset_varnames = c(0, -0.6, 0, 0), tl_labels = NULL, ...)

</code></pre>
#### Arguments

* ``labels``: vector of logicals indicating whether labels should be drawn for a particular
dimension.

* varnames vector of logicals indicating whether variable names should be drawn for a particular dimension.
set_labels An optional character vector with named components replacing the so-specified variable names. The component names must exactly match the variable names to be replaced.
set_varnames An optional list with named components of character vectors replacing the labels
of the so-specified variables. The component names must exactly match the variable names whose labels should be replaced.
tl_* ``labels``: vector of logicals indicating whether labels should be positioned on top (column
labels) / left (row labels) for a particular dimension.
alternate_labels
vector of logicals indicating whether labels should be alternated on the top/bottom (left/right) side of the plot for a particular dimension.
tl_varnames vector of logicals indicating whether variable names should be positioned on top
(column labels) / on left (row labels) for a particular dimension.
* gp_labels list of objects of class "gpar" used for drawing the labels.
* gp_varnames list of objects of class "gpar" used for drawing the variable names.
* rot_labels vector of rotation angles for the labels for each of the four sides of the plot.
* rot_varnames vector of rotation angles for the variable names for each of the four sides of the plot.
* pos_labels character string of label positions ("left", "center", "right") for each of the variables.
* pos_varnames character string of variable names positions ("left", "center", "right") for each of the four sides of the plot.
* just_labels character string of label justifications ("left", "center", "right") for each of the variables.
* just_varnames character string of variable names justifications ("left", "center", "right") for each of the four sides of the plot.
* boxes vector of logicals indicating whether boxes should be drawn around the labels for a particular dimension.
* fill_boxes Either a vector of logicals, or a vector of characters, or a list of such vectors, specifying the fill colors for the boxes. "TRUE" and "FALSE"  Value 
 are transformed
into "grey" and "white", respectively. If fill_boxes is atomic, each
component specifies a basic color for the corresponding dimension. This color
is transformed into its HSV representation, and the value is varied from 50% to
100% to give a sequential color palette for the levels. For NA components, no
palette is produced (no fill color). If fill_boxes is a list of vectors, each vector
specifies the level colors of the corresponding dimension.
offset_labels, offset_varnames
numeric vector of length 4 indicating the offset of the labels (variable names)
for each of the four sides of the plot.
labbl_varnames vector of logicals indicating whether variable names should be drawn on the left
(column variables) / on top (row variables) of the corresponding labels.
labeling_border 57
labels_varnames
vector of logicals indicating, for each dimension, whether the variable name
should be added to the corresponding labels or not.
sep separator used if any component of "labels_varnames" is TRUE.
abbreviate_labs
vector of integers or logicals indicating, for each dimension, the number of characters
the labels should be abbreviated to. TRUE means 1 character, FALSE causes
no abbreviation. 
#### Value 
 are recycled as needed.
rep vector of logicals indicating, for each dimension, whether labels should be repeated
for all conditioning strata, or appear only once.
clip vector of integers indicating, for each dimension, whether labels should be clipped
to not overlap.
lab_pos character string switching between "top" or "bottom" position of the labels
(only used for labeling_doubledecker).
* dep_varname logical or character string. If logical, this is indicating whether the name of the dependent variable should be printed or not. A character string will be printed instead of the variable name taken from the dimnames.
value_type character string specifying which Value 
 should be displayed in the cells.
suppress numeric vector of length 2 specifying an interval of 
#### Value 
 that are not displayed.
0 
#### Value 
 are never displayed. A single number, k, is treated as c(-k, k).
The default for labeling residuals is c(-2,2). Use suppresto show all
non-zero 
#### Value 
.
digits integer specifying the number of digits used for rounding.
clip_cells logical indicating whether the 
#### Value 
 should be clipped at the cell borders.
... only used for labeling_conditional and labeling_doubledecker: parameters
passed to labeling_cells and labeling_border.
Details
These functions generate labeling functions called by strucplot for their side-effect of adding
labels to the plot. They suppose that a strucplot has been drawn and the corresponding viewport
structure is pushed, since the positions of the viewports are used for the label positioning. Note that
the functions can also be used 'stand-alone' as shown in the 
#### Examples
```{r}
.
All 
#### Value 
 supplied to vectorized 
</code></pre>
#### Arguments
 can be 'abbreviated' by using named components which
override the default component 
#### Value 
. In addition, these defaults can be overloaded by the sequence
of non-named components which are recycled as needed (see examples).
This help page only documents labeling_border and derived functions, more functions are described
on the help page for labeling_cells and labeling_list.
labeling_left, labeling_left2, labeling_cboxed, and labeling_lboxed are really just wrappers
to labeling_border, and good examples for the parameter Usage.
labeling_residuals is a trivial wrapper for labeling_Value
, which in turn calls labeling_border
by additionally adding the observed or expected frequencies or residuals to the cells.

Value
A function with 
</code></pre>
#### Arguments
:
d "dimnames" attribute from the visualized contingency table, or the visualized
table itself from which the "dimnames" attributes will then be extracted.
split_vertical vector of logicals indicating the split directions.
condvars integer vector of conditioning dimensions.
Author(s)
David Meyer <David.Meyer@R-project.org>

#### References
Meyer, D., Zeileis, A., and Hornik, K. (2006), The strucplot framework: Visualizing multi-way contingency
tables with vcd. Journal of Statistical Software, 17(3), 1-48. URL http://www.jstatsoft.org/v17/i03/
and available as vignette("strucplot").
See Also
labeling_cells, labeling_list, structable, grid.text

#### Examples
```{r}

data("Titanic")
mosaic(Titanic)
mosaic(Titanic, labeling = labeling_left)
labeling_left
mosaic(Titanic, labeling = labeling_cboxed)
labeling_cboxed
mosaic(Titanic, labeling = labeling_lboxed)
labeling_lboxed
data("PreSex")
mosaic(~ PremaritalSex + ExtramaritalSex | Gender + MaritalStatus,
data = PreSex, labeling = labeling_conditional)
## specification of vectorized  Arguments

mosaic(Titanic, labeling_args = list(abbreviate_labs = c(Survived = TRUE)))
mosaic(Titanic, labeling_args = list(clip = TRUE, boxes = TRUE,
fill_boxes = c(Survived = "green", "red")))
mosaic(Titanic, labeling_args = list(clip = TRUE, boxes = TRUE,
fill_boxes = list(Sex = "red", "green")))
mosaic(Titanic, labeling_args = list(clip = TRUE, boxes = TRUE,
labeling_cells_list 59
fill_boxes = list(Sex = c(Male = "red", "blue"), "green")))
## change variable names
mosaic(Titanic, labeling_args = list(set_varnames = c(Sex = "Gender")))
## change labels
mosaic(Titanic, labeling_args = list(set_varnames = c(Survived = "Status"),
set_labels = list(Survived = c("Survived", "Not Survived")), rep = FALSE))
## show frequencies
mosaic(Titanic, labeling = labeling_
#### Value 
)
