---
title: "``mosaic ``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```

Extended Mosaic Plots

#### Desription

Plots (extended) mosaic displays.

#### Usage
<pre><code>

## Default S3 method:
mosaic(x, condvars = NULL,
split_vertical = NULL, direction = NULL, spacing = NULL,
spacing_args = list(), gp = NULL, expected = NULL, shade = NULL,
highlighting = NULL, highlighting_fill = grey.colors, highlighting_direction = NULL,
zero_size = 0.5, zero_split = FALSE, zero_shade = NULL,
zero_gp = gpar(col = 0), panel = NULL, main = NULL, sub = NULL, ...)
## S3 method for class 'formula'
mosaic(formula, data, highlighting = NULL,
..., main = NULL, sub = NULL, subset = NULL, na.action = NULL)

</code></pre>
#### Arguments

* ``x``: a contingency table in array form, with optional category labels specified in the
dimnames(x) attribute, or an object of class "structable".
* condvars vector of integers or character strings indicating conditioning variables, if any.
The table will be permuted to order them first.
* formula a formula specifying the variables used to create a contingency table from data.
For convenience, conditioning formulas can be specified; the conditioning variables
will then be used first for splitting. If any, a specified response variable
will be highlighted in the cells.
data either a data frame, or an object of class "table" or "ftable".
#### Examples
```{r}

library(MASS)
data("Titanic")
mosaic(Titanic)
## Formula interface for tabulated data plus shading and legend:
mosaic(~ Sex + Age + Survived, data = Titanic,
main = "Survival on the Titanic", shade = TRUE, legend = TRUE)
data("HairEyeColor")
mosaic(HairEyeColor, shade = TRUE)
## Independence model of hair and eye color and sex. Indicates that
## there are significantly more blue eyed blond females than expected
## in the case of independence (and too few brown eyed blond females).
mosaic(HairEyeColor, shade = TRUE, expected = list(c(1,2), 3))
## Model of joint independence of sex from hair and eye color. Males
## are underrepresented among people with brown hair and eyes, and are
## overrepresented among people with brown hair and blue eyes, but not
## "significantly".
## Formula interface for raw data: visualize crosstabulation of numbers
## of gears and carburettors in Motor Trend car data.
data("mtcars")
mosaic(~ gear + carb, data = mtcars, shade = TRUE)
data("PreSex")
mosaic(PreSex, condvars = c(1,4))
mosaic(~ ExtramaritalSex + PremaritalSex | MaritalStatus + Gender,
data = PreSex)
## Highlighting:
mosaic(Survived ~ ., data = Titanic)
data("Arthritis")
mosaic(Improved ~ Treatment | Sex, data = Arthritis, zero_size = 0)
mosaic(Improved ~ Treatment | Sex, data = Arthritis, zero_size = 0,
highlighting_direction = "right")
```
#### Other Arguments
subset an optional vector specifying a subset of observations to be used.
na.action a function which indicates what should happen when the data contain NAs. Ignored
if data is a contingency table.
zero_size size of the bullets used for zero entries (if 0, no bullets are drawn).
74 mosaic
zero_split logical controlling whether zero cells should be further split. If FALSE and
zero_shade is FALSE, only one bullet is drawn (centered) for unsplit zero cells.
If FALSE and zero_shade is TRUE, a bullet for each zero cell is drawn to allow,
e.g., residual-based shadings to be effective also for zero cells.
zero_shade logical controlling whether zero bullets should be shaded. The default is TRUE if
shade is TRUE or expected is not null or gp is not null, and FALSE otherwise.
zero_gp object of class "gpar" used for zero bullets in case they are not shaded.
split_vertical vector of logicals of length k, where k is the number of margins of x (default:
FALSE). 
#### Value 
 are recycled as needed. A TRUE component indicates that the
tile(s) of the corresponding dimension should be split vertically, FALSE means
horizontal splits. Ignored if direction is not NULL.
direction character vector of length k, where k is the number of margins of x (
#### Value 

are recycled as needed). For each component, a value of "h" indicates that the
tile(s) of the corresponding dimension should be split horizontally, whereas "v"
indicates vertical split(s).
spacing spacing object, spacing function, or corresponding generating function (see strucplot
for more information). The default is spacing_equal if x has two dimensions,
spacing_increase for more dimensions, and spacing_conditional if conditioning
variables are specified using condvars or the formula interface.
spacing_args list of  Arguments
 for the generating function, if specified (see strucplot for
more information).
gp object of class "gpar", shading function or a corresponding generating function
(see details and shadings). Components of "gpar" objects are recycled as
needed along the last splitting dimension. Ignored if shade = FALSE.
shade logical specifying whether gp should be used or not (see gp). If TRUE and
expected is unspecified, a default model is fitted: if condvars (see strucplot)
is specified, a corresponding conditional independence model, and else the total
independence model.
expected optionally, an array of expected 
#### Value 
 of the same dimension as x, or alternatively
the corresponding independence model specification as used by loglin
or loglm (see strucplot).
highlighting character vector or integer specifying a variable to be highlighted in the cells.
highlighting_fill
color vector or palette function used for a highlighted variable, if any.
highlighting_direction
Either "left", "right", "top", or "bottom" specifying the direction of highlighting
in the cells.
panel Optional function with 
</code></pre>
#### Arguments
: residuals, observed, expected, index,
gp, and name called by the struc_mosaic workhorse for each tile that is drawn
in the mosaic. index is an integer vector with the tile's coordinates in the contingency
table, gp a gpar object for the tile, and name a label to be assigned to
the drawn grid object.
main, sub either a logical, or a character string used for plotting the main (sub) title. If
logical and TRUE, the name of the data object is used.
... Other 
</code></pre>
#### Arguments
 passed to strucplot
mosaic 75
Details
Mosaic displays have been suggested in the statistical literature by Hartigan and Kleiner (1984) and
have been extended by Friendly (1994). mosaicplot is a base graphics implementation and mosaic
is a much more flexible and extensible grid implementation.
mosaic is a generic function which currently has a default method and a formula interface. Both
are high-level interfaces to the strucplot function, and produce (extended) mosaic displays. Most
of the functionality is described there, such as specification of the independence model, labeling,
legend, spacing, shading, and other graphical parameters.
A mosaic plot is an area proportional visualization of a (possibly higher-dimensional) table of expected
frequencies. It is composed of tiles (corresponding to the cells) created by recursive vertical
and horizontal splits of a square. The area of each tile is proportional to the corresponding cell
entry, given the dimensions of previous splits.
An extended mosaic plot, in addition, visualizes the fit of a particular log-linear model. Typically,
this is done by residual-based shadings where color and/or outline of the tiles visualize sign, size
and possibly significance of the corresponding residual.
The layout is very flexible: the specification of shading, labeling, spacing, and legend is modularized
(see strucplot for details).
In contrast to the mosaicplot function in graphics, the splits start with the horizontal direction by
default to match the printed output of structable.
Value
The "structable" visualized is returned invisibly.
Author(s)
David Meyer <David.Meyer@R-project.org>
References
Hartigan, J.A., and Kleiner, B. (1984), A mosaic of television ratings. The American Statistician,
38, 32-35.
Emerson, J. W. (1998), Mosaic displays in S-PLUS: A general implementation and a case study.
Statistical Computing and Graphics Newsletter (ASA), 9, 1, 17-23.
Friendly, M. (1994), Mosaic displays for multi-way contingency tables. Journal of the American
Statistical Association, 89, 190-200.
Meyer, D., Zeileis, A., and Hornik, K. (2006), The strucplot framework: Visualizing multi-way
contingency tables with vcd. Journal of Statistical Software, 17(3), 1-48. URL http://www.
jstatsoft.org/v17/i03/ and available as vignette("strucplot", package = "vcd").
The home page of * Michael Friendly (http://datavis.ca) provides in
</code></pre>
#### Format 
ion on various aspects
of graphical methods for analyzing categorical data, including mosaic plots. In particular, there are
many materials for his course "Visualizing Categorical Data with SAS and R" at http://datavis.
ca/courses/VCD/.
76 mplot
See Also
assoc, strucplot, mosaicplot, structable, doubledecker

