---
title: "plot.loddsratio "
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```

Plotting (Log) Odds Ratios

#### Desription

Produces a (conditional) line plot of extended (log) odds ratios.

#### Usage
<pre><code>

## S3 method for class 'loddsratio'
plot(x, baseline = TRUE, gp_baseline = gpar(lty = 2),
lines = TRUE, lwd_lines = 3, confidence = TRUE, conf_level = 0.95, lwd_confidence = 2,
whiskers = 0, transpose = FALSE, col = NULL, cex = 0.8, pch = NULL,
bars = NULL, gp_bars = gpar(fill = "lightgray", alpha = 0.5),
bar_width = unit(0.05, "npc"), legend = TRUE, legend_pos = "topright", legend_inset = c(0, 0),
legend_vgap = unit(0.5, "lines"), gp_legend_frame = gpar(lwd = 1, col = "black"),
gp_legend_title = gpar(fontface = "bold"),
gp_legend = gpar(), legend_lwd = 1, legend_size = 1,
xlab = NULL, ylab = NULL, xlim = NULL, ylim = NULL,
main = NULL, gp_main = gpar(fontsize = 12, fontface = "bold"),
newpage = TRUE, pop = FALSE, return_grob = FALSE,
add = FALSE, prefix = "", ...)

## S3 method for class 'loddsratio'
lines(x, legend = FALSE, confidence = FALSE, cex = 0, ...)

</code></pre>
#### Arguments

* ``x``: an object of class loddsratio.
baseline if TRUE, a dashed line is plotted at a value of 1 (in case of odds) or 0 (in case of
log-odds).
gp_baseline object of class "gpar" used for the baseline.
lines if TRUE, the points are connected by lines (only sensible if the variable represented
by the x-axis is ordinal).
* lwd_lines Width of the connecting lines (in char units).
* confidence logical; shall confindence intervals be plotted?
* conf_level : confidence level used for confidence intervals.
lwd_confidence Line width of the confidence interval bars (in char units).
whiskers width of the confidence interval whiskers.
* transpose if TRUE, the plot is transposed.
col character vector specifying the colors of the fitted lines, by default chosen with
rainbow_hcl.
cex size of the plot symbols (in lines).
pch character or numeric vector of symbols used for plotting the (possibly conditioned)
observed Value 
, recycled as needed.
bars logical; shall bars be plotted additionally to the points? Defaults to TRUE in case
of only one conditioning variable.
gp_bars object of class "gpar" used for the bars.
bar_width Width of the bars, if drawn.
legend logical; if TRUE (default), a legend is drawn.
legend_pos numeric vector of length 2, specifying x and y coordinates of the legend, or a
character string (e.g., "topleft", "center" etc.). Defaults to "topleft" if the
fitted curve’s slope is positive, and "topright" else.
legend_inset numeric vector or length 2 specifying the inset from the legend’s x and y coordinates
in npc units.
legend_vgap vertical space between the legend’s line entries.
gp_legend_frame
object of class "gpar" used for the legend frame.
gp_legend_title
object of class "gpar" used for the legend title.
gp_legend object of class "gpar" used for the legend defaults.
legend_lwd line width used in the legend for the different groups.
legend_size size used for the group symbols (in char units).
xlab label for the x-axis. Defaults to "Strata" if transpose is FALSE.
ylab label for the y-axis. Defaults to "Strata" if transpose is TRUE.
xlim x-axis limits. Ignored if transpose is FALSE.
ylim y-axis limits. Ignored if transpose is TRUE.
main user-specified main title.
gp_main object of class "gpar" used for the main title.
newpage logical; if TRUE, the plot is drawn on a new page.
pop logical; if TRUE, all newly generated viewports are popped after plotting.
return_grob logical. Should a snapshot of the display be returned as a grid grob?
add logical; should the plot added to an existing log odds ratio plot?
prefix character string used as prefix for the viewport name.
... other graphics parameters (see par).

##### Details
The function basically produces conditioned line plots of the (log) odds ratios structure provided in
x.
The lines method can be used to overlay different plots (for example, observed and expected

#### Value 
).
cotabplot can be used for stratified analyses (see 
#### Examples
.
Value
if return_grob is TRUE, a grob object corresponding to the plot. NULL (invisibly) else.
Author(s)
David Meyer <David.Meyer@R-project.org>
References
M. Friendly (2000), Visualizing Categorical Data. SAS Institute, Cary, NC.
See Also
loddsratio

#### Examples
```{r}

#cases
data(CoalMiners, package = "vcd")
lor_CM <- loddsratio(CoalMiners)
plot(lor_CM)
lor_CM_df <- as.data.frame(lor_CM)
# fit linear models using WLS
age <- seq(20, 60, by = 5)
lmod <- lm(LOR ~ age, weight/ ASE^2, data = lor_CM_df)
grid.lines(seq_along(age), fitted(lmod), gp = gpar(col = "blue", lwd = 2), default.units = "native")
qmod <- lm(LOR ~ poly(age, 2), weight/ ASE^2, data = lor_CM_df)
grid.lines(seq_along(age), fitted(qmod), gp = gpar(col = "red", lwd = 2), default.units = "native")
#k x 2
lor_Emp <-loddsratio(Employment)
plot(lor_Emp)
## 4 way tables
data(Punishment, package = "vcd")
mosaic(attitude ~ age + education + memory, data = Punishment,
highlighting_direction="left", rep = c(attitude = FALSE))
# visualize the log odds ratios, by education
plot(loddsratio(~ attitude + memory | education, data = Punishment))
# visualize the log odds ratios, by age
plot.loglm 93
plot(loddsratio(~ attitude + memory | age, data = Punishment))
# visualize the log odds ratios, by age and education
plot(loddsratio(~ attitude + memory | age + education, data = Punishment))
# same, transposed
plot(loddsratio(~ attitude + memory | age + education, data = Punishment), transpose = TRUE)
# alternative visualization methods
image(loddsratio(Freq ~ ., data = Punishment))
tile(loddsratio(Freq ~ ., data = Punishment))
## cotabplots for more complex tables
cotabplot(Titanic, cond = c("Age","Sex"), panel = cotab_loddsratio)
cotabplot(Freq ~ opinion + grade + year | gender, data = JointSports, panel = cotab_loddsratio)
cotabplot(Freq ~ opinion + grade | year + gender, data = JointSports, panel = cotab_loddsratio)
