---
title: "``vcd::plot.loglm`` "
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```
## Visualize Fitted Log-linear Models

#### Desription

Visualize fitted "loglm" objects by mosaic or association plots.

#### Usage
<pre><code>

## S3 method for class 'loglm'
plot(x, panel = mosaic, type = c("observed", "expected"),
residuals_type = c("pearson", "deviance"), gp = shading_hcl, gp_args = list(),
...)

</code></pre>
#### Arguments

* ``x`` : a fitted "loglm" object, see ``loglm``.
* ``panel`` :  a panel function for visualizing the observed Value , residuals and expected Value. Currently, mosaic and assoc in vcd.
* ``type`` :  a character string indicating whether the observed or the expected value of the table should be visualized.
* ``residuals_type``:  a character string indicating the type of residuals to be computed.
* ``gp`` :  object of class "gpar", shading function or a corresponding generating function (see details and shadings). Ignored if ``shade = FALSE``.
* ``gp_args``:  list of arguments for the shading-generating function, if specified.
* ``...`` :  Other arguments passed to the panel function.




#### Examples
```{r}

library(MASS)
## mosaic display for PreSex model
data("PreSex")
fm <- loglm(~ PremaritalSex * ExtramaritalSex * (Gender + MaritalStatus),
data = aperm(PreSex, c(3, 2, 4, 1)))
fm
```
```{r}
## visualize Pearson statistic
plot(fm, split_vertical = TRUE)
## visualize LR statistic
plot(fm, split_vertical = TRUE, residuals_type = "deviance")
```
```{r}
## conditional independence in UCB admissions data
data("UCBAdmissions")
fm <- loglm(~ Dept * (Gender + Admit), data = aperm(UCBAdmissions))
## use mosaic display
plot(fm, labeling_args = list(abbreviate = c(Admit = 3)))
```
```{r}
## and association plot
plot(fm, panel = assoc)
assoc(fm)
```

#### Details

The plot method for "loglm" objects by default visualizes the model using a mosaic plot (can be changed to an association plot by setting ``panel = assoc``) with a shading based on the residuals of this model. The legend also reports the corresponding p value of the associated goodness-of-fit test. The mosaic and assoc methods are simple convenience interfaces to this plot method, setting the panel argument accordingly.

#### Value
The "structable" visualized is returned invisibly.

#### Author(s)
Achim Zeileis <Achim.Zeileis@R-project.org>

