---
title: "``vcd::struc_sieve``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```

### Core-generating Function for Sieve Plots

#### Desription

Core-generating function for strucplot returning a function producing sieve plots.

#### Usage
<pre><code>

struc_sieve(sievetype = c("observed","expected"), 
  gp_tile = gpar(),
  scale = 1)

</code></pre>
#### Arguments

* ``sievetype``: logical indicating whether rectangles should be filled according to observed or expected frequencies.
* ``gp_tile``: object of class "gpar", controlling the appearance of all static elements of the cells (e.g., border and fill color).
* ``scale``: Scaling factor for the sieve.

#### Examples
```{r}

## Titanic data
data("Titanic")
strucplot(Titanic, core = struc_sieve)
```

#### Details
This function is usually called by strucplot (typically when called by sieve) and returns a function
used by strucplot to produce sieve plots.

#### Value
A function with Arguments:

* ``residuals``: table of residuals.
* ``observed``: table of observed Value.
* ``expected``: not used by ``struc_sieve``.
* ``spacing``: object of class "unit" specifying the space between the tiles.
* ``gp``: list of gpar objects used for the drawing the tiles.
* ``split_vertical``: vector of logicals indicating, for each dimension of the table, the split direction.

#### Author(s)
David Meyer <David.Meyer@R-project.org>


#### References
* Riedwyl, H., and Schüpbach, M. (1994), Parquet diagram to plot contingency tables. In F. Faulbaum (ed.), Softstat-93: Advances in Statistical Software, 293-299. Gustav Fischer, New York.
* Friendly, M. (2000), Visualizing Categorical Data, SAS Institute, Cary, NC.
* Meyer, D., Zeileis, A., and Hornik, K. (2006), The strucplot framework: Visualizing multi-way contingency tables with vcd. Journal of Statistical Software, 17(3), 1-48. URL http://www.jstatsoft.org/v17/i03/
and available as vignette("strucplot").

#### See Also
sieve, strucplot, structable
