---
title: "strucplot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
strucplot Structured Displays of Contingency Tables

#### Desription

This modular function visualizes certain aspects of high-dimensional contingency tables in a hierarchical way.

#### Usage
<pre><code>

strucplot(x, residuals = NULL, expected = NULL,
condvars = NULL, shade = NULL, type = c("observed", "expected"),
residuals_type = NULL, df = NULL, split_vertical = NULL,
spacing = spacing_equal, spacing_args = list(),
gp = NULL, gp_args = list(),
labeling = labeling_border, labeling_args = list(),
core = struc_mosaic, core_args = list(),
legend = NULL, legend_args = list(),
main = NULL, sub = NULL, margins = unit(3, "lines"),
116 strucplot
title_margins = NULL, legend_width = NULL,
main_gp = gpar(fontsize = 20), sub_gp = gpar(fontsize = 15),
newpage = TRUE, pop = TRUE, return_grob = FALSE,
keep_aspect_ratio = NULL, prefix = "", ...)

</code></pre>
#### Arguments

* ``x``: a contingency table in array form, with optional category labels specified in the
dimnames attribute.
residuals optionally, an array of residuals of the same dimension as x (see details).
expected optionally, an array of expected 
#### Value 
 of the same dimension as x, or alternatively
the corresponding independence model specification as used by loglin
or loglm (see details).
df degrees of freedom passed to the shading functions used for inference. Will be
calculated (and overwritten if specified) if both expected and residuals are
NULL, or if expected is given a formula.
condvars number of conditioning variables, if any; those are expected to be ordered first
in the table. This in
</code></pre>
#### Format 
ion is used for computing the expected 
#### Value 
, and is
also passed to the spacing functions (see spacings).
shade logical specifying whether gp should be used or not (see gp). If TRUE and
expected is unspecified, a default model is fitted: if condvars is specified, a
corresponding conditional independence model, and else the total independence
model.
residuals_type a character string indicating the type of residuals to be computed when none are
supplied. If residuals is NULL, residuals_type must be one of "pearson"
(default; giving components of Pearson's chi-squared), "deviance" (giving components
of the likelihood ratio chi-squared), or "FT" for the Freeman-Tukey
residuals. The value of this argument can be abbreviated. If residuals are specified, the value of residuals_type is just passed "as is" to the legend function.
* type a character string indicating whether the observed or the expected 
#### Value 
 of the
table should be visualized.
split_vertical vector of logicals of length k, where k is the number of margins of x (
#### Value 

are recycled as needed). A TRUE component indicates that the tile(s) of the corresponding dimension should be split vertically, FALSE means horizontal splits.
Default is FALSE.
* spacing spacing object, spacing function, or a corresponding generating function (see
details and spacings).
spacing_args list of arguments  for the spacing-generating function, if specified.
gp object of class "gpar", shading function or a corresponding generating function (see details and shadings). Components of "gpar" objects are recycled as
needed along the last splitting dimension. Ignored if shade = FALSE.
* gp_args list of arguments for the shading-generating function, if specified.
* labeling either a logical, or a labeling function, or a corresponding generating function (see details and labelings. If FALSE or NULL, no labeling is produced.
strucplot 117
* labeling_args list of arguments for the labeling-generating function, if specified.
* core either a core function, or a corresponding generating function (see details). Currently, generating functions for mosaic plots (struc_mosaic), association plots
(struc_assoc), and sieve plots (struc_sieve) are provided.
core_argslist of arguments 
 for the core-generating function, if specified.
legend either a legend-generating function, or a legend function (see details and legends), or a logical. If legend is NULL or TRUE and gp is a function, legend defaults to legend_resbased.
legend_args list of arguments for the legend-generating function, if specified.
main either a logical, or a character string used for plotting the main title. If main is a logical and TRUE, the name of the object supplied as x is used.
* sub a character string used for plotting the subtitle. If sub is a logical and TRUE and main is unspecified, the name of the object supplied as x is used.
* margins either an object of class "unit" of length 4, or a numeric vector of length 4. The elements are recycled as needed. The four components specify the top,right, bottom, and left margin of the plot, respectively. When a numeric vector
is supplied, the numbers are interpreted as "lines" units. In addition, the
unit or numeric vector may have named 
</code></pre>
#### Arguments
 ('top', 'right', 'bottom',
and 'left'), in which case the non-named 
</code></pre>
#### Arguments
 specify the default 
Value 

(recycled as needed), overloaded by the named arguments
.
title_margins either an object of class "unit" of length 2, or a numeric vector of length 2. The
elements are recycled as needed. The two components specify the top and bottom
title margin of the plot, respectively. The default for each specified title are 2 lines (and 0 else), except when a legend is plotted and keep_aspect_ratio
is TRUE: in this case, the default 
#### Value 
 of both margins are set as to align the
heights of legend and actual plot. When a numeric vector is supplied, the numbers
are interpreted as "lines" units. In addition, the unit or numeric vector may
have named arguments
 ('top' and 'bottom'), in which case the non-named argument
specify the default value (recycled as needed), overloaded by the named

</code></pre>
#### Arguments
.
legend_width An object of class "unit" of length 1 specifying the width of the legend (if any).
Default: 5 lines.
pop logical indicating whether the generated viewport tree should be removed at the
end of the drawing or not.
main_gp, sub_gp
object of class "gpar" containing the graphical parameters used for the main
(sub) title, if specified.
newpage logical indicating whether a new page should be created for the plot or not.
return_grob logical. Should a snapshot of the display be returned as a grid grob?
keep_aspect_ratio
logical indicating whether the aspect ratio should be fixed or not. If unspecified,
the default is TRUE for two-dimensional tables and FALSE otherwise.
prefix optional character string used as a prefix for the generated viewport and grob
names.
... For convenience, list of 
</code></pre>
#### Arguments
 passed to the labeling-generating function
used.
118 strucplot
Details
This function-usually called by higher-level functions such as assoc and mosaic-generates conditioning
plots of contingency tables. First, it sets up a set of viewports for main- and subtitles,
legend, and the actual plot region. Then, residuals are computed as needed from observed and
expected frequencies, where the expected frequencies are optionally computed for a specified independence
model. 


Finally, the specified functions for spacing, gp, main plot, legend, and labeling
are called to produce the plot. The function invisibly returns the "structable" object visualized.
Most elements of the plot, such as the core function, the spacing between the tiles, the shading
of the tiles, the labeling, and the legend, are modularized in graphical appearance control ("grapcon")
functions and specified as parameters. For each element foo (= spacing, labeling, core,
or legend), strucplot takes two 
</code></pre>
#### Arguments
: foo and foo\_args, which can be used to specify the
parameters in the following alternative ways:
1. Passing a suitable function to foo which subsequently will be called from strucplot to compute
shadings, labelings, etc.
2. Passing a corresponding generating function to foo, along with parameters passed to foo\_args,
that generates such a function. Generating functions must inherit from classes "grapcon_generator"
and "}foo\code{".
3. Except for the shading functions (shading\_bar), passing foo(foo\_args) to the foo argument.
4. For shadings and spacings, passing the final parameter object itself; see the corresponding
help pages for more details on the data structures.
If legends are drawn, a 'cinemascope'-like layout is used for the plot to preserve the 1:1 aspect ratio.
If type = "expected", the expected 
#### Value 
 are passed to the observed argument of the core
function, and the observed 
#### Value 
 to the expected argument.
Although the gp argument is typically used for shading, it can be used for arbitrary modifications
of the tiles' graphics parameters (e.g., for highlighting particular cells, etc.).
Value
Invisibly, an object of class "structable" corresponding to the plot. If return_grob is TRUE,
additionally, the plot as a grob object is returned in a grob attribute.
Note
The created viewports, as well as the tiles and bullets, are named and thus can conveniently modified
after a plot has been drawn (and pop = FALSE).
Author(s)
David Meyer <David.Meyer@R-project.org>
References
Meyer D., Zeileis A., and Hornik K. (2006), The strucplot framework: Visualizing multi-way contingency
tables with vcd. Journal of Statistical Software, 17(3), 1-48. URL http://www.jstatsoft.org/v17/i03/
and available as vignette("strucplot").
structable 119
See Also
assoc, mosaic, sieve, struc_assoc, struc_sieve, struc_mosaic, structable, doubledecker,
labelings, shadings, legends, spacings

#### Examples
```{r}
library(vcd)

data("Titanic")
strucplot(Titanic)
```

```{r}
strucplot(Titanic, core = struc_assoc)
```

```{r}
strucplot(Titanic, spacing = spacing_increase,
spacing_args = list(start = 0.5, rate = 1.5))
```

```{r}
strucplot(Titanic, spacing = spacing_increase(start = 0.5, rate = 1.5))
```

```{r}
## modify a tile's color
strucplot(Titanic, pop = FALSE)
grid.edit("rect:Class=1st,Sex=Male,Age=Adult,Survived=Yes",
gp = gpar(fill = "red"))

```
