---
title: "``struc_mosaic``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcd)
```

### Core-generating Function for Mosaic Plots

#### Desription

Core-generating function for strucplot returning a function producing mosaic plots.

#### Usage
<pre><code>

struc_mosaic(zero_size = 0.5, zero_split = FALSE, zero_shade = TRUE,
zero_gp = gpar(col = 0), panel = NULL)

</code></pre>
#### Arguments

* zero_size size of the bullets used for zero-entries in the contingency table (if 0, no bullets
are drawn).
* zero_split logical controlling whether zero cells should be further split. If FALSE and
zero_shade is FALSE, only one bullet is drawn (centered) for unsplit zero cells.
If FALSE and zero_shade is TRUE, a bullet for each zero cell is drawn to allow,
e.g., residual-based shadings to be effective also for zero cells.
zero_shade logical controlling whether zero bullets should be shaded.
* zero_gp object of class "gpar" used for zero bullets in case they are not shaded.
* panel Optional function with  Arguments
: residuals, observed, expected, index,
gp, and name called by the struc_mosaic workhorse for each tile that is drawn
in the mosaic. index is an integer vector with the tile’s coordinates in the contingency
table, gp a gpar object for the tile, and name a label to be assigned to
the drawn grid object.


#### Examples
```{r}

## Titanic data
data("Titanic")
## mosaic plot with large zeros
strucplot(Titanic, core = struc_mosaic(zero_size = 1))



Details
This function is usually called by strucplot (typically when called by mosaic) and returns a
function used by strucplot to produce mosaic plots.
Value
A function with Arguments
:
residuals table of residuals.
observed table of observed 
#### Value 
.
expected not used by struc_mosaic.
spacing object of class "unit" specifying the space between the tiles.
gp list of gpar objects used for the drawing the tiles.
split_vertical vector of logicals indicating, for each dimension of the table, the split direction.

Author(s)
David Meyer <David.Meyer@R-project.org>
References
Cohen, A. (1980), On the graphical display of the significant components in a two-way contingency
table. Communications in Statistics—Theory and Methods, A9, 1025–1041.
Friendly, M. (1992), Graphical methods for categorical data. SAS User Group International Conference
Proceedings, 17, 190–200. http://datavis.ca/papers/sugi/sugi17.pdf
Meyer, D., Zeileis, A., and Hornik, K. (2006), The strucplot framework: Visualizing multi-way contingency
tables with vcd. Journal of Statistical Software, 17(3), 1-48. URL http://www.jstatsoft.org/v17/i03/
and available as vignette("strucplot").
See Also
mosaic, strucplot, structable
