---
title: "vcd"
author: "Kevin O'Brien"
date: "28 December 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
tile Tile Plot

#### Desription

Plots a tile display.

#### Usage
<pre><code>

## Default S3 method:
tile(x,
tile_type = c("area", "squaredarea", "height", "width"),
halign = c("left", "center", "right"),
valign = c("bottom", "center", "top"),
tile 131
split_vertical = NULL,
shade = FALSE,
spacing = spacing_equal(unit(1, "lines")),
set_labels = NULL,
margins = unit(3, "lines"),
keep_aspect_ratio = FALSE,
legend = NULL,
legend_width = NULL,
squared_tiles = TRUE,
main = NULL, sub = NULL, ...)
## S3 method for class 'formula'
tile(formula, data,
..., main = NULL, sub = NULL, subset = NULL, na.action = NULL)

</code></pre>
#### Arguments

* ``x``: a contingency table, or an object coercible to one.
formula a formula specifying the variables used to create a contingency table from data.
data either a data frame, or an object of class "table" or "ftable".
subset an optional vector specifying a subset of observations to be used.
na.action a function which indicates what should happen when the data contain NAs. Ignored
if data is a contingency table.
tile_type character string indicating how the tiles should reflect the table frequencies (see
details).
halign, valign character string specifying the horizontal and vertical alignment of the tiles.
split_vertical vector of logicals of length k, where k is the number of margins of x (
#### Value 

are recycled as needed). A TRUE component indicates that the tile(s) of the corresponding
dimension should be split vertically, FALSE means horizontal splits.
Default is FALSE.
spacing spacing object, spacing function, or corresponding generating function (see strucplot
for more in
</code></pre>
#### Format 
ion).
set_labels An optional character vector with named components replacing the so-specified
variable names. The component names must exactly match the variable names
to be replaced.
shade logical specifying whether shading should be enabled or not (see strucplot).
margins either an object of class "unit" of length 4, or a numeric vector of length 4.
The elements are recycled as needed. The four components specify the top,
right, bottom, and left margin of the plot, respectively. When a numeric vector
is supplied, the numbers are interpreted as "lines" units. In addition, the
unit or numeric vector may have named 
</code></pre>
#### Arguments
 ('top', 'right', 'bottom',
and 'left'), in which case the non-named 
</code></pre>
#### Arguments
 specify the default 
#### Value 

(recycled as needed), overloaded by the named 
</code></pre>
#### Arguments
.
legend either a legend-generating function, or a legend function (see details and legends),
or a logical. If legend is NULL or TRUE and gp is a function or missing, legend
defaults to legend_resbased.
132 tile
legend_width An object of class "unit" of length 1 specifying the width of the legend (if any).
Default: 5 lines.
keep_aspect_ratio
logical indicating whether the aspect ratio should be fixed or not. The default is
FALSE to enable the creation of squared tiles.
squared_tiles logical indicating whether white space should be added as needed to rows or
columns to obtain squared tiles in case of an unequal number of row and column
labels.
main, sub either a logical, or a character string used for plotting the main (sub) title. If
logical and TRUE, the name of the data object is used.
... Other Arguments
 passed to strucplot

#### Examples
```{r}

data("Titanic")
## default plot
tile(Titanic)
tile(Titanic, type = "expected")
tile(Titanic, shade = TRUE)
## some variations
```
```{r}
tile(Titanic, tile_type = "squaredarea")
tile(Titanic, tile_type = "width", squared_tiles = FALSE)
tile(Titanic, tile_type = "height", squared_tiles = FALSE)
tile(Titanic, tile_type = "area", halign = "center", valign = "center")
## repeat levels
tile(Titanic[,,,c(1,2,1,2)])
```


#### Details
A tile plot is a matrix of tiles. For each tile, either the "width", "height", "area", or squared area
is proportional to the corresponding entry. The first three options allow column-wise, row-wise and
overall comparisons, respectively. The last variant allows to compare the tiles both column-wise
and row-wise, considering either the width or the height, respectively.
In contrast to other high-level strucplot functions, tile also accepts a table with duplicated levels
(see 
Examples

). In this case, artificial dimnames will be created, and the actual ones are drawn using
set_labels.
Note that multiway-tables are first "flattened" using structable.

#### Value
The "structable" visualized is returned invisibly.
Author(s)
David Meyer <David.Meyer@R-project.org>
See Also
assoc, strucplot, mosaic, structable,
