### Summary: The Cox Proportional Hazards Model

#### Introduction:
The Cox Proportional Hazards (Cox-PH) model is a regression model used for analyzing survival data and estimating the hazard function. Unlike non-parametric methods, Cox-PH can handle multiple variables simultaneously.

#### Key Concepts:

1. **Proportional Hazards**:
   - Assumes a baseline hazard function, $h_0(t)$, representing the hazard function of time.
   - Hazard function for covariates: \( h(t) = \psi \, h_0(t) \), where $\psi$ is a constant factor.
   - $\psi$ is modeled as a linear function: \( \psi = \exp(\beta \, X) \).

2. **Cox-PH Model**:
   - Full model: \( h(t | X) = h_0(t) \exp(\beta \, X) \).
   - Uses partial likelihood methods to account for censored data.
   
3. **Model Assessment**:
   - **R-Squared**: Modified for survival models to measure variance explained.
   - **Concordance Index**: Measures the proportion of pairwise comparisons correctly identified by the model.

4. **Building and Assessing Models**:
   - **Initial Model (Voicemail)**:
```{r}
telco_model01_coxph <- coxph(Surv(accountdur, churned) ~ vmailplan, data = telco_churn_tbl)
summary(telco_model01_coxph)
```
- Model 1 has a $c$-index of 0.565 and $R^2$ of 0.011.

   - **Adding International Plans**:
```{r}
telco_model02_coxph <- coxph(Surv(accountdur, churned) ~ intlplan, data = telco_churn_tbl)
summary(telco_model02_coxph)
```
- Model 2 has a $c$-index of 0.591, indicating better predictive power.

   - **Combining Variables**:
```{r}
telco_model03_coxph <- coxph(Surv(accountdur, churned) ~ vmailplan + intlplan, data = telco_churn_tbl)
summary(telco_model03_coxph)
```
- Model 3 has a $c$-index of 0.643 and $R^2$ of 0.05.

5. **Validating Proportional Hazards Assumption**:
   - **Null Model**:
```{r}
telco_model00_coxph <- coxph(Surv(accountdur, churned) ~ 1, data = telco_churn_tbl)
summary(telco_model00_coxph)
```
   - Uses residuals analysis to validate the model.

---

Let me know if you need further details or have additional questions on Cox-PH models!