Grubbs Procedure
========================================================

### Fuse Data Set
This Data set is the running times of 20 mechanical fuses measured 
by operators stopping two independent clocks.

```{r}
Fuse <- 1:20
###Measuerement by first Instrument
Y1 <- c(4.85,4.93,4.75,4.77,4.67,4.87,4.67,4.94,4.85,4.75,
        4.83,4.92,4.74,4.99,4.88,4.95,4.95,4.93,4.92,4.89)
###Measuerement by second Instrument
Y2 <- c(5.09,5.04,4.95,5.02,4.90,5.05,4.90,5.15,5.08,4.98,
        5.04,5.12,4.95,5.23,5.07,5.23,5.16,5.11,5.11,5.08)
```
### Identity Plot
```{r fig.width=7, fig.height=7}
par(font.lab=2,font.axis=2,cex=1.7)
plot(Y1,Y2,pch=16,col="red",ylim=c(4.5,5.5),xlim=c(4.5,5.5))
title("Identity Plot - Grubbs's Fuse Data")
abline(a=0,b=1,col="blue",lwd=2)
```

### Preliminary Calculations / Bland-Altman Analysis
```{r}
Ds = Y2-Y1
Ss = Y2+Y1
```
According to conventionally used procedure, there is a significant inter-method bias between both methods.
```{r}
## Mean of Case-wise Differences
mean(Ds)
## Limits of Agreement
c(mean(Ds)-1.96*sd(Ds),mean(Ds)+1.96*sd(Ds))
## Paired t-test
t.test(Y2,Y1,paired=T)$p.value
```
### Using the BB procedure
Case-wise difference regressed on case-wise sums.
```{r}

FitBB <- lm(Ds ~ Ss)

summary(FitBB)$coefficients
anova(FitBB)
```
### Scatterplot of BB regression


```{r fig.width=10, fig.height=7}
par(font.lab=2,font.axis=2,cex=1.7)
plot(Ss,Ds,pch=16,col="red",ylab="Case-wise Differences",xlab="Case-wise Sums",ylim=c(-0.05,0.35))
title("Grubbs's Fuse Data")
abline(coef(FitBB),col="blue",lwd=2)
abline(h=0,col="green",lwd=1.5)
```

