---
title: "{modelr} R package"
susbtitl: Linear Models with R
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r include=FALSE, echo=FALSE}
library(knitr)
library(broom)
library(modelr)
library(magrittr)
library(faraway)
data(cheddar)
library(car)

fit <- lm(taste ~ Acetic + H2S + Lactic, data = cheddar)
```

{modelr}
=======================
Compute model quality for a given dataset

Three summaries are immediately interpretible on the scale of the response variable:

* ``rmse()`` is the root-mean-squared-error
* ``mae()`` is the mean absolute error
* ``qae()`` is quantiles of absolute error.



### Root Mean Square Error

```{r}
rmse(fit, data= cheddar)
```


### mean absolute error

```{r}
mae(fit, data= cheddar)
```



```{r}
qae(fit, data= cheddar)
```


### Other summaries

* ``mape()`` mean absolute percentage error.
* ``rsae()`` is the relative sum of absolute errors.
* ``mse()`` is the mean-squared-error.
* ``rsquare()`` is the variance of the predictions divided by the variance of the response.



```{r}
rsquare(fit, data= cheddar)
```