---
title: The {gvlma} package 
subtitle: Model Diagnostics for Linear Models
author: DragonflyStats.github.io
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE, echo=FALSE}
library(knitr)
library(broom)
library(magrittr)
library(faraway)
data(cheddar)

Fit_4 <- lm(taste ~ Acetic + H2S + Lactic, data = cheddar)
```

{gvlma}
========================

* The {gvlma} package is a comprehensive, automatic testing suite for many of the assumptions of general linear models. 
* It does both statistical tests and diagnostic plots using an extremely simple implementation for powerful results.




The package is an implementation of a paper by Pena & Slate called ***Global Validation of Linear Model Assumptions*** and allows you to quickly check for:

* ***Linearity*** - the Global Stat tests for the null hypothesis that our model is a linear combination of its predictors.

* ***Homoscedasticity*** - the respective stat tests for the null that the residial variance is relatively constant over the range of values.

* ***Normality*** - skewness and kurtosis tests help you understand if the residuals fits a normal distribution. 

* If the null is rejected you probably need to transform your data in some way (like a log transform). 
This can also be assessed by looking at the normal probability plot it generates.	

### ***<tt>gvlma( )</tt>***

The <tt>gvlma( )</tt> function in the ***{gvlma}*** package, performs a global validation of linear model assumptions as well separate evaluations of skewness, kurtosis, and heteroscedasticity.

```{r eval=FALSE}
# Global test of model assumptions

gvmodel <- gvlma(fit) 
summary(gvmodel)
```

### Example 

This examples uses the cheddar data set, available in the {faraway} R package.

```{r echo=FALSE,include=FALSE}
model <- Fit_4
```

```{r}
library("gvlma")

# model <- lm(y ~ x, data)
  
summary(gvlma(model))
```


### Diagnostic Plots for {gvlma}

* The diagnostic plots also let you understand the relation between your data and these assumptions visually. 
* Other useful capabilities are the link function test which is used for understanding whether the underlying data is categorical or continuous.





```{r eval=FALSE }

  plot(gvlma(model),which=1)

```





