---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---

### InspectDF

#### InspectDF

inspectdf: Inspection, Comparison and Visualisation of Data Frames

inspectdf is collection of utilities for columnwise summary, comparison and visualisation of data frames. 

Functions are provided to summarise missingness, categorical levels, numeric distribution, correlation, column types and memory usage. 

The package has three aims: to speed up repetitive checking and exploratory tasks for data frames


---


```{r echo=FALSE, include=FALSE}
library(dplyr)
library(inspectdf)
```

<pre><code>
library(dplyr)
library(inspectdf)
</code></pre>


---

#### InspectDF

Key functions

* ``inspect_types()`` -  summary of column types
* ``inspect_mem()`` -  summary of memory usage of columns
* ``inspect_na()`` -  columnwise prevalence of missing values
* ``inspect_cor()`` -  correlation coefficients of numeric columns
* ``inspect_imb()`` -  feature imbalance of categorical columns
* ``inspect_num()`` -  summaries of numeric columns
* ``inspect_cat()`` -  summaries of categorical columns

---

### InspectDF


```{r}
# Load dplyr for starwars data & pipe

library(dplyr)

# Single dataframe summary

inspect_cat(starwars)

```

---

### InspectDF


```{r}
library(dplyr)
# Paired dataframe comparison
inspect_cat(starwars, starwars[1:20, ])
```

---

### InspectDF

```{r}
# Grouped dataframe summary
starwars %>% group_by(species) %>% inspect_cat()
```

