
Boxplot Boxplots With Point Identification

#### Description

Boxplot is a wrapper for the standard R boxplot function, providing point identification, axis labels, and a formula interface for boxplots without a grouping variable.

#### Usage
<pre><code>
Boxplot(y, ...)

## Default S3 method:
Boxplot(y, g, id=TRUE, xlab, ylab, ...)

## S3 method for class 'formula'
Boxplot(formula, data=NULL, subset, na.action=NULL,id=TRUE, xlab, ylab, ...)

## S3 method for class 'list'
Boxplot(y, xlab="", ylab="", ...)

## S3 method for class 'data.frame'
Boxplot(y, id=TRUE, ...)

## S3 method for class 'matrix'
Boxplot(y, ...)

</code></pre>

#### Arguments

* `` y ``:   a numeric variable for which the boxplot is to be constructed; a list of numeric
* variables, each element of which will be treated as a group; a numeric data frame or a numeric matrix, each of whose columns will be treated as a group.
* ``g``: a grouping variable, usually a factor, for constructing parallel boxplots.
* ``id``: a list of named elements giving one or more specifications for labels of individual points ("outliers"): n, the maximum number of points to label (default 10); location, "lr" (left or right) of points or "avoid" to try to avoid overplotting; method, one of "y" (automatic, the default), "identify" (interactive), or "none"; col for labels (default is the first color in ``carPalette()`` ); and cex size of labels (default is 1). 
- Can be FALSE to suppress point identification or TRUE (the default) to use all defaults. This is similar to how showLabels handles
* ``point``: labels for other functions in the car package, except that the usual default is id=FALSE.
* ``xlab``, ``ylab``: text labels for the horizontal and vertical axes; if missing, Boxplot will use the variable names, or, in the case of a list, data frame, or matrix, empty labels.
* formula a ‘model’ formula, of the form ~ y to produce a boxplot for the variable y, or of the form y ~ g, ``y ~ g1*g2*...``, or ``y ~ g1 + g2 + ...`` to produce parallel boxplots for y within levels of the grouping variable(s) g, etc., usually factors.
* data, subset, na.action as for statistical modeling functions (see, e.g., lm).
* ``...``: further arguments, such as at, to be passed to boxplot.



####Examples

```{r}

Boxplot(~income, data=Prestige, id=list(n=Inf)) # identify all outliers
Boxplot(income ~ type, data=Prestige)
```

```{r}
Boxplot(income ~ type, data=Prestige, at=c(1, 3, 2))
Boxplot(k5 + k618 ~ lfp*wc, data=Mroz)
```

```{r}
with(Prestige, Boxplot(income, id=list(labels=rownames(Prestige))))
with(Prestige, Boxplot(income, type, id=list(labels=rownames(Prestige))))
```

```{r}
Boxplot(scale(Prestige[, 1:4]))
```
#### Author(s)
John Fox <jfox@mcmaster.ca>, with a contribution from Steve Ellison to handle at argument (see
boxplot).

#### References

Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.

#### See Also
boxplot
