---
title: "``hist.boot``:  Methods Functions to Support boot Objects"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```



####Description

The Boot function in the car package uses the boot function from the boot package to do a straightforward
case or residual bootstrap for many regression objects. These are method functions for standard
generics to summarize the results of the bootstrap. Other tools for this purpose are available in
the boot package.

####Usage
<pre><code>

## S3 method for class 'boot'
hist(x, parm, layout = NULL, ask, main = "", freq = FALSE,
estPoint = TRUE, point.col = carPalette()[1], point.lty = 2, point.lwd = 2,
estDensity = !freq, den.col = carPalette()[2], den.lty = 1, den.lwd = 2,
estNormal = !freq, nor.col = carPalette()[3], nor.lty = 2, nor.lwd = 2,
ci = c("bca", "none", "perc", "norm"), level = 0.95,
legend = c("top", "none", "separate"), box = TRUE, ...)
## S3 method for class 'boot'
summary(object, parm, high.moments = FALSE, extremes = FALSE, ...)
## S3 method for class 'boot'
confint(object, parm, level = 0.95, type = c("bca", "norm",
"basic", "perc"), ...)
hist.boot 57
## S3 method for class 'boot'
Confint(object, parm, level = 0.95, type = c("bca", "norm",
"basic", "perc"), ...)
## S3 method for class 'boot'
vcov(object, ...)

</code></pre>

####Arguments

x, object An object created by a call to boot in the boot package, or to Boot in the car
package of class "boot".
parm A vector of numbers or coefficient names giving the coefficients for which a
histogram or confidence interval is desired. If numbers are used, 1 corresponds
to the intercept, if any. The default is all coefficients.
* ``layout`` :  If set to a value like c(1, 1) or c(4, 3), the layout of the graph will have
this many rows and columns. If not set, the program will select an appropriate
layout. If the number of graphs exceed nine, you must select the layout yourself,
or you will get a maximum of nine per page. If layout=NA, the function does
not set the layout and the user can use the par function to control the layout, for
example to have plots from two models in the same graphics window.
ask If TRUE, ask the user before drawing the next plot; if FALSE, don’t ask.
main Main title for the graphs. The default is main="" for no title.
freq The default for the generic hist function is freq=TRUE to give a frequency histogram.
The default for hist.boot is freq=FALSE to give a density histogram.
A density estimate and/or a fitted normal density can be added to the graph if
freq=FALSE but not if freq=TRUE.
estPoint, point.col, point.lty, point.lwd
If estPoint=TRUE, the default, a vertical line is drawn on the histgram at the
value of the point estimate computed from the complete data. The remaining
three optional arguments set the color, line type and line width of the line that is
drawn.
estDensity, den.col, den.lty, den.lwd
If estDensity=TRUE andfreq=FALSE, the default, a kernel density estimate is
drawn on the plot with a call to the density function with no additional arguments.
The remaining three optional arguments set the color, line type and line
width of the lines that are drawn.
estNormal, nor.col, nor.lty, nor.lwd
If estNormal=TRUE andfreq=FALSE, the default, a normal density with mean
and sd computed from the data is drawn on the plot. The remaining three optional
arguments set the color, line type and line width of the lines that are drawn.
ci A confidence interval based on the bootstrap will be added to the histogram
using the BCa method if ci="bca" the percentile method if ci="perc", or the
normal method if ci="norm". No interval is drawn if ci="none". The default
is "bca". The interval is indicated by a thick horizontal line at y=0. For some
bootstraps the BCa method is unavailable, in which case a warning is issued and
ci="perc" is substituted. If you wish to see all the options at once, see boot.ci.
The normal method is computed as the (estimate from the original data) minus
58 hist.boot
the bootstrap bias plus or minus the standard deviation of the bootstrap replicates
times the appropriate quantile of the standard normal distribution.
legend A legend can be added to the (array of) histograms. The value “"top"” puts at
the top-left of the plots. The value “"separate"” puts the legend in its own graph
following all the histograms. The value “"none"” suppresses the legend.
box Add a box around each histogram.
... Additional arguments passed to hist; for other methods this is included for
compatibility with the generic method. For example, the argument border=par()$bg
in hist will draw the histogram transparently, leaving only the density estimates.
With the vcov function, the additional arguments are passed to cov. See
the Value section, below.
high.moments Should the skewness and kurtosis be included in the summary? Default is
FALSE.
extremes Should the minimum, maximum and range be included in the summary? Default
is FALSE.
level Confidence level, a number between 0 and 1. In confint, level can be a vector;
for example level=c(.50, .90, .95) will return the following estimated
quantiles: c(.025, .05, .25, .75, .95, .975).
type Selects the confidence interval type. The types implemented are the "percentile"
method, which uses the function quantile to return the appropriate quantiles
for the confidence limit specified, the default bca which uses the bias-corrected
and accelerated method presented by Efron and Tibshirani (1993, Chapter 14).
For the other types, see the documentation for boot.
Value
hist is used for the side-effect of drawing an array of historgams of each column of the first argument.
summary returns a matrix of summary statistics for each of the columns in the bootstrap
object. The confint method returns confidence intervals. Confint appends the estimates based on
the original fitted model to the left of the confidence intervals.
The vcov returns the sample covariance of the bootstrap sample estimates. If any of the bootstrap
replications returned an NA value, then vcov will return a matrix of NAs. You can compute the
covariance of the complete cases by adding the argument use="complete.obs" to the call to vcov;
see cov.
#### Author(s)
Sanford Weisberg, <sandy@umn.edu>

####References

Efron, B. and Tibsharini, R. (1993) An Introduction to the Bootstrap. New York: Chapman and
Hall.
Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition. Thousand
Oaks: Sage.

* Fox, J. and Weisberg, S. (2017) Bootstrapping, http://socserv.mcmaster.ca/jfox/Books/Companion/
appendix/Appendix-Bootstrapping.pdf.
Weisberg, S. (2013) Applied Linear Regression, Fourth Edition, Wiley
See Also
See Also Boot, hist, density, Fox and Weisberg (2017), cited above

####Examples

```{r}

m1 <- lm(Fertility ~ ., swiss)
betahat.boot <- Boot(m1, R=99) # 99 bootstrap samples--too small to be useful
summary(betahat.boot) # default summary
confint(betahat.boot)
hist(betahat.boot)
```

