---
title: "``outlierTest``:  Bonferroni Outlier Test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```




####Description

Reports the Bonferroni p-values for testing each observation in turn to be a mean-shift outliner,
based Studentized residuals in linear (t-tests) and generalized linear models (normal tests).

####Usage
<pre><code>

outlierTest(model, ...)
## S3 method for class 'lm'
outlierTest(model, cutoff=0.05, n.max=10, order=TRUE,
labels=names(rstudent), ...)
## S3 method for class 'outlierTest'
print(x, digits=5, ...)

</code></pre>

####Arguments

model an lm or glm model object.
cutoff observations with Bonferonni p-values exceeding cutoff are not reported, unless
no observations are nominated, in which case the one with the largest Studentized
residual is reported.

n.max maximum number of observations to report (default, 10).
order report Studenized residuals in descending order of magnitude? (default, TRUE).
labels an optional vector of observation names.
... arguments passed down to methods functions.
x outlierTest object.
digits number of digits for reported p-values.


####Examples

```{r}

outlierTest(lm(prestige ~ income + education, data=Duncan))
```
####Details

For a linear model, p-values reported use the t distribution with degrees of freedom one less than the
residual df for the model. For a generalized linear model, p-values are based on the standard-normal
distribution. The Bonferroni adjustment multiplies the usual two-sided p-value by the number of
observations. The lm method works for glm objects. To show all of the observations set cutoff=Inf
and n.max=Inf.
Value
an object of class outlierTest, which is normally just printed.
#### Author(s)
John Fox <jfox@mcmaster.ca> and Sanford Weisberg

####References

Cook, R. D. and Weisberg, S. (1982) Residuals and Influence in Regression. Chapman and Hall,
https://conservancy.umn.edu/handle/11299/37076.
* Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.
Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.
* Weisberg, S. (2014) Applied Linear Regression, Fourth Edition, Wiley.
Williams, D. A. (1987) Generalized linear model diagnostics using the deviance and single case
deletions. Applied Statistics 36, 181-191.
