---
title: "``scatterplotMatrix``:  Scatterplot Matrices"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```




#### Description

This function provides a convenient interface to the pairs function to produce enhanced scatterplot
matrices, including univariate displays on the diagonal and a variety of fitted lines, smoothers,
variance functions, and concentration ellipsoids. spm is an abbreviation for scatterplotMatrix.

#### Usage

<pre><code>

scatterplotMatrix(x, ...)
## S3 method for class 'formula'
scatterplotMatrix(formula, data=NULL, subset, ...)
## Default S3 method:
scatterplotMatrix(x, smooth = TRUE,
id = FALSE, legend = TRUE, regLine = TRUE,
ellipse = FALSE, var.labels = colnames(x), diagonal = TRUE,
plot.points = TRUE, groups = NULL, by.groups = TRUE,
use = c("complete.obs", "pairwise.complete.obs"), col =
carPalette()[-1], pch = 1:n.groups, cex = par("cex"),
cex.axis = par("cex.axis"), cex.labels = NULL,
cex.main = par("cex.main"), row1attop = TRUE, ...)
spm(x, ...)

</code></pre>

#### Examples

```{r}

scatterplotMatrix(~ income + education + prestige | type, data=Duncan)
scatterplotMatrix(~ income + education + prestige | type, data=Duncan,
regLine=FALSE, smooth=list(spread=FALSE))
scatterplotMatrix(~ income + education + prestige,
data=Duncan, id=TRUE, smooth=list(method=gamLine))
```

#### Arguments
* x a data matrix or a numeric data frame.
* formula a one-sided “model” formula, of the form ~ x1 + x2 + ... + xk or ~ x1 + x2 + ... + xk | z where z evaluates to a factor or other variable to divide the data into groups.
data for scatterplotMatrix.formula, a data frame within which to evaluate the
formula.

* subset expression defining a subset of observations.
* smooth specifies a nonparametric estimate of the mean or median function of the vertical axis variable given the horizontal axis variable and optionally a nonparametric estimate of the spread or variance function. If smooth=FALSE neither function
is drawn. If smooth=TRUE, then both the mean function and variance funtions are drawn for ungrouped data, and the mean function only is drawn for grouped data. The default smoother is loessLine, which uses the loess function from
the stats package. This smoother is fast and reliable. See the details below
for changing the smoother, line type, width and color, of the added lines, and
adding arguments for the smoother.
id controls point identification; if FALSE (the default), no points are identified; can be a list of named arguments to the showLabels function; TRUE is equivalent to list(method="mahal", n=2, cex=1, location="lr"), which identifies the
2 points (in each group, if by.groups=TRUE) with the largest Mahalanobis distances
from the center of the data; list(method="identify") for interactive
point identification is not allowed.
* legend controls placement of a legend if the plot is drawn by groups; if FALSE, the legend is suppressed. Can be a list with the named elementcoords specifying the position of the legend in any form acceptable to the legend function; TRUE (the
default) is equivalent to list(coords=NULL), for which placement will vary by
the the value of the diagonal argument—e.g., "topright" for diagonal=TRUE.
* regLine controls adding a fitted regression line to each plot, or to each group of points if
by.groups=TRUE. If regLine=FALSE, no line is drawn. This argument can also
be a list with named list, with default regLine=TRUE equivalent to regLine = list(method=lm, lty=1, specifying the name of the function that computes the line, with line type 1 (solid) of relative line width 2 and the color equal to the first value in the argument
col. Setting method=MASS::rlm would fit using a robust regression.

* ellipse controls plotting data-concentration ellipses. If FALSE (the default), no ellipses are plotted. Can be a list of named values giving levels, a vector of one or more bivariate-normal probability-contour levels at which to plot the ellipses; robust,
a logical value determing whether to use the cov.trob function in the MASS
package to calculate the center and covariance matrix for the data ellipses; and
fill and fill.alpha, which control whether the ellipse is filled and the transparency
of the fill. TRUE is equivalent to list(levels=c(.5, .95), robust=TRUE, fill=TRUE, fill.var.labels variable labels (for the diagonal of the plot).
diagonal contents of the diagonal panels of the plot. If diagonal=TRUE adaptive kernel
density estimates are plotted, separately for each group if grouping is present.
diagonal=FALSE suppresses the diagonal entries. See details below for other
choices for the diagonal.
plot.points if TRUE the points are plotted in each off-diagonal panel.
groups a factor or other variable dividing the data into groups; groups are plotted with
different colors and plotting characters.
by.groups if TRUE, the default, regression lines and smooths are fit by groups.
use if "complete.obs" (the default), cases with missing data are omitted; if "pairwise.complete.obs"), all in each panel of the plot.
pch plotting characters for points; default is the plotting characters in order (see par).
col colors for points; the default is carPalette starting at the second color. The
color of the regLine and smooth are the same as for points but can be changed
using the the regLine and smooth arguments.
cex relative size of plotted points
cex.axis relative size of axis labels
cex.labels relative size of labels on the diagonal
cex.main relative size of the main title, if any
row1attop If TRUE (the default) the first row is at the top, as in a matrix, as opposed to at
the bottom, as in graph (argument suggested by Richard Heiberger).
... arguments to pass down.

#### Details

Many arguments to scatterplotMatrix were changed in version 3 of car, to simplify use of this
function.
The smooth argument is usually either set to TRUE or FALSE to draw, or omit, the smoother. Alternatively
smooth can be set to a list of arguments. The default behavior of smooth=TRUE is equivalent to
smooth=list(smoother=loessLine, spread=TRUE, lty.smooth=1, lwd.smooth=1.5, lty.spread=3, lwd.spread=1),
specifying the smoother to be used, including the spread or variance smooth, and the line widths
and types for the curves. You can also specify the colors you want to use for the mean and variance
smooths with the arguments col.smooth and col.spread. Alternative smoothers are gamline
which uses the gam function from the mgcv package, and quantregLine which uses quantile regression
to estimate the median and quartile functions using rqss from the quantreg package. All of these smoothers have one or more arguments described on their help pages, and these arguments can be added to the smooth argument; for example, smooth = list(span=1/2) would use the
default loessLine smoother, include the variance smooth, and change the value of the smoothing
parameter to 1/2. For loessLine and gamLine the variance smooth is estimated by separately
smoothing the squared positive and negative residuals from the mean smooth, using the same type
of smoother. The displayed curves are equal to the mean smooth plus the square root of the fit to the positive squared residuals, and the mean fit minus the square root of the smooth of the negative squared residuals. The lines therefore represent the comnditional variabiliity at each value on the horizontal axis. Because smoothing is done separately for positive and negative residuals, the variation shown will generally not be symmetric about the fitted mean function. For the quantregLine
method, the center estimates the median for each value on the horizontal axis, and the spread estimates the lower and upper quartiles of the estimated conditional distribution for each value of the horizontal axis.
The sub-arguments spread, lty.spread and col.spread of the smooth argument are equivalent to the newer var, col.var and lty.var, respectively, recognizing that the spread is a measuure of conditional variability.
By default the diagonal argument is used to draw kernel density estimates of the variables by setting
diagonal=TRUE, which is equivalent to setting diagonal = list(method="adaptiveDensity", bw=bw.nrd0, adjust=The additional arguments shown are descibed at adaptiveKernel. The other methods avaliable, with their default arguments, are diagonal=list(method="density", bw="nrd0", adjust=1, kernel="gaussian", na.which uses density for nonadaptive kernel density estimation; diagonal=list(method ="histogram", breaks="FD")
which uses hist for drawing a histogram that ignores grouping, if present; diagonal=list(method="boxplot")

with no additional arguments which draws (parallel) boxplots; diagonal=list(method="qqplot")
with no additional arguments which draws a normal QQ plot; and diagonal=list(method="oned")
with no additional arguments which draws a rug plot tilted to the diagonal, as suggested by Richard
Heiberger.
Earlier versions of scatterplotMatrix included arguments transform and family to estimate
power transformations using the powerTransform function before drawing the plot. The same
functionality can be achieved by calling powerTransform directly to estimate a transformation,
saving the transformed variables, and then plotting.
Value
NULL, returned invisibly. This function is used for its side effect: producing a plot. If point identification
is used, a vector of identified points is returned.
#### Author(s)
John Fox <jfox@mcmaster.ca>

#### References

Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.
See Also
pairs, scatterplot, dataEllipse, powerTransform, bcPower, yjPower, cov.trob, showLabels,
ScatterplotSmoothers.

