---
title: "``whichNames`` Position of Row Names"
author: "Kevin O Brien"
date: "22 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(car)
```



####Description

These functions return the indices of the supplied row names of a data frame or names of another
object, such as a vector or list. whichNames is just an alias for which.names.

####Usage
<pre><code>

whichNames(names, object, ...)
which.names(names, object, ...)
## S3 method for class 'data.frame'
whichNames(names, object, ...)
## Default S3 method:
whichNames(names, object, ...)

</code></pre>

####Arguments

names a name or character vector of names.
object a data frame or an object with a names attribute.
... not used.
Value
Returns the index or indices of names in row names of the data frame or names of an object of
another class.
#### Author(s)
John Fox <jfox@mcmaster.ca>
142 whichNames

####References

Fox, J. and Weisberg, S. (2019) An R Companion to Applied Regression, Third Edition, Sage.

####Examples

```{r}

whichNames(c('minister', 'conductor'), Duncan)
```